\documentclass[main.tex]{subfiles}
\begin{document}
\section{The Completion}
\label{S:completion}
\subsection{Introduction}
\noindent
Valuation systems
were introduced in
 Section~\ref{S:valuation-systems}
to give meaning to the phrase 
``the Lebesgue integral $\Lphi$ is a completion of~$\Sphi$'',
namely,
\begin{equation*}
\text{$\vsLF$ is complete}\qquad\text{and}\qquad
\text{$\Lphi$ extends~$\Sphi$}.
\end{equation*}
In this section we replace~$\Sphi$
and~$\smash{\E^\R}$
by any valuation system $\vs{V}{L}\varphi{E}$
and study when, so to say, $\varphi$ has 
\emph{a completion with respect to~$V$},
i.e., when there is a sublattice~$C$ of~$V$
and a valuation~$\psi\colon C\ra E$
such that 
\begin{equation*}
\vs{V}{C}\psi{E}\text{ is complete}
\qquad\text{and}\qquad
\text{$\psi$ extends~$\varphi$}.
\end{equation*}
There is not always a completion (see Example~\ref{E:no-completion}).
However,
if~$\varphi$ has a completion with respect to~$V$,
then~$\varphi$ also has a smallest\footnote{``Smallest''
with respect to the ordering on partial functions 
given by $f\leq g$ iff $g$ extends~$f$.} %
completion
with respect to~$V$,
\begin{equation*}
\vs{V}{\ol{L}}{\ol\varphi}{E}.
\end{equation*}
We describe $\ol\varphi$
in Subsection~\ref{SS:hierarchy},
and call~$\ol\varphi$ \emph{\underline{the} completion} of~$\varphi$
with respect to~$V$.

\subsubsection*{$\Pi$-Completion}
If we replace ``complete'' by ``$\Pi$-complete''
in the above discussion
the situation is much easier
(see Definition~\ref{D:system-complete}).
Indeed,
we will see  in Lemma~\ref{L:Pi-continuity}
that~$\varphi$ has a $\Pi$-completion with respect
to~$V$ iff
\begin{equation*}
\label{eq:Pi-cont}
\left[ \quad
\begin{minipage}{.7\columnwidth}
for every $b\in L$ and $\varphi$-convergent 
 $a_1 \geq a_2 \geq \dotsb$
we have
\begin{equation*}
\bw_n a_n \leq b
\quad\implies\quad
\bw_n \varphi (a_n) \leq \varphi(b).
\end{equation*}
\end{minipage}
\right.
\end{equation*}
Similar to before, if there is a $\Pi$-completion
of~$\varphi$ with respect to~$V$,
then there is also a smallest $\Pi$-completion
with respect to~$V$,
which will be denoted by
\begin{equation*}
\vs{V}{\Pi L}{\Pi \varphi}{E}.
\end{equation*}
We will study~$\Pi\varphi$ in Subsection~\ref{S:Pi-extension}.

\subsubsection*{$\Sigma$-Completion}
We can replace ``complete'' by ``$\Sigma$-complete''
as well.\\
If it exists,
the smallest $\Sigma$-completion of~$\varphi$
with respect to~$V$ is denoted by
\begin{equation*}
\vs{V}{\Sigma L}{\Sigma \varphi}{E}.
\end{equation*}
Since~$\Pi \varphi$ and~$\Sigma\varphi$
are very similar,
we will only study~$\Pi\varphi$.
All the results that we obtain about~$\Pi\varphi$
and all the definitions  for~$\Pi\varphi$
can be easily translated
to results about~$\Sigma\varphi$
and definitions for~$\Sigma\varphi$,
respectively.
We leave this to the reader.

\subsubsection*{Hierarchy of Extensions}
Recall that~$\varphi$ is complete with respect to~$V$
if and only if~$\varphi$
is both $\Pi$-complete
and $\Sigma$-complete with respect to~$V$
(see Definition~\ref{D:system-complete}).

So if we want to find a completion of~$\varphi$
with respect to~$V$ it is natural to try and see if $\Sigma\Pi\varphi$
is complete 
(when it exists) with respect to~$V$.
Unfortunately,
while $\Pi\varphi$ is $\Pi$-complete
with respect to~$V$,
the valuation
$\Sigma\Pi\varphi$ need not be $\Pi$-complete.
Nevertheless,
we can continue to apply ``$\Sigma$'' and ``$\Pi$''
whenever possible,
and we will see in Subsection~\ref{SS:hierarchy-abstract}
that this leads to a `hierarchy' of the following shape.
\begin{equation*}
\xymatrix @=10pt {
& \Sigma\varphi \ar @{-} [rr] \ar @{-} [rrdd]
&& \Sigma \Pi \varphi  \ar @{-} [rr] \ar @{-} [rrdd]
&& \Sigma \Pi \Sigma\varphi  \ar @{-} [rr] \ar @{-} [rrdd]
&& \dotsb
\\  
\varphi \ar @{-} [ru] \ar @{-} [rd] 
&&&&&&&\dotsb\\
& \Pi\varphi \ar @{-} [rr] \ar @{-} [rruu]
&& \Pi \Sigma \varphi \ar @{-} [rr]\ar @{-} [rruu]
&& \Pi \Sigma \Pi \varphi \ar @{-} [rr]\ar @{-} [rruu]
&& \dotsb
}
\end{equation*}
It will become clear,
that if the making of this hierarchy is hindered, e.g.,
if $\Pi\Sigma\varphi$ has no $\Sigma$-completion
with respect to~$V$
(and so $\Sigma\Pi\Sigma\varphi$
does not exist),
then $\varphi$ cannot have a completion
with respect to~$V$.
On the other hand,
we will see, loosely speaking, 
that if the making of the hierarchy
can proceed
unhindered
even if we go on endlessly using ordinal numbers,
that then we eventually obtain
the smallest completion~$\ol\varphi$
of~$\varphi$ with respect to~$V$.
This will all become clear 
in Subsection~\ref{SS:hierarchy-abstract}.

With some luck, the valuation $\Sigma\Pi\Sigma \varphi$
might already be complete with respect to~$V$.
We could then say that ``we hit $\ol\varphi$ in $3$ steps''.
We will prove in Subsection~\ref{SS:borel-hierarchy}
that in general
we need to make \emph{uncountably} many steps
before we hit~$\ol\varphi$. {}


\subsubsection*{No Completion}
Let us end the introducion
with an example
of a valuation
system $\vs{V}{L}\varphi{E}$
such that~$\varphi$ has no completion with respect to~$V$.
\begin{ex}
\label{E:no-completion}
Let $L$ be the sublattice of~$\R$ given by
\begin{equation*}
L\ \eqdf\ \{ \ n^{-1}\colon\ n\in\N\ \} \,\cup\, \{0\}.
\end{equation*}
Let $\varphi\colon L \longrightarrow \R$
be the valuation given by, for all~$n\in\N$,
\begin{alignat*}{3}
\varphi(\,n^{-1}\,) \ &=\ 1&\qquad\text{and}\qquad
\varphi(\,0\,) \ &=\ 0.
\end{alignat*}
Then we have a valuation system $\vs{\R}{L}\varphi{\R}$.

Let $C$ be a sublattice of~$\R$,
and let $\psi\colon C \longrightarrow \R$
be the valuation given by
\begin{equation*}
\vs{\R}{C}\psi\R\text{ \ is complete }
\qquad\text{and}\qquad
\psi\text{ extends }\varphi.
\end{equation*}
We will prove that this is not possible.

Consider the sequence $a_1 \geq a_2 \geq \dotsb$ in~$L$
given by $a_n = n^{-1}$.
We have 
\begin{equation*}
\psi(a_n) \ =\  \varphi(a_n) \ =\  1.
\end{equation*}
So $\bw_n \psi(a_n) = 1$.
In particular,
$a_1 \geq a_2 \geq \dotsb$ is $\psi$-convergent
(see Definition~\ref{D:phi-conv}).
Since $\psi$ is complete with respect to~$\R$
(see Definition~\ref{D:system-complete}), 
we get
\begin{equation*}
1\ = \ 
\bw_n \psi( a_n ) \ = \  \psi(\, \bw_n a_n \,)
\ =\ \psi(\,\bw_n n^{-1}\,)
\ =\ \psi(\, 0\, ) \ =\  0,
\end{equation*}
which is absurd.
Hence $\varphi$ has no completion with respect to~$\varphi$.

\end{ex}

%
%                     PI L
%

\subsection{The $\Pi$-Extension}
\label{S:Pi-extension}
Let $\vs{V}{L}\varphi{E}$ be a valuation system.
Before we try to extend $\vs{V}{L}\varphi{E}$ to a complete
valuation system, 
we study whether $\vs{V}{L}\varphi{E}$
can be extended to a $\Pi$-complete valuation system.

\begin{dfn}
\label{D:PiL}
$\ \Pi L \ \eqdf \ \{ 
\bw_n a_n \colon \ \text{$\varphi$-convergent }
 a_1\geq a_2 \geq\dotsb \text{ from $L$} \}.$
\end{dfn}
%
%                  Pi(L) is lattice
%
\begin{lem}
\label{L:PiL-lattice}
$\Pi L$ is a sublattice of~$V$,
and $L$ is a sublattice of~$\Pi L$.
\end{lem}
\begin{proof}
We first prove that~$\Pi L$ is a sublattice of~$V$.
Let $a,b\in \Pi L$ be given;
we need to prove that $a\wedge b \in \Pi L$
and $a \vee b \in \Pi L$.
Choose $\varphi$-convergent 
$a_1 \geq a_2 \geq\dotsb$ and
$b_1 \geq b_2 \geq\dotsb$
with $a = \bw_n a_n$ and  $b=\bw_n b_n$.
Then $a_1 \wedge b_1 \geq a_2 \wedge b_2 \geq \dotsb$
is $\varphi$-convergent by Proposition~\ref{P:R-main},
and we have $\bw_n a_n \wedge b_n = a\wedge b$.
Hence $a\wedge b \in \Pi L$.
Similarly,
$a_1 \vee b_1 \geq a_2 \vee b_2 \geq \dotsb$
is $\varphi$-convergent by Proposition~\ref{P:R-main}
and
using $\sigma$-distributivity 
one can prove that $a \vee b  =\bw_n a_n \vee b_n$.
Hence $a\vee b \in \Pi L$.

To prove that~$L$ is a sublattice of~$\Pi L$,
we first note that~$L$ is a subset of~$\Pi L$.
Now,
since both $L$ and~$\Pi L$ are sublattices of~$V$,
and $L$ is a subset of~$\Pi L$,
we know that $L$ must be a sublattice of~$\Pi L$.
\end{proof}
\begin{rem}
In the proof of Lemma~\ref{L:PiL-lattice},
we have used the fact that~$V$ is $\sigma$-distrubutive
and the fact that~$E$ is $R$-complete
(via Proposition~\ref{P:R-main}).
\end{rem}

\begin{dfn}
\label{D:Pi-extendible}
We say~$\varphi$ is \keyword{$\Pi$-extendible}
if there is a valuation $\psi \colon \Pi L \ra E$ with
\begin{equation*}
\psi ( \bw_n a_n )
\ =\ 
\bw_n \varphi(a_n) 
\qquad
\text{ for all $\varphi$-convergent }
a_1 \geq a_2 \geq \dotsb.
\end{equation*}
Clearly,
there can be at most one such map~$\psi$;
if it exists, we denote it by 
\begin{equation*}
\Pi \varphi\colon \Pi L \ra E.
\end{equation*}
Finally,
note that if $\varphi$ is $\Pi$-extendible,
then $\Pi\varphi$ extends $\varphi$ (hence the name).
\end{dfn}
%
%                  PI EXTENDIBLE IMPLIES PI COMPLETE
%
\begin{lem}
\label{L:Pi-complete}
If $\varphi$ is $\Pi$-extendible,
then $\vs{V}{\Pi L}{\Pi\varphi}{E}$ is $\Pi$-complete.
\end{lem}
\begin{proof}
Let $a^1 \geq a^2 \geq \dotsb $ from $\Pi L$
be given and suppose  $\bw_n \Pi \varphi (a^n)$ exists;
we need to prove that $\bw_n a^n \in \Pi L$
and $\Pi\varphi(\bw_n a_n) = \bw_n \Pi \varphi(a_n )$.
To begin, write
$a^n = \bw_n a^n_m$
for each~$n$,
where $a^n_1 \geq a^n_2 \geq \dotsb$
is $\varphi$-convergent, and define
\begin{equation*}
b_i \ \eqdf\ \bw\{\,a^n_m\colon n,m\leq i \,\}.
\end{equation*}
Then $b_1 \geq b_2 \geq \dotsb$
and $\bw_n b_n = \bw_n a^n$.
Recall that  $\bw_n \Pi\varphi(a^n)$
exists.
We claim that $\bw_n \Pi\varphi(a^n)$
is the infimum of $\varphi b_1 \geq \varphi b_2 \geq \dotsb$.
If we can prove this, we are done.
Indeed,
then $b_1 \geq b_2 \geq\dotsb$ is $\varphi$-convergent,
so $\bw_n a^n = \bw_n b_n \in \Pi L$,
and
\begin{alignat*}{3}
\Pi\varphi( \bw_n a^n ) \ &=\  \Pi\varphi( \bw_n b_n )
 \qquad&& \text{since $\bw_n a^n = \bw_n b_n$}, \\
&=\ \bw_n \varphi (b_n)  
  && \text{since $\varphi$ is $\Pi$-extendible,}\\
&=\ \bw_n \Pi\varphi(a^n) 
  && \text{by the claim.}
\end{alignat*}

For each~$i$,
we have $b_i \geq \bw_{n\leq i} a^n = a^i$,
so $\varphi(b_i) = \Pi\varphi(b_i) \geq \Pi\varphi(a^i)$.
Hence we see that
$\bw_n \Pi\varphi(a^n)$ is a lower bound of
$\varphi(b_1)\geq \varphi(b_2) \geq \dotsb$.

On the other hand:
Let $\ell$ be a lower bound of $\varphi b_1 \geq \varphi b_2 \geq \dotsb$;
we need to prove that $\ell \leq \bw_n \Pi\varphi(a^n)$.
For all~$n$ and~$m$,
we have $a_m^n \geq b_{n\vee m}$
and so $\varphi(a_m^n) \geq \varphi(b_{n\vee m}) \geq \ell$.
Hence $\Pi\varphi(a^n) = \bw_m \varphi(a_m^n) \geq \ell$
for all~$n$.
So $\bw_n \Pi\varphi(a^n) \geq \ell$.

So $\bw_n \Pi\varphi(a^n)$ is the infimum of 
$\varphi b_1 \geq \varphi b_2 \geq \dotsb$,
and we are done.
\end{proof}
%
%                  MINIMALITY OF PI L
%
\begin{lem}
\label{L:Pi-minimal}
Let $C$ be a sublattice of~$V$.
Let $\psi\colon C\ra E$ be a valuation.
Assume
\begin{equation*}
\psi\text{ extends }\varphi
\qquad\text{and}\qquad
\vs{V}{C}\psi{E}\ \text{ is $\Pi$-complete.}
\end{equation*}
Then $\varphi$ is $\Pi$-extendible and
$\psi$ extends $\Pi\varphi$.
\end{lem}
\begin{proof}
Let $\varphi$-convergent $a_1 \geq a_2 \geq \dotsb$
be given.
We prove that
\begin{equation}
\label{eq:L:PiL-minimal}
\bw_n a_n \in C\qquad\text{and}\qquad \psi (\bw_n a_n) = \bw_n \varphi(a_n).
\end{equation}
We know that $\bw_n \varphi (a_n)$ exists
(since $a_1 \geq a_2 \geq \dotsb$ is $\varphi$-convergent),
and that $\varphi(a_n)= \psi(a_n)$
(since $\psi$ extends $\varphi$).
So $\bw_n \psi(a_n)$ exists too.
Hence $a_1 \geq a_2 \geq \dotsb$ is $\psi$-convergent.
Because $\vs{V}{C}\psi{E}$ is $\Pi$-complete
this implies that $\bw_n a_n \in C$ and 
$\psi (\bw_n a_n) = \bw_n \psi(a_n)$
(see Definition~\ref{D:system-complete2}\ref{D:system-complete2-pi}).
Hence we have proven Equation~\eqref{eq:L:PiL-minimal}.

Equation~\eqref{eq:L:PiL-minimal} implies that $\Pi L \subseteq C$.
So in order to prove that $\varphi$
is $\Pi$-extendible,
let us consider the valuation $\mu \eqdf \psi \,|\,\Pi L$.
We must prove that
$\mu(\bw_n a_n) = \bw_n \varphi(a_n)$
for every $\varphi$-convergent $a_1 \geq a_2 \geq \dotsb$
(see Definition~\ref{D:Pi-extendible}),
but this follows immediately from Equation~\eqref{eq:L:PiL-minimal}.
Hence $\varphi$ is $\Pi$-extendible.

Finally,
since we know that~$\varphi$ is $\Pi$-extendible,
we can talk about~$\Pi\varphi$,
and write the second part of
Equation~\eqref{eq:L:PiL-minimal}
as $\psi(\bw_n a_n) = \Pi \varphi(\bw_n a_n)$.
Hence $\psi$ extends $\Pi\varphi$.
\end{proof}
%
%                  MONOTONICITY OF PI
%
\begin{lem}
\label{L:Pi-monotonous}
Let $\vs{V}{K}\psi{E}$ be
a valuation system
such that $\psi$ extends $\varphi$.\\
Suppose that $\psi$ is $\Pi$-extendible.
Then $\varphi$ is $\Pi$-extendible
and  $\Pi\psi$ extends $\Pi\varphi$.
\end{lem}
\begin{proof}
Note that $\Pi\psi$ extends $\varphi$,
and $\vs{V}{\Pi K}{\Pi \psi}{E}$
is $\Pi$-complete (see Lemma~\ref{L:Pi-complete}).
So Lemma~\ref{L:Pi-minimal}
implies that
$\varphi$ is $\Pi$-extendible
and $\Pi\psi$ extends $\Pi\varphi$.
\end{proof}
%
%                  PI EXTENDIBLE IFF PI-CONTINUOUS
%
\begin{lem}
\label{L:Pi-continuity}
$\varphi$ is $\Pi$-extendible
if and only if  $\varphi$
has the following property.
\begin{equation}
\label{eq:Pi-cont}
\left[ \quad
\begin{minipage}{.7\columnwidth}
For every $b\in L$ and $\varphi$-convergent 
 $a_1 \geq a_2 \geq \dotsb$,
we have
\begin{equation*}
\bw_n a_n \leq b
\quad\implies\quad
\bw_n \varphi (a_n) \leq \varphi(b).
\end{equation*}
\end{minipage}
\right.
\end{equation}
\end{lem}
\begin{proof}
$\implies$\ 
Suppose $\varphi$ is $\Pi$-extendible.
Then $\varphi$ has Property~\eqref{eq:Pi-cont},
because if $b\in L$ and $\varphi$-convergent $a_1 \geq a_2 \geq \dotsb$
with $\bw_n a_n \leq b$ are given,
then we have
\begin{equation*}
\bw_n \varphi(a_n) 
\ =\ 
\Pi\varphi(\bw_n a_n)
\ \leq\ 
\varphi(b).
\end{equation*}

\noindent$\Longleftarrow$\ 
Suppose~$\varphi$ has Property~\eqref{eq:Pi-cont};
we prove $\varphi$ is $\Pi$-extendible.
We claim that
\begin{equation}
\label{eq:Piphi-order-preserving}
\bw_n a_n \ \leq\ \bw_n b_n 
\quad\implies\quad
\bw_n \varphi(a_n) \ \leq\ \bw_n \varphi(b_n),
\end{equation}
where $a_1 \geq a_2 \geq \dotsb$ and $b_1 \geq b_2 \geq \dotsb$
are $\varphi$-convergent.

Indeed,
if $\bw_n a_n \leq \bw_n b_n$,
then $\bw_n a_n \leq b_m$ for all~$m$,
so $\bw_n \varphi(a_n) \leq \varphi(b_m)$ for all~$m$
(by Property~\eqref{eq:Pi-cont}),
and hence $\bw_n \varphi(a_n) \leq \bw_n \varphi(b_m)$.

Equation~\eqref{eq:Piphi-order-preserving} implies that
 $\bw_n a_n = \bw_n b_n \implies \bw_n\varphi(a_n) = \bw_n \varphi(b_n)$,
so there is a unique map $\psi\colon \Pi L \ra E$ such that
\begin{equation*}
\psi(\bw_n a_n) \ =\ \bw_n \varphi(a_n)
\qquad
\text{ for all $\varphi$-convergent }a_1 \geq a_2 \geq \dotsb.
\end{equation*}
In fact, 
Equation~\eqref{eq:Piphi-order-preserving}
also implies that~$\psi$ is order preserving.

To prove that~$\varphi$ is $\Pi$-extendible
(see Definition~\ref{D:Pi-extendible}),
it suffices to show that~$\psi$ is a valuation.
For this,
it remains to be shown that~$\psi$ is modular
(see Definition~\ref{D:val}).

Let $a,b\in \Pi L$ be given; 
we need to prove $\psi(a\wedge b) + \psi(a\vee b)= \psi(a)+\psi(b)$.
Write  $a = \bw_n a_n$ and $b = \bw_n b_n$
where 
 $a_1 \geq a_2 \geq \dotsb$ and $b_1 \geq b_2 \geq \dotsb$
are $\varphi$-convergent sequences.
We have
\begin{alignat*}{3}
\varphi(a\wedge b) + \varphi(a \vee b) 
\ &=\ \psi(\bw_n a_n \wedge \bw_n b_n) + \psi(\bw_n a_n \vee \bw_n b_n) \\
\ &=\ \psi(\bw_n a_n \wedge b_n) + \psi(\bw a_n \vee b_n) \\
  &=\ \bw_n \varphi(a_n \wedge b_n) + \bw_n \varphi(\bw a_n \vee b_n) \\
  &=\ \bw_n \varphi(a_n \wedge b_n) + \varphi(a_n \vee b_n) \\
  &=\ \bw_n \varphi(a_n) + \varphi(b_n) \\
  &=\ \psi(\bw_n a_n) + \psi(\bw_n b_n).
\end{alignat*}
Hence $\psi$ is modular, which completes the proof
that $\varphi$ is $\Pi$-extendible.
\end{proof}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%j%
%
%                  THE HIERARCHY OF EXTENSIONS
%
\subsection{The Minimal Complete Extension} $\,$\\
\label{SS:hierarchy}
Let $\vs{V}{L}\varphi{E}$ 
and $\vs{V}{C}\psi{E}$
be valuation systems
such that 
\begin{equation*}
\text{$\psi$ extends~$\varphi$}
\qquad\text{and}\qquad
\text{$\psi$ is complete.}
\end{equation*}
Then $\vs{V}{C}{\psi}{E}$ is also $\Pi$-complete
(see Definition~\ref{D:system-complete2}).
Hence Lemma~\ref{L:Pi-minimal}
implies that $\varphi$ is $\Pi$-extendible 
and that $\psi$ extends  $\Pi \varphi$.

Thus, loosely speaking, $\Pi\varphi$ is the minimal extension of~$\varphi$
which is $\Pi$-complete.
In this subsection,
we identify the minimal extension~$\overline \varphi$ of~$\varphi$
which is complete.
We tackle this problem
in order to familiarize
the reader with the notions needed to define
``$\vs{V}{L}{\varphi}{E}$ is extendible''
(see Definition~\ref{D:extendible}).
These notions, which we introduce
 rather informally in this subsection,
will be defined rigorously and in a more general setting later on.

Let us begin. Note that
$\vs{V}{C}{\psi}{E}$ is also $\Sigma$-complete.
Hence $\varphi$ is $\Sigma$-extendible,
and $\psi$ extends $\Sigma\varphi$.
So we have the following situation.
\begin{equation*}
\text{$\psi$ extends~$\Pi \varphi$ and~$\Sigma\varphi$}
\qquad\text{and}\qquad
\text{$\vs{V}{C}{\psi}{E}$ is complete.}
\end{equation*}

By similar reasoning,
we see that $\Pi\varphi$ is $\Sigma$-extendible,
and that $\Sigma\varphi$ is $\Pi$-extendible
and that~$\psi$ extends both $\Sigma\Pi\varphi$ and $\Pi\Sigma\varphi$.
(Note that $\Pi(\Pi\varphi) =  \Pi\varphi$.)
So we have the following situation.
\begin{equation*}
\text{$\psi$ extends~$\Sigma\Pi \varphi$ and~$\Pi\Sigma\varphi$}
\qquad\text{and}\qquad
\text{$\vs{V}{C}{\psi}{E}$ is complete.}
\end{equation*}

Of course,
we can continue this proces.
More formally,
the clauses
\begin{alignat*}{5}
\Pi_{n+1} \varphi \, &=\, \Pi(\Sigma_n\varphi) &\qquad\quad
\Sigma_{n+1} \varphi \,&=\, \Sigma(\Pi_n\varphi) &\qquad\quad 
\Pi_0 \varphi \,&=\, \varphi \,&&=\, \Sigma_0 \varphi \\
\Pi_{n+1} L \, &=\, \Pi(\Sigma_n L) &\qquad\quad
\Sigma_{n+1} L \, &=\, \Sigma(\Pi_n L) &\qquad
\Pi_0 L \,&=\, L \,&&=\, \Sigma_0 L,
\end{alignat*}
give us for every~$n$ valuations
$\vs{V}{\Pi_n L}{\Pi_n\varphi}{E}$ and
$\vs{V}{\Sigma_n L}{\Sigma_n\varphi}{E}$.

Note that $\Pi\varphi$ extends $\varphi$.
Hence $\Sigma_2\varphi$ extends $\Sigma \varphi$ by
Lemma~\ref{L:Pi-monotonous}.
Hence $\Pi_3\varphi$ extends $\Pi_2 \varphi$.
Etcetera.
Similarly,
$\Sigma\varphi$ extends~$\varphi$,
so $\Pi_2 \varphi$ extends~$\Pi\varphi$,
and so on.

The hierarchy which we have obtained 
is shown in the following diagram.
\begin{equation*}
\xymatrix @=10pt {
& \Sigma\varphi \ar @{-} [rr] \ar @{-} [rrdd]
&& \Sigma_2\varphi  \ar @{-} [rr] \ar @{-} [rrdd]
&& \Sigma_3\varphi  \ar @{-} [rr] \ar @{-} [rrdd]
&& \Sigma_4\varphi  \ar @{-} [rr] \ar @{-} [rrdd]
&& \Sigma_5 \varphi  \ar @{-} [r]\ar @{-} [rd]
&& \dotsb
\\  
\varphi \ar @{-} [ru] \ar @{-} [rd] 
&&&&&&&&&&&\dotsb\\
& \Pi\varphi \ar @{-} [rr] \ar @{-} [rruu]
&& \Pi_2\varphi \ar @{-} [rr]\ar @{-} [rruu]
&& \Pi_3\varphi \ar @{-} [rr]\ar @{-} [rruu]
&& \Pi_4\varphi \ar @{-} [rr]\ar @{-} [rruu]
&& \Pi_5 \varphi \ar @{-} [r] \ar @{-} [ru]
&& \dotsb
}
\end{equation*}
We say that the \emph{hierarchy collapsed at~$Q$}, where $Q  \in 
\{\,  L,\, \Pi_1 L, \, \Sigma_1 L,\,\Pi_2 L,\,\dotsc\,\}$, if
\begin{equation*}
\Pi( Q ) \,=\, Q \,=\, \Sigma(Q).
\end{equation*}
In that case, let~$q\colon Q\ra E$ be the associated valuation
(either $\Pi_n\varphi$ or $\Sigma_n\varphi$ for some~$n$).
Then $\vs{V}{Q}{q}{E}$ is complete,
since it is both $\Pi$-complete 
and $\Sigma$-complete.

Note that the definition of $\Pi_n \varphi$
and $\Sigma_n \varphi$ does not depend
on which complete extension~$\psi$ of~$\varphi$ is given,
only on the fact that such~$\psi$ exists.
In particular,
if $\vs{V}{C'}{\psi'}{E}$ is any complete valuation system
such that~$\psi'$ extends~$\varphi$,
then $\psi'$ extends~$\Pi_n L$ and $\Sigma_n L$.
In particular,
such~$\psi'$ extends~$q$.
Hence~$q$ is the smallest complete extension of~$\varphi$ we sought.

However,
in general the hierarchy need not have collapsed at
any $\Pi_n L$ or~$\Sigma_n L$,
as we will show lateron \todo{add reference}.
So to find our minimal complete extension of~$\varphi$,
we need to push forwards.
To this end, consider the sets $\Pi_\omega L$
and $\Sigma_\omega L$ given by
\begin{equation*}
\Pi_\omega L \,\eqdf\,\textstyle{\bigcup_n}\, \Pi_n L
\qquad\text{and}\qquad
\Sigma_\omega L \,\eqdf\, \textstyle{\bigcup_n}\, \Sigma_n L.
\end{equation*}
Since $\Pi_{n} L \subseteq \Sigma_{n+1} L$
and $\Sigma_{n} L \subseteq \Pi_{n+1}L$ for all~$n$,
we see that $\Pi_\omega L = \Sigma_\omega L$.

Now,
since $\Pi_n\varphi$ extends $\Pi_m\varphi$
for $n\geq m$,
there is a unique map~$\Pi_\omega \varphi\colon \Pi_\omega L \ra E$
which extends all~$\Pi_n \varphi$.
One can easily see that $\vs{V}{\Pi_\omega L}{\Pi_\omega\varphi}{E}$
is a valuation system.
Similarly, there is a unique map 
$\Sigma_\omega \varphi \colon \Sigma_\omega L \ra E$
which extends all~$\Sigma_n\varphi$.
Then $\vs{V}{\Sigma_\omega L}{\Sigma_\omega\varphi}{E}$
is a valuation system.

Since $\Pi_{n+1}\varphi$ extends $\Sigma_{n}\varphi$
for all~$n$, one sees that $\Pi_\omega\varphi = \Sigma_\omega\varphi$.

Again, the hierarchy might have collapsed at~$\Pi_\omega L$,
i.e.,
\begin{equation*}
\Pi(\Pi_\omega L ) \,=\, \Pi_\omega L \,=\, \Sigma(\Pi_\omega L).
\end{equation*}
In that case $\Pi_\omega\varphi$ the minimal completion of~$\varphi$
we sought.

However,
again the hierarchy
might not have collapsed at~$\Pi_\omega L$,
so we consider the valuations
 $\Pi_{\omega+n}\varphi\eqdf\Pi_{n} (\Pi_\omega \varphi)$
and $\Sigma_{\omega+n}\varphi \eqdf \Sigma_{n}(\Pi_\omega\varphi)$.

\begin{equation*}
\xymatrix @=10pt {
& \Sigma\varphi \ar @{-} [rr] \ar @{-} [rrdd]
&& \Sigma_2 \varphi  \ar @{-} [r]\ar @{-} [rd]
&& \dotsb
& \Sigma_{\omega} \varphi \ar @{-} [rr] \ar @{-} [rrdd]
                          \ar @{=} [dd]
&& \Sigma_{\omega+1}\varphi \ar @{-} [rr] \ar @{-} [rrdd]
&& \Sigma_{\omega+2} \varphi  \ar @{-} [r]\ar @{-} [rd]
&& \dotsb
\\  
\varphi \ar @{-} [ru] \ar @{-} [rd] 
&&&&&\dotsb
&&
&&&&&\dotsb\\
& \Pi\varphi \ar @{-} [rr] \ar @{-} [rruu]
&& \Pi_{2} \varphi \ar @{-} [r] \ar @{-} [ru]
&& \dotsb
& \Pi_{\omega}\varphi \ar @{-} [rr] \ar @{-} [rruu]
&& \Pi_{\omega+1}\varphi \ar @{-} [rr] \ar @{-} [rruu]
&& \Pi_{\omega+2} \varphi \ar @{-} [r] \ar @{-} [ru]
&& \dotsb
}
\end{equation*}
With induction on ordinal numbers,
we can repeat this process endlessly.
However, 
the collapse of the hierarchy
can not be postponed indefinitely.

More formally,
let $\overline{L}\eqdf \{\, c\in C\colon
\exists\alpha[c\in \Pi_\alpha L]\, \}$.
Then we have $\Pi_\alpha L \subseteq \overline L$
for every~$\alpha$.
We want to prove that~$\Pi_\alpha L=\overline L$
for some~$\alpha$.
Define
\begin{equation*}
\alpha(c) \ =\ \min\, \{\  \beta\colon\   c\in \Pi_\beta L \ \}
\qquad\qquad(c\in \overline L).
\end{equation*}
Then the set of ordinal numbers $\{\,\alpha(c)\colon\,c\in \overline L\,\}$
has a supremum, say~$\xi$.
We have 
\begin{equation*}
c\in \Pi_{\alpha(c)} L \ \subseteq \ \Pi_\xi L
\qquad\qquad
(c\in \overline L).
\end{equation*}
So $\overline L \subseteq \Pi_{\xi} L$.
But we alread had $\Pi_{\xi}L\subseteq \overline L$.
Hence $\Pi_{\xi} L = \overline L$.

We claim that the hierarchy has collapsed at~$\xi$,
i.e., 
\begin{equation*}
\Pi(\Pi_\xi L) \,=\, \Pi_\xi L \,=\, \Sigma(\Pi_\xi L).
\end{equation*}
Indeed,
we have 
\begin{equation*}
\Pi_\xi L \ \subseteq\ \Sigma(\Pi_\xi L ) \ \subseteq \ \overline L
\ = \ \Pi_\xi L.
\end{equation*}
So $\Sigma(\Pi_\xi L) = \Pi_\xi L$.
Similarly, $\Pi_\xi L = \Pi(\Pi_\xi L)$.

One can easily verify that
$\overline\varphi \eqdf \Pi_\xi \varphi$
is the smallest complete  extension of~$\varphi$.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%                  THE NEED FOR THE HIERARCHY
%
\subsection{The Borel Hierarchy Theorem}$\,$\\
\label{SS:borel-hierarchy}
\subsubsection{Introduction}$\,$\\
Before we continue our study of the hierarchy introduced in the
Subsection~\ref{SS:hierarchy}, let us take a step back and wonder: 
 is this all --- the endless hierarchy --- neccesary?

Indeed,
using the terminology of Subsection~\ref{SS:hierarchy},
it is not unthinkable
that the hierarchy is always collapsed
at, say $\Sigma_{37} L$.
In that case 
the theory would be much simpler;
we would only need to
use the symbols up to~``$\Sigma_{37}$''.
In particular,
the involvement of the (infinite) ordinal numbers would not be required.

It turns out that we \emph{do} need a large amount of symbols
to desribe the hierarchy.
In this subsection we will give an example
where the hierarchy can only be collaped at~$\Pi_\alpha L$
or at~$\Sigma_\alpha L$
if the ordinal number~$\alpha$ is \emph{uncountable}.

On the bright side,
it does not get worse than this:
we will see 
(in Lemma~\ref{L:aleph1})
that 
the hierarchy is always collapsed at~$\Pi_{\aleph_1} L$,
where~$\aleph_1$
is the set of all countable ordinal numbers,
i.e., the smallest uncountable ordinal number.

The material in the subsection will not be used later on.
So the reader can safely skip this subsection
and proceed to Subsection~\ref{SS:hierarchy-abstract} 
on page~\pageref{SS:hierarchy-abstract}
if so desired.

\subsubsection{Borel Subsets}$\,$\\
Our example involves Borel sets.
Recall that the \emph{Borel subsets}
of a topological space~$X$
(such as~$\R$)
are those subsets one can form using countable intersection
and countable union starting from the open subsets.

Instead of~$\R$,
we  work with the Borel subsets
of the 
 \emph{Baire space},
 $\N^\N$.
In short,
the topology on~$\N^\N$
is the product topology
when~$\N$ is given the discrete toplogy.
To understand these words,
one might look at~\cite{Willard70},
but this is not necessary as we will
give a more direct description of~$\mathcal{T}$
in Subsubsection~\ref{SSS:bhier-open}.

While we could do the following
for~$\R$ as well,
it is much easier for~$\N^\N$.
\begin{nt}
Let $\mathcal{T}$
denote the set of open subsets of~$\N^\N$, \\
and let~$\mathcal{B}$
denote the set of Borel subsets of~$\N^\N$.
\end{nt}
Note that $\mathcal{B}$
is a sublattice of~$\wp(\N^\N)$,
and~$\mathcal{B}$ is a sublattice of~$\mathcal{T}$.

\begin{dfn}
Let $\psi\colon \mathcal{B}\ra \R$
be the map given by, for all~$A\in \mathcal{B}$,
\begin{equation*}
\psi(A)\ =\ 0.
\end{equation*}
\end{dfn}
Then $\psi$ is a valuation,
and we have the following valuation system.
\begin{equation*}
\vs{\wp(\N^\N)}{\mathcal{B}}{\psi}{\R}.
\end{equation*}

\begin{lem}
The valuation $\psi$ is complete with respect to~$\wp(\N^\N)$
(see Def.~\ref{D:system-complete}).
\end{lem}
\begin{proof}
Let $A_1 \supseteq A_2 \supseteq \dotsb$
and $B_1 \subseteq B_2 \subseteq \dotsb$
be a $\psi$-convergent sequences in~$\mathcal{B}$.
To prove that~$\psi$ is complete relative to~$\wp(\N^\N)$
we must show that 
\begin{equation*}
\textstyle 
\bigcap_n A_n \,\in\, \mathcal{B}
\qquad\text{and}\qquad
\bigcup_n B_n \,\in\, \mathcal{B}.
\end{equation*}
This follows immediately by definition of
the Borel subsets.
\end{proof}

\begin{rem}
\label{R:bhier-A}
Let $\mathcal{A}$ be a sublattice of~$\mathcal{B}$
and let
\begin{equation*}
\varphi\colon\mathcal{A}\longrightarrow \R
\end{equation*}
be the restriction of~$\psi$ to~$\mathcal{A}$.
Note that we are in the setting of Subsection~\ref{SS:hierarchy},
\begin{equation*}
\text{$\psi$ extends $\varphi$}
\qquad\text{and}\qquad
\text{$\psi$ is complete}.
\end{equation*}
Using the notation from Subsection~\ref{SS:hierarchy},
let us see what~$\Pi\mathcal{A}$
and $\Sigma\mathcal{A}$ are.

Note that every sequence $A_1 \supseteq A_2 \supseteq \dotsb$
in~$\mathcal{A}$ is $\varphi$-convergent,
as~$\psi$ is constant.
Further, given $A_1,A_2,\dotsc \in \mathcal{A}$,
we have $\bigcap_n A_n = \bigcap_n A_n'$,
where $A_1' \supseteq A_2' \supseteq \dotsb$
are defined by $A_n' = A_1 \cap \dotsb \cap A_n$.
So we see that
\begin{alignat}{3}
\label{eq:bhier-PiA}
\Pi\mathcal{A}\ &=\ 
\{ \ \textstyle{\bigcap_n} A_n\colon\ 
A_1,A_2,\dotsc \,\in\, \mathcal{A}\ \}.
\shortintertext{%
By a similar reasoning it is not hard to see that }
\label{eq:bhier-SigmaA}
\Sigma\mathcal{A}\ &=\ 
\{ \ \textstyle{\bigcup_n} A_n\colon\ 
A_1,A_2,\dotsc \,\in\, \mathcal{A}\ \}.
\end{alignat}
\end{rem}

\subsubsection{Statement of the Borel Hierarchy Theorem}$\,$\\
Let us spend words on where we are headed.
We will define a sublattice~$\mathcal{A}$
of~$\wp(\N^\N)$
in such a way that,
using the notation of Remark~\ref{R:bhier-A},
we have that
$\Sigma \mathcal{A}$
is  precisely the family of open subsets of~$\N^\N$,
while~$\Pi \mathcal{A}$ is the family
of closed subsets of~$\N^\N$.
From this information,
the reader can deduce with induction and the principle
of the excluded middle, that for every ordinal~$\alpha>0$,
and all $A\subseteq\N^\N$, 
\begin{equation}
\label{eq:bhier-complement}
A \,\in\, \Pi_\alpha \mathcal{A}
\qquad\iff\qquad
\N^\N\backslash A \,\in\, \Sigma_\alpha \mathcal{A}.
\end{equation}
The aim of this subsection
is to prove the following statement.
\begin{equation}
\label{eq:hierarchy-theorem}
\left[\quad
\begin{minipage}{.7\columnwidth}
\textbf{%
Let~$\alpha$ be a \emph{countable} ordinal number.\\
There is a set~$S\in \Sigma(\Pi_\alpha\mathcal{A})$
with $S\notin \Pi(\Sigma_\alpha\mathcal{A})$, and \\
there is a set $P\in\Pi(\Sigma_\alpha\mathcal{A})$
with $P\notin \Sigma(\Pi_\alpha\mathcal{A})$.}
\end{minipage}
\right.
\end{equation}
In particular
this means that if the hierarchy has collapsed
at~$\Pi_\alpha\mathcal{A}$
or at~$\Sigma_\alpha\mathcal{A}$
for some ordinal number~$\alpha$
then $\alpha$ must be \emph{uncountable}.

Statement~\eqref{eq:hierarchy-theorem}
is known in descriptive set theory
as the \emph{Borel Hierarchy Theorem}.
We will give a proof of Statement~\eqref{eq:hierarchy-theorem}
 in this subsection that
uses the principle of excluded middle
and is based on a beautiful paper by Veldman~\cite[paragraph 5]{Veldman08}.%
\footnote{
In this paper~\cite{Veldman08},
Veldman (also) gives
a proof of a variant of the Borel Hierarchy Theorem
using Brouwer's Continuity Principle
instead of the principle of excluded middle.}

\subsubsection{Open Subsets of $\N^\N$}$\,$\label{SSS:bhier-open}\\
Before we give a definition of~$\mathcal{A}$,
and start with the proof of Statement~\eqref{eq:hierarchy-theorem}
let us describe the topology~$\mathcal{T}$ 
on the Baire space~$\N^\N$ in more detail.
\begin{dfn}
Given $m,n\in\N$, define $B^m_n$ by
\begin{equation*}
B_n^m\ \eqdf\ \{ \ f\in\N^\N\colon\  f(n) \,=\, m \ \}.
\end{equation*}
\end{dfn}

\begin{rem}
\label{R:bhier-open}
For $A\subseteq \N^\N$,
we have $A\in \mathcal{T}$
if and only if for each $f\in A$
we have,
\begin{equation*}
f\,\in\, B^{m_1}_{n_1} \cap \dotsb \cap B^{m_K}_{n_K} \ \subseteq\ A,
\end{equation*}
for some $K\in\N$ and $m_1,\dotsc,m_K\in\N$
and $n_1,\dotsc,n_K\in\N$.
\end{rem}

We can formulate~Remark~\ref{R:bhier-open} more abstractly
with some notation.
\begin{dfn}
Let $\mathcal{S}$ and $\mathcal{S}_\cap$
be families of subsets of~$\N^\N$
given by
\begin{alignat*}{3}
\mathcal{S} \ &\eqdf\  
\{\ B^m_n \colon \ m,n\in\N \ \} \\
\mathcal{S}_\cap \ &\eqdf \ 
\{ \ S_1 \cap \dotsb\cap S_K\colon\ 
K\in\N,\  S_k\in \mathcal{S}\ \}.
\end{alignat*}
\end{dfn}

\begin{rem}\label{R:bhier-basis}
By Remark~\ref{R:bhier-open},
we see that~$\mathcal{S}$
is a subbasis for the topology~$\mathcal{T}$
on~$\N^\N$,
and we see that~$\mathcal{S}_\cap$ is a basis
for~$\mathcal{T}$.
In particular, we get
\begin{equation}
\label{eq:expression-t}
\mathcal{T}\ =\ 
\{\ \textstyle{\bigcup_n} A_n\colon \ A_1,A_2,\dotsc \in \mathcal{S}_\cap\ \}.
\end{equation}
\end{rem}
\begin{rem}\label{R:bhier-clopen}
Let $m,n\in\N$. Then  $B^m_n \in \mathcal{T}$ by Remark~\ref{R:bhier-basis}.
More suprisingly,
\begin{equation*}
\N^\N \backslash B^m_n\,\in\, \mathcal{T},
\end{equation*}
that is, $B^m_n$ is closed as well.
Indeed, this follows by the following equality.
\begin{equation*}
\N^\N \backslash B^m_n \ =\ 
\textstyle{\bigcup} \{\  B^k_n\colon\ k\in\N,\ k\neq m\ \}.
\end{equation*}

\end{rem}
\subsubsection{Definition of the Sublattice~$\mathcal{A}$ of~$\N^\N$}$\,$\\
Recall that we want to define a sublattice~$\mathcal{A}$
of~$\wp(\N^\N)$
so that $\Sigma \mathcal{A}$
are the open subsets of~$\N^\N$,
while~$\Pi \mathcal{A}$  are the closed subsets
(see Remark~\ref{R:bhier-A}).

Since the elements of~$\mathcal{S}$ are both open and closed,
we let~$\mathcal{A}$ be the sub-Boolean algebra
of~$\wp(\N^\N)$
generated by~$\mathcal{S}$.
More concretely:
\begin{dfn}\label{D:bhier-A}
Let $\mathcal{S}'$, $\mathcal{S}_\cap'$
and $\mathcal{A}$ be the families of subsets of~$\N^\N$ given by
\begin{alignat*}{3}
\mathcal{S}' 
\ &\eqdf\ 
\{\ B^m_n \colon \ m,n\in\N \ \}
\ \cup\ 
\{\ \N^\N\backslash B^m_n \colon \ m,n\in\N \ \}\\
\mathcal{S}_\cap'
\ &\eqdf\ 
\{ \ S_1 \cap \dotsb\cap S_K\colon\ 
K\in\N,\  S_k\in \mathcal{S}'\ \} \\ 
\mathcal{A}
\ &\eqdf\ 
\{ \ T_1 \cup \dotsb\cup T_L\colon\ 
L\in\N,\  T_\ell\in \mathcal{S}_\cap'\ \}.
\end{alignat*}
\end{dfn}
\begin{lem}
The family
$\mathcal{A}$
is a sublattice of~$\wp(\N^\N)$,
and 
\begin{equation*}
\N^\N\backslash A\,\in\,\mathcal{A}
\qquad\iff\qquad
A\,\in\,\mathcal{A}
\end{equation*}
for every~$A\subseteq\N^\N$,
and we have the following equalities.
\begin{alignat*}{3}
\Sigma\mathcal{A} \ &=\ 
\{ \ U\subseteq \N^\N\colon\ U\text{ is open}\ \},\\
\Pi\mathcal{A} \ &=\ 
\{ \ C\subseteq \N^\N\colon \ C\text{ is closed}\ \}.
\end{alignat*}
\end{lem}
\begin{proof}
We leave this to the reader.
\end{proof}


\subsubsection{Encoding the Elements of~$\mathcal{A}$}$\,$\\
Now that we have defined~$\mathcal{A}$,
we can start the proof of
 Statement~\eqref{eq:hierarchy-theorem}.
Maybe the most important idea behind the proof presented here
is that 
we can
encode the  Borel subsets~$\mathcal{B}$ of~$\N^\N$
as elements of~$\N^\N$.

To warm up
let us see how we can encode 
a tuple $a_1 \dotsb a_n$ of natural numbers as a natural number.
We leave it to the reader to 
find a bijection
\begin{equation*}
\left< -,-\right>\colon \ \N\times \N \longrightarrow \N \backslash\{1\}.
\end{equation*}
Let $\N^*$ be the set of tuples on~$\N$.
Given a tuple $a_1\dotsb a_n \in \N^*$, define
\begin{equation*}
\left< a_1a_2  \dotsb a_n \right>\ \eqdf\ 
\left<a_1, \left<a_2,\dotsc \,\left<a_n, 1 \right>\,\dotsb\,\right>\right>.
\end{equation*}
Then the resulting map  $\left<-\right>\colon \N^* \ra \N$
is a bijection.

Let us now encode the elements of~$\mathcal{A}$
 (see Def.~\ref{D:bhier-A}).
Given $k \in \N$, let
\begin{equation*}
 \decode{k}^{\mathcal{S}'}
\ \eqdf \ 
\begin{cases}
\ \N^\N \backslash B^m_n & \text{ if $k\equiv\left<1mn\right>$}, \\
\ B^m_n & \text{ if $k\equiv\left<2mn\right>$}, \\
\ \varnothing &\text{ otherwise.}
\end{cases}
\end{equation*}
Then $\decode-^{\mathcal{S}'}\colon\N\ra \mathcal{S}'$
is a surjection.
Given~$k\in \N$ with $k\equiv \left<a_1\dotsb a_K\right>$, let
\begin{equation*}
 \decode{k}^{\mathcal{S}_\cap'}
\ \eqdf\ 
\decode{a_1}^{\mathcal{S}'} \,\cap\,\dotsb\,\cap\,\decode{a_K}^{\mathcal{S}'}.
\end{equation*}
Then $\decode-^{\mathcal{S}_\cap'}\colon \N\ra {\mathcal{S}_\cap'}$
is a surjection.
Given~$k\in \N$ with $k\equiv \left<a_1\dotsb a_K\right>$, let
\begin{equation*}
 \decode{k}^{\mathcal{A}}
\ \eqdf\ 
\decode{a_1}^{\mathcal{S}_\cap'} \,\cup\,
   \dotsb\,\cup\,\decode{a_K}^{\mathcal{S}_\cap'}.
\end{equation*}
Then $\decode-^{\mathcal{A}}\colon \N\ra {\mathcal{A}}$
is a surjection.

Let $A\in \mathcal{A}$ be given.
If
$\decode{k}^\mathcal{A}=A$
for some 
 $k\in \N$ 
we say
that~$k$ is a \emph{code} for~$A$.
Note that~$A$ might have multiple codes.
This will not be a problem.

\subsubsection{Encoding Countable Ordinal Numbers}$\,$\\
Before we can make the step
to encode all Borel subsets of~$\N^\N$
we need an encoding for
the countable ordinal numbers.
We need some notation.
\begin{nt}
Let $f\in\N^\N$
and $n\in \N$ be given.
Define $f^{[n]}\in \N^\N$ by, for $m\in\N$,
\begin{equation*}
f^{[n]} (m) \ =\ f(\,\left<n,m\right>\,).
\end{equation*}
\end{nt}
Since $\left< -,-\right>$
is a bijection from~$\N\times\N$
to $\N\backslash\{1\}$,
an element $f\in\N^\N$
is completely determined
by $f^{[1]},f^{[2]},\dotsc$
\emph{and}~$f(1)$.
More precisely,
the assignment
\begin{equation*}
f\quad\mapsto\quad
 f(1),\ f^{[1]},\  f^{[2]},\ \dotsc.
\end{equation*}
gives a bijection from
$\N^\N$ to $\N \times (\N^\N)^\N$.

To encode the countable ordinal numbers,
we use special elements of~$\N^\N$.
\begin{dfn}
\label{D:bhier-stump}
Let~$\Stump$ be the subset  of~$\N^\N$
inductively given by:
\begin{enumerate}
\item
If $f\in\N^\N$ and $f(1)\neq 1$, then $f\in \Stump$.

\item
If $f\in\N^\N$
and $f(1)=1$
and $f^{[n]}\in\Stump$ for all~$n\in\N$,
then $f\in \Stump$.
\end{enumerate}
The elements of~$\Stump$
are called \keyword{stumps}
and are used in Intuitionistic Mathematics
as a replacement for the countable ordinal numbers.
\end{dfn}
\begin{dfn}
\label{D:bhier-alpha}
Let $\alpha[-]\colon \Stump \ra \aleph_1$
be the map recursively defined by
\begin{equation*}
\alpha[f]\ =\ 
\begin{cases}
\ 0&
\text{if $f(1)\neq1$;}\\
\ \bv_{n\in \N}\ \alpha[f^{[n]}]+1&
\text{if $f(1)= 1$.}
\end{cases}
\end{equation*}
Recall that~$\aleph_1$
is the set of all countable ordinal numbers.
\end{dfn}

\begin{lem}
The map $\alpha[-]\colon \Stump\longrightarrow \aleph_1$ is surjective.
\end{lem}
\begin{proof}
To prove that $\alpha[-]$ is surjective,
we must 
show that
for each~$\alpha\in\aleph_1$
there is an~$f\in \Stump$ with $\alpha[f]=\alpha$.
We use induction on~$\alpha\in\aleph_1$.

First we must find an~$f\in\Stump$
with $\alpha[f]=0$.
Simply take the $f \in \N^\N$
with $f(n)=37$ for all~$n\in \N$.
Then $f(1)\neq 0$, so~$f\in\Stump$, and $\alpha[f]=0$.

Let $\alpha\in \aleph_1$
be given, 
and assume that $\alpha = \alpha[f]$
for some~$f\in\Stump$.
We need to find a~$g\in \Stump$
such that $\alpha[g]=\alpha+1$.
Define $g\in\N^\N$ by 
\begin{equation*}
g(1)\,=\,1,\qquad\text{and}\qquad
g^{[n]}\,=\, f\quad\text{for all~$n\in\N$}.
\end{equation*}
Then $g\in\Stump$, and $\alpha[g]=\bv_{n\in\N} \ \alpha[f]+1$.
Since $\alpha[f]=\alpha$, we have $\alpha[g]=\alpha+1$.

Let $\lambda\in\aleph_1$
be a limit ordinal,
and assume that $\alpha[-]$ is surjective on~$\lambda$.
We must find an~$f\in \Stump$ such that~$\alpha[f]=\lambda$.
Since $\lambda\in\aleph_1$
there are  $\alpha_1 ,\alpha_2 ,\dotsc \in \lambda$
such that $\lambda = \bv_{n\in\N}\ \alpha_n+1$.
Since $\alpha[-]$
is surjective on~$\lambda$,
we know that there are  $f_1,f_2,\dotsc \in \Stump$
with $\alpha[f_n] = \alpha_n$.
Define $f\in\N^\N$ by 
\begin{equation*}
f(1)\,=\,1\qquad\text{and}\qquad
f^{[n]} \,=\, f_n\quad\text{for all $n\in\N$}.
\end{equation*}
Then $f\in\Stump$,
and $\alpha[f] = \bv_{n\in\N}\  \alpha[f^{[n]}]+1 
= \bv_{n\in\N}\ \alpha_n +1 = \lambda$.
\end{proof}

\subsubsection{Encoding Borel Subsets of~$\N^\N$}$\,$\\
We are now ready to encode the Borel subsets of~$\N^\N$.
\begin{dfn}
With recursion on~$\Stump$
define for each~$f\in\Stump$
maps
\begin{equation*}
\decode{-}^\Pi_f \colon \N^\N \longrightarrow \Pi(\Sigma_{\alpha[f]}\mathcal{A})
\qquad
\text{and}
\qquad
\decode{-}^\Sigma_f \colon \N^\N 
\longrightarrow \Sigma(\Pi_{\alpha[f]}\mathcal{A})
\end{equation*}
such that the following clauses hold.
\begin{enumerate}
\item
For $f\in \Stump$ with $f(1)\neq 1$ we have
\begin{alignat*}{3}
\decode{g}^\Pi_f \ &=\  
\bigcap_{n\geq 2} \ \decode{\,g(n)\,}^\mathcal{A}&
\qquad\text{and}\qquad
\decode{g}^\Sigma_f \ &=\ 
\bigcup_{n\geq 2}\  \decode{\,g(n)\,}^\mathcal{A}.
\end{alignat*}

\item
For $f\in \Stump$ with $f(1)= 1$
we have
\begin{equation*}
\decode{g}^\Pi_f \ =\ 
\bigcap_{n\in\N}\ \decode{\,g^{[n]}\,}^\Sigma_{f^{[n]}}
\qquad\text{and}\qquad
\decode{g}^\Sigma_f \ =\ 
{\bigcup_{n\in\N}}\ \decode{\,g^{[n]}\,}^\Pi_{f^{[n]}}.
\end{equation*}
\end{enumerate}
\end{dfn}
\begin{lem}
For each~$f\in\Stump$
the maps
$\decode{-}^\Pi_f$
and 
$\decode{-}^\Sigma_f$
are surjective.
\end{lem}
\begin{proof}
We leave this to the reader.
\end{proof}
\begin{rem}
\label{R:bhier-g1}
Let $f\in \Stump$
and $g\in \N^\N$ be given.
Note that $\decode{g}^\Pi_f$
and $\decode{g}^\Sigma_f$
do not depend on~$g(1)$.
More precisely,
given $g'\in\N^\N$ 
with $g'(n)=g(n)$ for all~$n\geq 2$
--- so  possibly $g(1) \neq g'(1)$.
Then we have  $\decode{g}^\Pi_f = \decode{g'}^\Pi_f$
and $\decode{g}^\Sigma_f = \decode{g'}^\Sigma_f$.
\end{rem}
We use Remark~\ref{R:bhier-g1}
to combine the maps~$\decode{-}^\Pi_f$
and $\decode{-}^\Sigma_f$
into one map~$\decode{-}^\mathcal{B}_f$.
\begin{dfn}
Let
$\decode{-}^{\mathcal{B}}_{-}\colon 
  \N^\N\times \Stump \longrightarrow \mathcal{B}$
be given by, for $f\in\Stump$ and~$g\in\N^\N$,
\begin{equation}
\label{eq:decode-Borel}
\decode{g}^{\mathcal{B}}_f \ = \ 
\begin{cases}
\  \decode{g}^{\Pi}_f  \quad
& \text{if $g(1)= 37$},\\

\ \decode{g}^{\Sigma}_f
& \text{if $g(1)\neq 37$}.
\end{cases}
\end{equation}
\end{dfn}
We want to prove that $\decode{-}_{-}^\mathcal{B}$
is surjective. 
To do this,
we need a lemma.
\begin{lem}
\label{L:Borel-aleph1}
We have the following equality.
\begin{equation}
\label{eq:Borel-aleph1}
\mathcal{B} \ =\  \Pi_{\aleph_1} \mathcal{A}.
\end{equation}
\end{lem}
\begin{proof}
Note that we have already proven 
(at the end of
in Subsection~\ref{SS:hierarchy})
 that $\mathcal{B} = \Pi_\xi \mathcal{A}$
for \emph{some} ordinal number~$\xi$.
Statement~\eqref{eq:Borel-aleph1}
is an improvement.

Recall that $\Pi_{\aleph_1}\mathcal{A}
\equiv \bigcup_{\alpha\in\aleph_1} \Pi_\alpha \mathcal{A}$.
Since $\Pi \mathcal{A} \subseteq \Sigma_2 \mathcal{A}$,
we have
\begin{equation*}
\mathcal{T} \ \equiv\  \Sigma \mathcal{A} 
\ \subseteq\  \Pi_{\aleph_1}\mathcal{A}\ \subseteq\ \mathcal{B}.
\end{equation*}
Recall that~$\mathcal{B}$
is the family of all subsets of~$\N^\N$
that can be formed using countable unions and countable intersections
starting from~$\mathcal{T}$.
So to prove that $\Pi_{\aleph_1}\mathcal{A} = \mathcal{B}$
it suffices
to show that~$\Pi_{\aleph_1}\mathcal{A}$
is `closed' under countable unions and  intersections,
i.e.,
given $A_1,A_2,\dotsc \in \Pi_{\aleph_1}\mathcal{A}$
we must show that  $\bigcup_n A_n \in \Pi_{\aleph_1}\mathcal{A}$
and  $\bigcap_n A_n \in \Pi_{\aleph_1}\mathcal{A}$.
We will only prove that~$\bigcup_n A_n \in \Pi_{\aleph_1}\mathcal{A}$;
the proof of~$\bigcap_n A_n \in \Pi_{\aleph_1}\mathcal{A}$
is similar.

Let $A_n' \eqdf A_1 \cup \dotsb \cup A_n$
for each~$n\in\N$.
Then $\bigcup_n A_n' = \bigcup_n A_n$.
Since $\Pi_{\aleph_1}\mathcal{A}$
is a sublattice of~$\wp(\N^\N)$,
we get that 
$A_n'\in \Pi_{\aleph_1}\mathcal{A}
\equiv\bigcup_{\alpha\in\aleph_1}\Pi_{\alpha}\mathcal{A}$
for all~$n\in\N$.

Pick $\alpha_1,\alpha_2,\dotsc \in \aleph_1$
such that $A_n' \in \Pi_{\alpha_n} \mathcal{A}$
for all~$n\in\N$.
Let $\alpha\eqdf \bigvee_n \alpha_n$.
Then 
for all~$n\in\N$ we have $\alpha_n \leq \alpha$,
and  $\Pi_{\alpha_n} \mathcal{A} \subseteq \Pi_{\alpha}\mathcal{A}$,
and so $A_n'\in\Pi_{\alpha}\mathcal{A}$.
By definition of 
$ \Sigma(\Pi_{\alpha}\mathcal{A})$
we have $\bigcup_n A_n' \,\in\,
\Sigma(\Pi_{\alpha}\mathcal{A})\equiv\Sigma_{\alpha+1}\mathcal{A}$.
Now, note that since $\alpha_1,\alpha_2,\dotsc \in\aleph_1$,
also $\alpha=\bigvee_n \alpha_n \in \aleph_1$,
and so $\alpha+1 \in\aleph_1$.
Hence 
\begin{equation*}
\textstyle{\bigcup_n A_n \,=\,  \bigcup_n A_n' 
\ \in\  \Pi_{\alpha+1}\mathcal{A}
\,\subseteq\, \Pi_{\aleph_1}\mathcal{A}.}
\end{equation*}
So $\Pi_{\aleph_1} \mathcal{A}$
is closed under countable union.
Similarly, $\Pi_{\aleph_1} \mathcal{A}$
is closed under countable intersection.
It follows that  $\mathcal{B} = \Pi_{\aleph_1}\mathcal{A}$.
We have proven Statement~\eqref{eq:Borel-aleph1}.
\end{proof}

\begin{prop}
\label{P:bhier-surjective}
The map $\decode{-}^\mathcal{B}_{-}\colon \N^\N\times \Stump
\longrightarrow \mathcal{B}$
is surjective.
\end{prop}
\begin{proof}
Let $A\in\mathcal{B}$ be given.
We must find $f\in\Stump$
and $g\in\N^\N$ such that $A=\decode{g}^{\mathcal{B}}_f$.
By Lemma~\ref{L:Borel-aleph1}
we know that $A\,\in\,\Pi_{\aleph1} \mathcal{A}
\equiv \bigcup_{\alpha\in\aleph_1} \Pi_\alpha\mathcal{A}$.
Pick an~$\alpha\in \aleph_1$ with $A\in \Pi_\alpha \mathcal{A}$.
Since the map $\alpha[-]\colon \Stump \ra \aleph_1$
is a surjection,
there is an $f\in\Stump$
such that $\alpha[f] = \alpha$.
Since the map $\decode{-}^{\Pi}_{f}
\colon \N^\N \longrightarrow \Pi(\Sigma_{\alpha[f]} \mathcal{A})$
is a surjection
and $A\,\in\, \Pi_{\alpha} \mathcal{A}
\equiv \Pi_{\alpha[f]}\mathcal{A}
\subseteq \Pi(\Sigma_{\alpha[f]}\mathcal{A})$,
there is an~$h\in \N^\N$
such that $\decode{h}^{\Pi}_f = A$.

Now, let $g\in\N^\N$ be given by
$g^{[n]} = h^{[n]}$ for all~$n\in\N$
and $g(1)= 37$.
Then
\begin{equation*}
\decode{g}^{\mathcal{B}}_f
\ =\ 
\decode{g}^{\Pi}_f
\ =\ 
\decode{h}^{\Pi}_f
\ =\ 
A.
\end{equation*}
So we see that $\decode{-}^{\mathcal{B}}_{-}
\colon \N^\N\times\Stump\longrightarrow \mathcal{B}$
is surjective.
\end{proof}

\subsubsection{Outstanding Debt}$\,$\\
We take a small detour,
because in Example~\ref{E:not-convex}
we used a fact, Statement~\eqref{eq:not-convex-fact},
without proof,
and we are now in a position
to correct this situation.
\begin{nt}
Let $\mathbb{D}$ denote the \emph{Cantor set},
see~\cite[Examples~17.9c]{Willard70}.
\end{nt}
\begin{lem}
\label{L:bhier-disc-injection}
$\mathbb{D}$ is Borel negligble
subset of~$\R$,
and there is an injective map
\begin{equation*}
p\colon \N^\N \times \Stump \longrightarrow \mathbb{D}.
\end{equation*}
\end{lem}
\begin{proof}
We leave this to the reader.
\end{proof}

\begin{cor}
\label{C:bhier-embed}
We have the following situation.
\begin{equation*}
\xymatrix{
\mathcal{B} && 
\N^\N\times \Stump
  \ar[rr]^p 
  \ar[ll]_{\decode{-}^\mathcal{B}_{-}}
&& 
\mathbb{D}
}
\end{equation*}
The map~$\decode{-}^\mathcal{B}_{-}$
is surjective,
the map~$p$ is injective,
and~$\mathbb{D}$ is Borel negligible.
\end{cor}
\begin{proof}
Combine Lemma~\ref{L:bhier-disc-injection}
and Proposition~\ref{P:bhier-surjective}.
\end{proof}
\noindent
Note that Corollary~\ref{C:bhier-embed}
implies Statement~\eqref{eq:not-convex-fact}.
This ends our detour.

\subsubsection{Cataloguing Borel Subsets of~$\N^\N$}$\,$\\
We have encoded the Borel subsets
using elements of~$\N^\N$.
To prove the Borel Hierarchy Theorem
(see Statement~\eqref{eq:hierarchy-theorem})
we use the encoding to go one step further.
\begin{dfn}
For each~$f\in \Stump$,
define 
\begin{alignat*}{3}
U^\Pi_f
\ &=\ 
\{\ h\in\N^\N\colon\   h^{[1]} \in \decode{h^{[2]}}^\Pi_f\ \},&\qquad
U^\Sigma_f
\ &=\ 
\{\ h\in\N^\N\colon\ 
h^{[1]} \in \decode{h^{[2]}}^\Sigma_f \ \}.
\end{alignat*}
\end{dfn}
\noindent
One can think of the set~$\smash{U^\Pi_f}$
as a \emph{catalogue} of~$\smash{\Pi(\Sigma_{\alpha[f]}\mathcal{A}})$.\\
Indeed,
given $A\in \Pi(\Sigma_{\alpha[f]}\mathcal{A})$
with 
$A=\decode{g}^\Pi_f$
for some $g\in\N^\N$,
we have
\begin{equation*}
A \ =\ \{ \ h^{[1]}\colon \ h\in U^\Pi_f \ \text{ and } \ h^{[2]}=g\ \}.
\end{equation*}
The following lemma
might be the essential 
part of the Borel Hierarchy Theorem.
\begin{lem}
\label{L:bhier-lemma}
Let $f\in\Stump$ be given.
Then we have
\begin{equation}
\label{eq:peer-catalogue}
U_f^\Pi \,\in\, \Pi(\Sigma_{\alpha[f]}\mathcal{A})
\qquad\text{and}\qquad
U_f^\Sigma \,\in\, \Sigma(\Pi_{\alpha[f]}\mathcal{A}).
\end{equation}
\end{lem}
To give a proof of Lemma~\ref{L:bhier-lemma}
we need some notation and a lemma.
\begin{dfn}
\label{D:bhier-continuous}
Let $F\colon \N^\N\longrightarrow \N^\N$
be given.
\begin{enumerate}
\item
Given $A\subseteq \N^\N$,
let
$F^*(A)\,\eqdf\,\{\ g\in \N^\N\colon \ F(g)\in A\ \}$.
\item
We say that~$F$ is \keyword{continuous}
if~$F^*(B^m_n) \in \mathcal{T}$
for all~$m,n\in \N$.
\end{enumerate}
\end{dfn}
\begin{lem}
\label{L:bhier-continuous}
Let $F\colon \N^\N\ra \N^\N$ be a continuous function.\\
Let $\alpha$ be an ordinal number with~$\alpha>0$.
Then, for all~$A\subseteq\N^\N$,
\begin{alignat*}{3}
F^*(A) \,&\in\, \Pi_\alpha\mathcal{A}&
\qquad\text{when}\qquad A\,&\in\,\Pi_\alpha\mathcal{A}, \\
F^*(A) \,&\in\, \Sigma_\alpha\mathcal{A}&
\qquad\text{when}\qquad A\,&\in\,\Sigma_\alpha\mathcal{A}.
\end{alignat*}
\end{lem}
\begin{proof}
We leave this to the reader.
\end{proof}


\begin{proof}[Proof of Lemma~\ref{L:bhier-lemma}]
We prove Statement~\eqref{eq:peer-catalogue} using induction 
over~$f\in\Stump$.

\vspace{.6em}
\noindent
Let $f\in\Stump$ with $f(1)\neq1$ be given.
We must prove that
\begin{equation*}
U^\Pi_f \ \in\ \Pi \mathcal{A}
\qquad\text{and}\qquad
U^\Sigma_f \ \in\ \Sigma \mathcal{A}.
\end{equation*}
We will only prove that $U^\Sigma_f \in \Sigma \mathcal{A}$
since the proof of~$U^\Pi_f \in \Pi\mathcal{A}$ is similar.

Let $h\in \N^\N$ be given.
Note that the following are equivalent.
\begin{alignat*}{3}
h \ &\in\ U^\Sigma_f, \\
h^{[1]} \ &\in\ \decode{\,h^{[2]}\,}^\Sigma_f, \\
h^{[1]} \ &\in\ \textstyle{\bigcup_{n\geq 2}}\ 
           \decode{\,h^{[2]}(n)\,}^\mathcal{A}, \\
h^{[1]} \ &\in\ \decode{\,h^{[2]}(n)\,}^\mathcal{A}
\qquad &&\text{for some~$n\geq 2$},\\
h^{[1]} \ &\in\ \decode{k}^\mathcal{A}
\quad\text{and}\quad
h^{[2]}(n)\,=\,k
\qquad &&\text{for some~$n\geq 2$,\  $k\in\N$}.
\end{alignat*}
Let $F\colon \N^\N\ra\N^\N$
be given by $F(g)=g^{[1]}$.
Then $h^{[1]} \in \decode{k}^{\mathcal{A}}$
iff $h\in F^*(\decode{k}^\mathcal{A})$.
Further, note that $h^{[2]}(n)=k$ iff $h\in B^{k}_{\left<2,n\right>}$.
All in all, we get 
\begin{equation}
\label{eq:bhier-Usigmaf0}
U^\Sigma_f
\ =\ 
\bigcup_{k\in\N} \bigcup_{n\geq 2}
\ F^*(\,\decode{k}^\mathcal{A}\,) \ \cap\  B^k_{\left<2,n\right>}.
\end{equation}
By Equation~\eqref{eq:bhier-Usigmaf0}
to prove~$U^\Sigma_f \in\Sigma \mathcal{A}$
it suffices to show 
that~$F^*(\,\decode{k}^\mathcal{A}\,) \in \Sigma\mathcal{A}$.

Since it is not hard to see that~$F$ is continuous
(see Definition~\ref{D:bhier-continuous})
and $\decode{k}^\mathcal{A} \in \Sigma\mathcal{A}$
we get by Lemma~\ref{L:bhier-continuous} that
$F^*(\,\decode{k}^\mathcal{A}\,) \,\in\, \Sigma\mathcal{A}$.
Hence~$U^\Sigma_f \in \Sigma\mathcal{A}$.

\vspace{.6em}
\noindent
Recall that we are proving Statement~\eqref{eq:peer-catalogue}
using induction on~$f\in\Stump$.\\
Let $f\in\Stump$
with $f(1)=1$ be given.
Assume that for all~$n\in\N$,
\begin{equation*}
U^\Pi_{f^{[n]}} \,\in\, \Pi(\Sigma_{\alpha[f^{[n]}]}\mathcal{A})
\qquad\text{and}\qquad
U^\Sigma_{f^{[n]}} \,\in\, \Sigma(\Pi_{\alpha[f^{[n]}]}\mathcal{A}).
\end{equation*}
We must prove that 
$U^\Pi_f \,\in\, \Pi(\Sigma_{\alpha[f]} \mathcal{A})$
and $U^\Sigma_f \,\in\, \Sigma(\Pi_{\alpha[f]} \mathcal{A})$.
We will prove that
\begin{equation}
\label{eq:peer-catalogue-induction}
U^\Sigma_f \,\in\, \Sigma(\Pi_{\alpha[f]} \mathcal{A}),
\end{equation}
and we leave the proof of
$U^\Pi_f \,\in\, \Pi(\Sigma_{\alpha[f]} \mathcal{A})$
to the reader.

To proceed, we need some notation.
We will define a `pairing'
\begin{equation*}
P\colon \N^\N \times \N^\N\longrightarrow \N^\N.
\end{equation*}
Let $h_1,h_2\in\N^\N$ be given.
Define $P(h_1,h_2)\in\N^\N$
by: $P(h_1,h_2)(1) = 1$, and
\begin{equation*}
(\,P(h_1,h_2)\,)^{[1]} \,=\, h_1,
\qquad\text{and}\qquad
(\,P(h_1,h_2)\,)^{[2]} \,=\, h_2,
\end{equation*}
and $(\,P(h_1,h_2)\,)^{[n]}(m) = 1$
for all~$n,m\in\N$ with $n>2$.

Let $h\in\N^\N$ be given.
Note that the following are equivalent.
\begin{alignat*}{3}
h \ &\in\  U^\Sigma_f, \\
h^{[1]} \ &\in\ \decode{\,h^{[2]}\,}^\Sigma_f, \\
h^{[1]} \ &\in\ 
\textstyle{\bigcup_{n\in\N}}\ 
        \decode{\,h^{[2][n]}\,}^\Pi_{f^{n}}, \qquad\\
 h^{[1]} \ &\in\ 
        \decode{\,h^{[2][n]}\,}^\Pi_{f^{n}} 
\qquad&\text{for some $n\in\N$},\\
 P(h^{[1]},h^{[2][n]}) \ &\in\ 
        U^{\Pi}_{f^{[n]}} 
\qquad&\text{for some $n\in\N$}.
\end{alignat*}
Now,
for each~$n\in\N$,
let $F_n\colon \N^\N \ra\N^\N$
be given by, for $h\in\N^\N$,
\begin{equation*}
F_n(h) \ =\  P(h^{[1]},h^{[2][n]}).
\end{equation*}
Then using the notation of Definition~\ref{D:bhier-continuous}
we see that
\begin{equation}
\label{eq:expression-us}
U^\Sigma_f \ =\ 
\textstyle{\bigcup_{n\in\N}} \ F_n^*(\, U^\Pi_{f^{[n]}}\,).
\end{equation}
Recall that we must prove
that $U^\Sigma_f \in \Sigma(\Pi_{\alpha[f]}\mathcal{A})$.
It suffices to prove that
\begin{equation}
\label{eq:hierarchy-theorem-last-step}
F_n^*(\, U^\Pi_{f^{[n]}}\,) \ \in\  \Pi(\Sigma_{\alpha[f^{[n]}]}\mathcal{A}),
\end{equation}
by  Statement~\eqref{eq:expression-us},
and because we have
\begin{equation*}
\Pi(\Sigma_{\alpha[f^{[n]}]}\mathcal{A}) 
\ \subseteq \ \Sigma(\Pi_{\alpha[f]}\mathcal{A}).
\end{equation*}
By Lemma~\ref{L:bhier-continuous}
to prove
that Statement~\eqref{eq:hierarchy-theorem-last-step}
holds
it suffices to show that~$F_n$
is continuous; we leave this to the reader.
Hence we have proven Statement~\eqref{eq:peer-catalogue-induction}.

This completes the proof of Statement~\eqref{eq:peer-catalogue}.
\end{proof}

\subsubsection{Diagonalization}$\,$\\
With the catalogues $U^\Sigma_f$ and $U^\Pi_f$
at our disposal
we use Cantor's diagonal argument
to prove the Borel Hierarchy Theorem
(see Statement~\eqref{eq:hierarchy-theorem}).

\begin{dfn}
Let $f\in\Stump$ be given.
Let $D^\Pi_f$
and $D^\Sigma_f$ be give by
\begin{equation*}
D^\Pi_f \ \eqdf \ 
\{\  g\in\N^\N\colon\ g\notin \decode{g}^\Pi_f\ \},
\qquad
D^\Sigma_f \ \eqdf \ 
\{\  g\in\N^\N\colon\ g\notin \decode{g}^\Sigma_f\ \}.
\end{equation*}
\end{dfn}

\begin{thm}
\label{T:borel-hierarchy}
Let~$f\in \Stump$
be given.
Then we have
\begin{alignat*}{3}
D^\Pi_f \,&\in\, \Sigma(\Pi_{\alpha[f]} \mathcal{A})
\qquad&&\text{and}\qquad&
D^\Pi_f \,&\notin\, \Pi(\Sigma_{\alpha[f]} \mathcal{A}),\\
D^\Sigma_f \,&\in\, \Pi(\Sigma_{\alpha[f]} \mathcal{A})
\qquad&&\text{and}\qquad&
D^\Sigma_f \,&\notin\, \Sigma(\Pi_{\alpha[f]} \mathcal{A}).
\end{alignat*}
\end{thm}
\begin{proof}
Let $f\in\Stump$ be given.
We will only prove 
that
\begin{equation}
\label{eq:bhier}
D^\Pi_f \,\in\, \Sigma(\Pi_{\alpha[f]} \mathcal{A})
\qquad\text{and}\qquad
D^\Pi_f \,\notin\, \Pi(\Sigma_{\alpha[f]} \mathcal{A}),
\end{equation}
because there is a similar proof of
$D^\Sigma_f \,\in\, \Pi(\Sigma_{\alpha[f]} \mathcal{A})$
and
$D^\Sigma_f \,\notin\, \Sigma(\Pi_{\alpha[f]} \mathcal{A})$.

Let us first prove that 
$D^\Pi_f\notin  \Pi(\Sigma_{\alpha[f]}\mathcal{A})$.
So assume 
$D^\Pi_f\in \Pi(\Sigma_{\alpha[f]}\mathcal{A})$
in order to reach a contradiction.
Since the map $\decode{-}^\Pi_f\colon \N^\N \longrightarrow \Pi(\Sigma_{\alpha[f]}\mathcal{A})$
is surjective,
there is a~$g_\delta \in \N^\N$
with $\decode{g_\delta}^\Pi_f = D^\Pi_f$.
Then
by definition of~$D^\Pi_f$,
\begin{equation}
\label{eq:diagonalization}
g_\delta \,\in\, D^\Pi_f
\quad\iff\quad
g_\delta \,\notin\, \decode{g_\delta}^\Pi_f = D^\Pi_f.
\end{equation}
Statement~\eqref{eq:diagonalization}
leads to a contradiction.
Hence we conclude that $D^\Pi_f\notin \Pi(\Sigma_{\alpha[f]}\mathcal{A})$.

Let us prove that $D^\Pi_f \in \Sigma (\Pi_{\alpha[f]}\mathcal{A})$.
Let~$\Delta\colon \N^\N \ra \N^\N$
be given by, for $g\in \N^\N$,
\begin{equation*}
\Delta(g) \ =\  P(g,g).
\end{equation*}
Then $\Delta$ is continuous
(see Definition~\ref{D:bhier-continuous})
and we have, for $g\in \N^\N$,
\begin{equation*}
g\,\in\, D^\Pi_f 
\quad\iff\quad
\Delta(g) \,\notin\, U^\Pi_f
\quad\iff\quad
g\,\in\,\Delta^*(\,\N^\N \backslash U^\Pi_f \,).
\end{equation*}
So we see that $D^\Pi_f = \Delta^*(\,\N^\N\backslash U_f^\Pi\,)$.
Recall that we must prove that
\begin{equation*}
D^\Pi_f \ \in\  \Sigma (\Pi_{\alpha[f]}\mathcal{A}).
\end{equation*}
Since~$\Delta$
is continuous
it suffices to show that 
$\N^\N\backslash U_f^\Pi \ \in\  \Sigma (\Pi_{\alpha[f]}\mathcal{A})$
by Lemma~\ref{L:bhier-continuous}.
We already know that $U^\Pi_f \in \Pi(\Sigma_{\alpha[f]} \mathcal{A})$,
so 
by Statement~\eqref{eq:bhier-complement}
we get
\begin{equation*}
\N^\N\backslash U^\Pi_f \ \in\  \Sigma(\Pi_{\alpha[f]} \mathcal{A}).
\end{equation*}
So
$D^\Pi_f \in \Sigma (\Pi_{\alpha[f]}\mathcal{A})$.
Hence we have proven Statement~\eqref{eq:bhier}.
\end{proof}

\begin{cor}
Statement~\eqref{eq:hierarchy-theorem}
holds.
\end{cor}









































%
%                  HIERARCHY OF EXTENSIONS
%
\subsection{The Hierarchy of Extensions}
\label{SS:hierarchy-abstract}
\begin{prop}
\label{P:hier}
Let $\vs{V}{L}{\varphi}{E}$
be a valuation system.

Consider the following `package' of objects.
\begin{enumerate}
\item \label{hier:first-cond}
For each ordinal number~$\alpha$,
statements
\begin{equation*}
\text{``\keyword{$\varphi$ is $\Pi_\alpha$-extendible}''}
\qquad\text{ and }\qquad
\text{``\keyword{$\varphi$ is $\Sigma_\alpha$-extendible}''}.
\end{equation*}
\item
For each~$\alpha$ such that~$\varphi$ 
is $\Pi_\alpha$-extendible,
a valuation system 
\begin{equation*}
\vs{V}{\Pi_\alpha L}{\Pi_\alpha \varphi}{E},
\end{equation*}

\item
and for each~$\alpha$ such that~$\varphi$ is $\Sigma_\alpha$-extendible,
a valuation system 
\begin{equation*}
\vs{V}{\Sigma_\alpha L}{\Sigma_\alpha \varphi}{E}.
\end{equation*}
\end{enumerate}
There is precisely one such package which satisfies
the following conditions.
\begin{enumerate}
\item
$\varphi$ is {$\Pi_0$-extendible}
and $\varphi$ is {$\Sigma_0$-extendible}.
Moreover,
\begin{equation*}
\Pi_0 L = L\qquad 
\Sigma_0 L = L \qquad 
\Pi_0 \varphi = \varphi \qquad
\Sigma_0 \varphi = \varphi.
\end{equation*}

\item
Let $\alpha$ be an ordinal number.
We have 
$\varphi$ is {$\Pi_{\alpha+1}$-extendible}
if and only if $\varphi$ is $\Sigma_{\alpha}$-extendible
and $\Sigma_\alpha \varphi$ is $\Pi$-extendible.
If $\varphi$ is $\Pi_{\alpha+1}$-extendible,
then 
\begin{equation*}
\Pi_{\alpha+1} L = \Pi(\Sigma_\alpha L)\qquad\qquad
\Pi_{\alpha+1} \varphi = \Pi(\Sigma_\alpha \varphi),
\end{equation*}
where 
 $\vs{V}{\Pi(\Sigma_\alpha L)}{\Pi(\Sigma_\alpha\varphi)}{E}$
is the valuation system from Definition~\ref{D:Pi-extendible}.

\item
Let $\alpha$ be an ordinal number.
We have 
$\varphi$ is {$\Sigma_{\alpha+1}$-extendible}
if and only if $\varphi$ is $\Pi_{\alpha}$-extendible
and $\Pi_\alpha \varphi$ is $\Sigma$-extendible.
If $\varphi$ is $\Sigma_{\alpha+1}$-extendible,
then 
\begin{equation*}
\Sigma_{\alpha+1} L = \Sigma(\Pi_\alpha L)\qquad\qquad
\Sigma_{\alpha+1} \varphi = \Sigma(\Pi_\alpha \varphi).
\end{equation*}

\item
Let~$\lambda$ be a limit ordinal.
We have $\varphi$ is {$\Pi_\lambda$-extendible}
if and only if $\varphi$ is $\Pi_\alpha$-extendible
for each~$\alpha < \lambda$.
If $\varphi$ is $\Pi_\lambda$-extendible,
then 
\begin{equation*}
\Pi_\lambda L = \textstyle{\bigcup_{\alpha < \lambda} \Pi_\alpha L}
\qquad\qquad
\Pi_\lambda \varphi (c) = \Pi_\beta \varphi(c),
\end{equation*}
where $\beta < \lambda$ and $c\in \Pi_\beta L$.

\item
\label{hier:last-cond}
Let~$\lambda$ be a limit ordinal.
We have $\varphi$ is {$\Sigma_\lambda$-extendible}
if and only if $\varphi$ is $\Sigma_\alpha$-extendible
for each~$\alpha < \lambda$.
If $\varphi$ is $\Sigma_\lambda$-extendible,
then 
\begin{equation*}
\Sigma_\lambda L = \textstyle{\bigcup_{\alpha < \lambda} \Sigma_\alpha L}
\qquad\qquad
\Sigma_\lambda \varphi (c) = \Sigma_\beta \varphi(c),
\end{equation*}
where $\beta < \lambda$ and $c\in \Sigma_\beta L$.
\end{enumerate}
Moreover,
the hierarchy we obtain has the following form.
\begin{equation*}
\xymatrix @=10pt {
& \Sigma\varphi \ar @{-} [rr] \ar @{-} [rrdd]
&& \Sigma_2 \varphi  \ar @{-} [r]\ar @{-} [rd]
&& \dotsb
& \Sigma_{\omega} \varphi \ar @{-} [r] \ar @{-} [rd]
                          \ar @{=} [dd]
&& \dotsb\ \dotsb
& \Sigma_{\lambda} \varphi \ar @{-} [r] \ar @{-} [rd]
                          \ar @{=} [dd]
&& \dotsb
\\  
\varphi \ar @{-} [ru] \ar @{-} [rd] 
&&&&&\dotsb
&&&\dotsb\ \dotsb
&&&\dotsb\\
& \Pi\varphi \ar @{-} [rr] \ar @{-} [rruu]
&& \Pi_{2} \varphi \ar @{-} [r] \ar @{-} [ru]
&& \dotsb
& \Pi_{\omega}\varphi \ar @{-} [r] \ar @{-} [ru]
&& \dotsb\ \dotsb
& \Pi_{\lambda}\varphi \ar @{-} [r] \ar @{-} [ru]
&& \dotsb
}
\end{equation*}
More precisely, the following statements are true.
\begin{enumerate}[label=(\Roman*)]
\item\label{req:hierarchy-A}
For all $\beta < \gamma$,
if  $\varphi$ is $\Pi_\gamma$-extendible,
then $\varphi$ is both $\Pi_\beta$- and $\Sigma_\beta$-extendible,
and the map $\Pi_\gamma\varphi$ extends 
both $\Pi_\beta\varphi$ and $\Sigma_\beta\varphi$.
\item\label{req:hierarchy-B}
For all $\beta < \gamma$,
if  $\varphi$ is $\Sigma_\gamma$-extendible,
then $\varphi$ is both $\Pi_\beta$- and $\Sigma_\beta$-extendible,
and the map $\Sigma_\gamma\varphi$ extends 
both $\Pi_\beta\varphi$ and $\Sigma_\beta\varphi$.
\item\label{req:hierarchy-C}
Let~$\lambda$
be zero or a a limit ordinal.
Then $\varphi$ is $\Pi_\lambda$-extendible
if and only if $\varphi$ is $\Sigma_\lambda$-extendible.
Furthermore,
if $\varphi$ is $\Pi_\lambda$-extendible,
then $\Pi_\gamma\varphi = \Sigma_\gamma\varphi$.
\end{enumerate}
\end{prop}
\begin{proof}
Straightforward with transfinite induction.
We leave this to the reader.
\end{proof}
%
%                  COLLAPSE
%
\begin{dfn}
\label{D:collapse}
Let $\vs{V}{L}\varphi{E}$ be a valuation system.
\begin{enumerate}
\item
\label{D:collapse-1}
We say the \keyword{hierarchy has collapsed at~$Q$} if
\begin{equation*}
\left[\quad
\begin{minipage}{.7\columnwidth}
$\varphi$ is $\Pi_{\alpha+1}$-extendible
 and $\Sigma_{\alpha+1}$-extendible,
and 
\begin{equation*}
\Pi(Q)\ =\ Q\ =\ \Sigma(Q),
\end{equation*}
where 
$Q=\Pi_\alpha \varphi$
or $Q=\Sigma_\alpha \varphi$.
\end{minipage}
\right.
\end{equation*}

\item
\label{D:collapse-2}
We say the \keyword{hierarchy collapses}
if the hierarchy has collapsed at some~$Q$.
\end{enumerate}
\end{dfn}

\begin{lem}
\label{L:collapse}
Let $\vs{V}{L}\varphi{E}$
be a valuation system.
\begin{enumerate}
\item
\label{L:collapse-1}
Let $\alpha<\beta$ be ordinal numbers.
Suppose the hierarchy has collapsed at~$\Pi_\alpha \varphi$.
Then $\varphi$ is $\Pi_\beta$-extendible and $\Sigma_\beta$-extendible,
and  
\begin{equation*}
\Pi_\beta \varphi \,=\, \Pi_\alpha \varphi \,=\, \Sigma_\beta \varphi.
\end{equation*}
\item
\label{L:collapse-2}
If the hierarchy collapses
at $\Pi_\alpha \varphi$ and at $\Sigma_\alpha\varphi$
for some~$\alpha$,
then $\Pi_\alpha\varphi = \Sigma_\alpha\varphi$.

\item
\label{L:collapse-3}
Suppose the hierarchy has collapsed at~$Q_1$ and at~$Q_2$.
Then $Q_1 = Q_2$.
\item
\label{L:collapse-4}
The hierarchy collapses
if and only if 
it has collapsed at $\Pi_\alpha \varphi$
for some~$\alpha$.
\end{enumerate}
\end{lem}
\begin{proof}
\noindent\ref{L:collapse-1}\ 
Suppose that the hierarchy has collapsed at~$\Pi_\alpha\varphi$
for some~$\alpha$.
We prove with induction on~$\beta$,
that for all~$\beta\geq \alpha$,
we have that
$\varphi$
is $\Pi_\beta$-extendible and $\Sigma_\beta$-extendible,
and that  $\Pi_\beta\varphi = \Pi_\alpha\varphi = \Sigma_\beta \varphi$.

If $\beta=\alpha$, the proposition is trivial.

Suppose $\varphi$ is $\Pi_\beta$-extendible
and $\Sigma_\beta$-extendible,
and that $\Pi_\beta\varphi = \Pi_\alpha\varphi = \Sigma_\beta \varphi$,
for some~$\beta$.
Since the hierarchy has collapsed at $\Pi_\alpha\varphi$,
we know that~$\varphi$ is $\Sigma_{\alpha+1}$-extendible.
(Hence $\Pi_\alpha \varphi = \Pi_\beta\varphi$
is $\Sigma$-extendible,
and so $\varphi$ is $\Sigma_{\beta+1}$-extendible.)
Moreover, we know that $\Sigma(\Pi_\alpha\varphi)=\Pi_\alpha\varphi$.
Hence  
\begin{equation*}
\Sigma_{\beta+1} \varphi \,=\, \Sigma(\Pi_\beta \varphi)
\,=\, \Sigma(\Pi_\alpha\varphi) \,=\, \Pi_\alpha\varphi.
\end{equation*}
Similarly, $\Pi_{\beta+1} \varphi = \Pi_\alpha\varphi$.

Suppose $\lambda > \alpha$ is a limit ordinal
and assume that
for all~$\beta<\lambda$ with $\alpha\leq\beta$,
$\varphi$ is $\Pi_\beta$-extendible
and $\Sigma_\beta$-extendible,
and $\Pi_\beta\varphi = \Pi_\alpha\varphi = \Sigma_\beta\varphi$.
We need to prove that $\varphi$ is $\Pi_\lambda$-extendible
and that $\Pi_\lambda \varphi = \Pi_\alpha \varphi$
(since $\Pi_\lambda \varphi = \Sigma_\lambda\varphi$).

Let $\beta<\lambda$ be given.
In order to prove that $\varphi$ is $\Pi_\lambda$-extendible,
we must show that~$\varphi$ is $\Pi_\beta$-extendible 
(see Proposition~\ref{P:hier}).
If $\beta\geq \alpha$,
then we know that $\varphi$ is $\Pi_\beta$-extendible by assumption.
If $\beta< \alpha$,
then we know that $\varphi$ is $\Pi_\beta$-extendible,
because $\varphi$ is $\Pi_\alpha$-extendible and $\alpha> \beta$
(see Proposition~\ref{P:hier}).

It remains to be show that $\Pi_\lambda \varphi = \Pi_\alpha \varphi$.
To this end, note that
\begin{equation*}
\Pi_\lambda L \ =\  \textstyle{\bigcup_{\beta<\lambda} \Pi_\beta L}
\ =\  \textstyle{\bigcup_{\alpha\leq\beta<\lambda} \Pi_\beta L }
\ =\  \textstyle{\bigcup_{\alpha\leq\beta<\lambda} \Pi_\alpha L }
\ =\  \Pi_\alpha L.
\end{equation*}
Since we already know that $\Pi_\lambda \varphi$
extends $\Pi_\alpha \varphi$,
this implies $\Pi_\alpha \varphi = \Pi_\lambda\varphi$.

\noindent\ref{L:collapse-2}.\ 
We know that $\Sigma_{\alpha+1}\varphi$ extends $\Sigma_\alpha\varphi$,
and that $\Sigma_{\alpha+1}\varphi=\Sigma(\Pi_\alpha\varphi)=\Pi_\alpha\varphi$
since the hierarchy has collapsed at $\Pi_\alpha\varphi$.
Hence $\Pi_\alpha\varphi$ extends $\Sigma_\alpha\varphi$.
With a similar reasoning
we see that $\Sigma_\alpha\varphi$ extends $\Pi_\alpha\varphi$.
Hence $\Pi_\alpha\varphi = \Sigma_\alpha\varphi$.

\noindent\ref{L:collapse-3}.\ 
Suppose the hierarchy has collapsed at~$Q_1$ and~$Q_2$.
We prove $Q_1 = Q_2$.

There is an~$\alpha$ such that  $Q_1 = \Pi_\alpha \varphi$
or $Q_1 = \Sigma_\alpha\varphi$,
and there is a~$\beta$ with 
$Q_2 = \Pi_\beta \varphi$ or $Q_2 = \Sigma_\beta \varphi$.
Without loss of generality, assume $\alpha \leq \beta$.

If $\alpha < \beta$, then we get $Q_1 = Q_2$ by~\ref{L:collapse-1}.
If $\alpha=\beta$, we get $Q_1 = Q_2$ by~\ref{L:collapse-2}.

\noindent\ref{L:collapse-4}.\ 
Suppose the hierarchy collapses.
We need to prove that the hierarchy has collapsed at some $\Pi_\alpha \varphi$.
Since the hierarchy collapses,
we know that the hierarchy has collapsed 
at some $\Pi_\alpha \varphi$ or at some $\Sigma_\alpha\varphi$.
In the former case, we are done.
In the latter case, $\Pi_{\alpha+1} \varphi = \Pi(\Sigma_\alpha \varphi)
= \Sigma_\alpha\varphi$,
so the hierarchy has collapsed at~$\Pi_{\alpha+1}\varphi$.
\end{proof}
%
%                  DEFINITION OF EXTENDIBLE
%
\begin{dfn}
\label{D:extendible}
Let $\vs{V}{L}\varphi{E}$ be a valuation system.
\begin{enumerate}
\item
\label{D:extendible-1}
We say that~$\varphi$ (or  $\vs{V}{L}\varphi{E}$) is \keyword{extendible}
if the hierarchy collapses.

\item
\label{D:extendible-2}
Suppose that $\varphi$ is extendible.
Then there is precisely one valuation
at which the hierarchy has collapsed
(see Lemma~\ref{L:collapse}\ref{L:collapse-3});
we denote it by
\begin{equation*}
\vs{V}{\overline L}{\overline\varphi}{E}.
\end{equation*}
\end{enumerate}
\end{dfn}
%
%                  OVERLINE VARPHI IS COMPLETE
%
\begin{lem}
\label{L:complete}
Let $\vs{V}{L}\varphi{E}$ be an extendible valuation system.
Then
\begin{equation*}
\vs{V}{\overline L}{\overline\varphi}{E}
\text{ is complete.}
\end{equation*}
\end{lem}
\begin{proof}
To prove that~$\overline\varphi$ is complete,
it suffices to show that~$\overline\varphi$
is $\Pi$-complete and $\Sigma$-complete.
We know that $\overline\varphi = \Pi\overline\varphi$
(since the hierarchy has collapsed at~$\overline\varphi$,
see Definition~\ref{D:extendible}\ref{D:extendible-2}
and Definition~\ref{D:collapse}\ref{D:collapse-1}),
and that $\Pi\overline\varphi$ is $\Pi$-complete
(see Lemma~\ref{L:Pi-complete}).
Hence $\overline\varphi$ is $\Pi$-complete.
Similarly,
$\overline\varphi$ is $\Sigma$-complete.
\end{proof}
%
%                  EXTENDIBLE IFF PI_ALEPH_1 EXTENDIBLE
%
\begin{lem}
\label{L:aleph1}
Let $\vs{V}{L}\varphi{E}$ be a valuation system.\\
If $\varphi$ is $\Pi_{\aleph_1}$-extendible,
then the hierarchy has collapsed at $\Pi_{\aleph_1}\varphi$.
\end{lem}
\begin{proof}
Suppose $\varphi$ is $\Pi_{\aleph_1}$-extendible.
We need to prove that the hierarchy has collapsed at~$\Pi_{\aleph_1}\varphi$.
For this, we must show that
(see Definition~\ref{D:collapse}\ref{D:collapse-1}),
\begin{equation*}
\Pi(\Pi_{\aleph_1}\varphi)
= \Pi_{\aleph_1}\varphi = \Sigma(\Pi_{\aleph_1}\varphi).
\end{equation*}

Let $a_1 \geq a_2 \geq \dotsb$ 
be a $\Pi_{\aleph_1}\varphi$-convergent sequence.
In order to show that $\Pi(\Pi_{\aleph_1}\varphi)=\Pi_{\aleph_1}\varphi$,
it suffices to prove that $\bw_n a_n \in \Pi_{\aleph_1}\varphi$.

Since $\aleph_1$ is a limit ordinal,
we know that $\Pi_{\aleph_1} L = \bigcup_{\alpha < \aleph_1} \Pi_\alpha L$
(see Proposition~\ref{P:hier}).
Define
\begin{equation*}
\alpha(n)\ \eqdf\ 
\min\,\{\, \alpha<\aleph_1\colon\, a_n \in \Pi_\alpha L\,\}
\end{equation*}
for every~$n$.
Now,
the set $\{\,\alpha(1),\,\alpha(2),\,\dotsc\,\}$ of
ordinals has a supremum,
\begin{equation*}
\xi\,\eqdf\,\bv_n\alpha(n).
\end{equation*}
Since~$\aleph_1$ is the smallest uncountable ordinal,
and $\alpha(n)<\aleph_1$,
we know that all~$\alpha(n)$ are countable.
Hence~$\xi$ is countable as well,
and so $\xi<\aleph_1$.

Now, we have $a_n \in \Pi_{\xi} L$
for all~$n\in \N$.  Hence
\begin{equation*}
\bw_n a_n \in \Pi(\Pi_{\xi} L) 
\ =\ \Pi_{\xi+1} L
\ \subseteq\ \Pi_{\aleph_1} L.\qedhere
\end{equation*}
\end{proof}
%
%
%
\begin{cor}
\label{C:aleph1}
Let $\vs{V}{L}\varphi{E}$ be a valuation system. Then
\begin{equation*}
\varphi\text{ is extendible }
\quad\iff\quad
\varphi\text{ is $\Pi_{\aleph_1}$-extendible.}
\end{equation*}
Moreover,
if $\varphi$ is extendible,
then $\overline\varphi = \Pi_{\aleph_1}\varphi$.
\end{cor}
\begin{proof}
Assume $\varphi$ is extendible
in order to show that $\varphi$ is $\Pi_{\aleph_1}$-extendible.
Then we know that the hierarchy collapses 
(see Definition~\ref{D:collapse}\ref{D:collapse-2}).
So it collapsed at some~$\Pi_\alpha \varphi$
(see Lemma~\ref{L:collapse}\ref{L:collapse-4}).
Pick an ordinal number~$\beta$ with $\beta > \alpha$
and $\beta > \aleph_1$.
Then $\varphi$ is $\Pi_\beta$-extendible
by Lemma~\ref{L:collapse}\ref{L:collapse-1}.
But $\aleph_1 <\beta$,
so $\varphi$ is also $\Pi_{\aleph_1}$-extendible.

Suppose $\varphi$ is $\Pi_{\aleph_1}$-extendible.
Then the hierarchy has collapsed at~$\Pi_{\aleph_1}$
by Lemma~\ref{L:aleph1}.
Hence 
$\varphi$ is extendible
and  $\overline\varphi = \Pi_{\aleph_1}\varphi$
(see Definition~\ref{D:extendible}\ref{D:extendible-2}).
\end{proof}

%
%                  Pi_\alpha is  MONOTONOUS
%
\begin{lem}
\label{L:alpha-monotonous}
Let $\vs{V}{L}\varphi{E}$
and $\vs{V}{C}\psi{E}$
 be a valuation systems.\\
Assume 
$\psi$ extends $\varphi$.
Then for every ordinal number $\alpha$, we have
\begin{alignat*}{3}
\text{$\psi$ is $\Pi_\alpha$-extendible}
\quad&\implies\quad
\text{$\varphi$ is $\Pi_\alpha$-extendible}
\quad&&\text{and}\quad
\text{$\Pi_\alpha\psi$ }&&\text{extends $\Pi_\alpha\varphi$} \\
\text{$\psi$ is $\Sigma_\alpha$-extendible}
\quad&\implies\quad
\text{$\varphi$ is $\Sigma_\alpha$-extendible}
\quad&&\text{and}\quad
\text{$\Sigma_\alpha\psi$ }&&\text{extends $\Sigma_\alpha\varphi$.}
\end{alignat*}
\end{lem}
\begin{proof}
We prove this lemma using induction on~$\alpha$.

\noindent\emph{(Zero)} For~$\alpha=0$, the proposition is trivial.

\noindent\emph{(Successor)} Let~$\alpha$ be an ordinal number such that
if $\psi$ is $\Sigma_\alpha$-extendible,
then
$\varphi$ is $\Sigma_\alpha$-extendible 
and $\Sigma_\alpha\psi$ extends $\Sigma_\alpha\varphi$.
Suppose~$\psi$
is $\Pi_{\alpha+1}$-extendible.
We prove
\begin{equation}
\label{eq:P:extension-1}
\text{ $\varphi$ is $\Pi_{\alpha+1}$-extendible}
\qquad\text{and}\qquad
\text{ $\Pi_{\alpha+1}\psi$ extends $\Pi_{\alpha+1}\varphi$.}
\end{equation}
Since $\psi$ is $\Pi_{\alpha+1}$-extendible,
we know that
\begin{equation*}
\text{ $\psi$ is $\Sigma_\alpha$-extendible}
\qquad\text{and}\qquad
\text{$\Sigma_\alpha\psi$ is~$\Pi$-extendible}.
\end{equation*}
By assumption, 
the former implies that $\varphi$ is $\Sigma_\alpha$-extendible
and $\Sigma_\alpha\psi$ extends $\Pi_\alpha\psi$;
by Lemma~\ref{L:Pi-monotonous},
the latter implies $\Sigma_\alpha\varphi$
is $\Pi$-extendible
and $\Pi(\Sigma_\alpha\psi)$ extends $\Pi(\Sigma_\alpha \varphi)$.
In other words,
we have proven Statement~\eqref{eq:P:extension-1}.

\noindent\emph{(Limit)}
Let $\lambda$ be a limit ordinal such that
for all $\alpha<\lambda$, we have
\begin{equation*}
\text{$\psi$ is $\Pi_\alpha$-extendible}
\quad\implies\quad
\text{$\varphi$ is $\Pi_\alpha$-extendible }
\quad\text{and}\quad
\text{$\Pi_\alpha\psi$ extends $\Pi_\alpha\varphi$.}
\end{equation*}
Further, 
assume $\psi$ is~$\Pi_\lambda$-extendible
in order to prove that
\begin{equation}
\label{eq:P:extension-2}
\text{ $\varphi$ is $\Pi_{\lambda}$-extendible}
\qquad\text{and}\qquad
\text{ $\Pi_{\lambda}\psi$ extends $\Pi_{\lambda}\varphi$.}
\end{equation}
Let $\alpha <\lambda$ be given.
Since $\psi$ is $\Pi_\lambda$-extendible,
we know that $\psi$ is $\Pi_\alpha$-extendible.
So by assumption, 
$\varphi$ is $\Pi_\alpha$-extendible,
and $\Pi_\alpha \psi$ extends $\Pi_\alpha \varphi$.

So we see that $\varphi$ is $\Pi_\lambda$-extendible.
Further
(since $\Pi_\lambda\psi$ extends $\Pi_\alpha \psi$),
we see that $\Pi_\lambda \psi$ extends all $\Pi_\alpha\varphi$.
Hence $\Pi_\lambda \psi$ extends $\Pi_\lambda \varphi$.
So we have proven~\eqref{eq:P:extension-2}.
\end{proof}
%
%                  OVERLINE PHI IS MONOTONOUS
%
\begin{prop}
\label{P:comp-monotonous}
Let $\vs{V}{L}\varphi{E}$
and $\vs{V}{C}\psi{E}$
 be a valuation systems.\\
Assume 
$\psi$ extends $\varphi$,
and 
$\psi$ is extendible.
Then 
\begin{alignat*}{3}
\text{$\varphi$ is extendible}
\qquad&&\text{and}\qquad
\text{$\overline\psi$ }&&\text{extends $\overline\varphi$}.
\end{alignat*}
\end{prop}
\begin{proof}
By Corollary~\ref{C:aleph1},
we get the conclusion from
Lemma~\ref{L:alpha-monotonous}
with $\alpha=\aleph_1$.
\end{proof}
%
%                  OVERLINE PHI IS MINIMAL
%
\begin{prop}
\label{P:comp-minimal}
Let $\vs{V}{L}\varphi{E}$
and $\vs{V}{C}\psi{E}$
 be a valuation systems.\\
Assume 
$\psi$ extends $\varphi$
and $\psi$ is complete.
Then 
\begin{alignat*}{3}
\text{$\varphi$ is extendible}
\qquad&&\text{and}\qquad
\text{$\psi$ }&&\text{extends $\overline\varphi$}.
\end{alignat*}
\end{prop}
\begin{proof}
Since~$\psi$ is complete,
$\psi$ is clearly extendible 
and $\overline\psi = \psi$.
(Indeed, the hierarchy (of~$\psi$) has already collapsed at~$\psi$ itself,
see Definition~\ref{D:extendible}.)
Hence $\varphi$ is extendible
and $\psi=\overline\psi$ extends $\overline\varphi$
by Proposition~\ref{P:comp-monotonous}.
\end{proof}
\end{document}
