\documentclass[main.tex]{subfiles}
\begin{document}
\section{Valuations}
\label{S:vals}
We derive some elementary properties of valuations in this section.
We start with some examples of valuations in Subsection~\ref{SS:vals_intro}.
We study the distance 
induced by a valuation~$\varphi$
in Subsection~\ref{SS:vals_d}, being
\begin{equation*}
\ld\varphi(x,y) \ =\ \varphi(x\vee y) - \varphi(x\wedge y).
\end{equation*}
In Subsection~\ref{SS:vals_eq}
we study the equivalence induced by this distance,
\begin{equation*}
x\approx y\quad\iff\quad \ld\varphi(x,y)=0.
\end{equation*}
The notion of distance 
is especially important to the remainder of this text.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Introduction}
\label{SS:vals_intro}
\noindent
%
%          DEFINITION OF MODULAR MAPS
%
\begin{dfn}

\label{D:val}
Let $L$ be a lattice. 
Let $E$ an ordered Abelian group
(see Section~\ref{S:ag}).
Let  $\varphi\colon L \ra E$ be a map.
We say that
\begin{enumerate}
\item
\label{D:val-mod}
$\varphi$ is \keyword{modular} provided that
\begin{equation*}
\varphi(a\wedge b) + \varphi(a \vee b)
\ =\ 
\varphi(a) + \varphi(b)
\qquad(a,b\in L);
\end{equation*}

\item
\label{D:val-val}
$\varphi$ is a \keyword{valuation}
provided that $\varphi$ is modular and order preserving.
\end{enumerate}
\end{dfn}

\begin{ex}
Let $\mathcal{F}$ be the set of finite subsets of~$\N$,
and for each $A\in \mathcal{F}$,
let $\#(A)$ be the number of elements of~$A$.
Then have 
\begin{equation*}
\#(A\cap B) + \#(A\cup B) \,=\, \#(A) + \#(B)
\qquad(A,B\in\mathcal{F}),
\end{equation*}
so obviously the map $\mathcal{F} \ra \N$
given by by $A\mapsto \#(A)$ is a valuation.
\end{ex}

%
%                  LEBESGUE MEASURE IS A VALUATION
%
\begin{ex}
\label{E:lmeas-val}
Let $\LA$ be the set of Lebesgue measurable
subsets of~$\R$ with finite Lebesgue measure.
Then~$\LA$ is a lattice of subsets of~$\R$.
Given~$A\in\LA$,
let $\Lmu$ denote the Lebesgue measure of~$A$.
Then $A\subseteq B \implies \Lmu(A)\leq \Lmu (B)$
and 
\begin{equation*}
\Lmu(A\cap B) + \Lmu(A\cup B) \,=\, \Lmu(A) + \Lmu(B)
\qquad(A,B\in\mathcal{A}).
\end{equation*}
So~$\Lmu$ is a valuation.
\end{ex}

%
%                  LEBESGUE INTEGRAL IS A VALUATION
%
\begin{ex}
\label{E:int-val}
Let~$\LF$ be the set of Lebesgue integrable functions on~$\R$.
It is a lattice ordered Abelian group.
The assignment $f\mapsto \int f$ 
yields an order preserving group homomorphism
$\Lphi\colon \LF\ra \R$.
To see that $\Lphi$ is modular (and a valuation),
note
\begin{equation*}
\min\{x,y\}+\max\{x,y\} \,=\, x+y\qquad(x,y\in \R).
\end{equation*}
So given $f,g\in \LF$,
we have $f\wedge g + f \vee g = f+ g$,
and hence
\begin{equation*}
\Lphi(f\wedge g) + \Lphi(f\vee g) 
\,=\,\Lphi(f\wedge g + f\vee g)
\,=\, \Lphi(f+g)
\,=\, \Lphi (f)+\Lphi(g).
\end{equation*}
\end{ex}

%
%                  ANY MAP ON A CHAIN IS MODULAR
%
\begin{ex}
Let $C$ be a chain,
i.e. a totally ordered set.
Then $C$ is a lattice with
\begin{equation*}
a\wedge b = \min\{a,b\},
 \qquad 
a\vee b = \max\{a,b\}.
\end{equation*}
One quickly sees that
\emph{any map $f\colon C\rightarrow E$
to an ordered Abelian group is modular.}
\end{ex}

%
%                  RING OF SUBSETS
%
\begin{ex}
\label{E:ring-val}
Let~$X$ be a set and let~$\mathcal{A}$ be a ring of subsets of~$X$.
That is,
\begin{equation*}
A\cap B,\qquad\qquad A\cup B,\qquad\qquad A\backslash B
\end{equation*}
are in~$\mathcal{A}$ for all $A,B\in\mathcal{A}$.
Then clearly~$\mathcal{A}$ is a lattice.

Let $E$ be an ordered Abelian group
and let~$\mu\colon \mathcal{A}\rightarrow E$ be a map.
Recall that~$\mu$ is \keyword{additive} if $\mu(A) + \mu(B) = \mu(A\cup B)$
for all $A,B\in\mathcal{A}$ with $A\cap B=\varnothing$.

\emph{If~$\mu$ is additive,
then $\mu$ is modular.}
Indeed,
let $A,B\in \mathcal{A}$ be given. We need to prove that
$\mu(A) + \mu(B) =\mu(A\cap B) + \mu(A\cup B)$
assuming~$\mu$ is additive.
We have
\begin{alignat*}{6}
\mu(A) + \mu(B) \,
  & =\, \mu(A\cap B \ \cup\ A\backslash B) + \mu(B) \\ 
  & =\, \mu(A\cap B) + \mu(A\backslash B)  + \mu(B)\qquad
    && \text{since } A\cap B \ \cap\ A\backslash B = \varnothing \\ 
  & =\, \mu(A\cap B) + \mu(A\backslash B \ \cup\ B ) 
    && \text{since } A\backslash B\ \cap\ B = \varnothing \\
  & =\, \mu(A\cap B) + \mu(A\cup B).
\end{alignat*}

Recall that~$\mu$ is \keyword{positive} whenever
$\mu(A)\in E^+$ for all~$A\in\mathcal{A}$.

\emph{If~$\mu$ is additive and positive,
then~$\mu$ is a valuation.}
Since~$\mu$ is additive,
$\mu$ is modular.
It remains to be shown that~$\mu$ is order preserving
(see Definition~\ref{D:val}).
Let $A\subseteq B$ from~$\mathcal{A}$ be given
in order to prove $\mu(A)\leq \mu(B)$.
We have
\begin{equation*}
(B\backslash A)\,\cup\, A\,=\,B,\qquad\qquad 
(B\backslash A)\,\cap\, A\,=\,\varnothing.
\end{equation*}
So by additivity, 
$\mu(B)=\mu(B\backslash A)+\mu(A)$.
Then $\mu(B)\geq \mu(A)$, since $\mu(B\backslash A)\geq 0$.
\end{ex}

%
%                  RING OF SIMPLE LEBESGUE SUBSETS
%
\begin{ex}
\label{E:smeas-val}
We describe a ring of subsets of~$\R$
and a positive and additive 
map~$\Smu\colon \SA\ra \R$ that 
will eventually
lead to the Lebesgue measure.

Let $\mathcal{S}$ be the set of all finite intervals of~$\R$,
\begin{equation*}
(a,b)\qquad [a,b)\qquad (a,b]\qquad [a,b]
\qquad\qquad\text{where }a<b.
\end{equation*}
Let~$\SA$ be the ring generated by~$\mathcal{S}$.
Every element~$A$ of~$\SA$ is of the form
\begin{equation*}
I_1 \,\cup \,\dotsb \,\cup\, I_N
\end{equation*}
where $I_1,\dotsc,I_N\in \SA$
are disjoint.
Let $\Smu(A)$ be given by
\begin{equation*}
\mu_{\mathrm L}(A) \ \eqdf\  |I_1| + \dotsb + |I_N|.
\end{equation*}
One can verify that the number $\Smu(A)$
only depends on~$A$ and not on the choice of~$I_1,\dotsc,I_N$.
Hence we obtain a map~$\Smu\colon \SA \ra \R$.
Almost by definition $\Smu$ is additive and positive.
Hence $\Smu\colon\SA\ra \R$ is a valuation 
(see Example~\ref{E:ring-val}).
\end{ex}


In Example~\ref{E:int-val},
we saw a group homomorphism that is modular.
In fact any group homomorphism
on a lattice ordered Abelian group is modular
(see Corollary~\ref{C:hom-val}\ref{C:hom-val-group}).
\begin{ex}
\label{E:1-valuation}
Let~$R$ be a lattice ordered Abelian group.
Then the identity map $\mathrm{id}_R$ is a valuation.
Indeed, $\mathrm{id}_R$ is modular by Lemma~\ref{L:1-valuation},
and clearly order preserving.
\end{ex}

\begin{lem}
\label{L:mod-comp}
Suppose we have the following situation
\begin{equation*}
\xymatrix{
L' \ar[r]^f&
L \ar[r]^\varphi&
E \ar[r]^g&
E',
}
\end{equation*}
where $L$, $L'$ are lattices,
$E$, $E'$ are ordered Abelian groups,
$f$ is a lattice homomorphism,
$\varphi$ a map,
and $g$ is a group homomorphism.
Then
\begin{enumerate}
\item
\label{L:mod-comp-mod}
$g\circ \varphi \circ f$ is modular
provided that $\varphi$ is modular;
\item
\label{L:mod-comp-val}
$g\circ \varphi \circ f$ is a valuation
provided that $\varphi$ is a valuation
and~$g$ is positive.
\end{enumerate}
\end{lem}
\begin{proof}
\noindent
\ref{L:mod-comp-mod}
\  Suppose~$\varphi$ is modular.
Let $a,b\in L'$ be given.
Writing $\varphi'= g\circ\varphi \circ f$,
we need to prove that
$\varphi'(a\wedge b)+\varphi'(a\vee b)=\varphi'(a)+\varphi'(b)$.
We have
\begin{alignat*}{3}
\varphi'(a) + \varphi'(b)
\ &=\ g(\varphi(f(a))) \,+\, g(\varphi(f(b))) \\
  &=\ g(\ \varphi( f(a)) + \varphi( f(b))\ ) \\
  &=\ g(\ \varphi(f(a)\wedge f(b)) + \varphi(f(a)\vee f(b))\ ) \\
  &=\ g(\  \varphi(f(a\wedge b)) \,+\, \varphi(f(a\vee b))\ ) \\
  &=\ g(\varphi(f(a\wedge b))) \,+\, g(\varphi(f(a\vee b))) \\ 
  &=\ \varphi'(a\wedge b) + \varphi'(a \vee b)
\end{alignat*}
\ref{L:mod-comp-val}
\  Suppose~$\varphi$ is a valuation
and~$g$ is positive.
We need to prove that~$\varphi'\eqdf g\circ \varphi\circ f$
is a valuation.
By part~\ref{L:mod-comp-mod}
we know that~$\varphi'$ is modular.
It remains to be shown that~$\varphi'$ is order preserving.
This is easy: $g$, $\varphi$, and~$f$ are all order preserving.
So $\varphi'=g\circ\varphi\circ f$ must be order preserving too.
\end{proof}

\begin{cor}
\label{C:hom-val}
Let~$R$ be a lattice ordered Abelian group.
\begin{enumerate}
\item
\label{C:hom-val-lat}
Let~$L$ be a lattice.
Any lattice homomorphism $f\colon L\ra R$ 
is a valuation;

\item
\label{C:hom-val-group}
Let $E$ be an ordered Abelian group
and $g\colon R\ra E$ a group homomorphism.
Then $g$ is modular.
Moreover,
if~$g$ is positive,
then~$g$ is a valuation.
\end{enumerate}
\end{cor}
\begin{proof}
Apply Lemma~\ref{L:mod-comp} to the following situations.
\begin{equation*}
\xymatrix{
L\ar[r]^f &
R\ar[r]^{\mathrm{id}_R} &
R\ar[r]^{\mathrm{id}_R} &
R &&
R\ar[r]^{\mathrm{id}_R} &
R\ar[r]^{\mathrm{id}_R} &
R\ar[r]^{g} &
E}
\end{equation*}
(Recall that $\mathrm{id}_R$ is a valuation,
see Example~\ref{E:1-valuation}.)
\end{proof}

%
%                  RIESZ SPACE OF FUNCTIONS
%
\begin{ex}
\label{E:val-riesz-space-of-functions}
Let $X$ be a set.
We say that $F\subseteq \R^X$
is  \emph{Riesz space of functions} if
\begin{equation*}
f\vee g,\quad\qquad 
f\wedge g,\quad\qquad
f+g,\quad\qquad 
\lambda \cdot f
\end{equation*}
are all in~$F$
where $f,g\in F$ and $\lambda \in \R$.
Then~$F$ is a lattice ordered Abelian group.

Let~$E$ be an ordered Abelian group
and let $\varphi\colon F\ra E$ be a positive linear map.
We see that~$\varphi$ is a valuation
 by Corollary~\ref{C:hom-val}\ref{C:hom-val-group}.
\end{ex}

%
%                  RIESZ SPACE OF STEP FUNCTIONS
%
\begin{ex}
\label{E:sint-val}
We  describe a Riesz space
of functions~$F_{\mathrm{L}}$ on~$\R$
and a positive linear map~$\varphi_{\mathrm{L}}\colon F_{\mathrm{L}}\ra \R$
that will eventually lead to the Lebesgue integral.

A \emph{step function} is a function~$f\colon \R\ra\R$
for which there are $s_1 < s_2 <\dotsb <s_N$ in~$\R$
such that $f$ is constant on each~$(s_n,s_{n+1})$
and $f$ is zero outside $[s_1,s_N]$.

Let $F_\mathrm{L}$ be the set of step functions.
One can easily see that~$F_\mathrm{L}$ is a Riesz space of functions.
Let $f\in F_\mathrm{L}$.
Let $s_1 < s_2 <\dotsb <s_N$
be such that $f$ is constant, say $c_n\in \R$,
 on~$(s_n,s_{n+1})$
and $f$ is zero outside $[s_1,s_N]$.
One can prove that 
\begin{equation}
\label{exp:step}
\sum_{n=1}^{N-1} \, c_n\cdot(s_{n+1} - s_n)
\end{equation}
does not depend on the choice
of~$s_1 < s_2 <\dotsb <s_N$.
So Expression~\eqref{exp:step}
given a map $\varphi_\mathrm{L}\colon F_\mathrm{L} \ra \R$.
This map is easily seen to be linear.

Consequently, $\varphi_\mathrm{L}\colon F_\mathrm{L}\ra \R$
is a valuation (see Example~\ref{E:val-riesz-space-of-functions}).
\end{ex}

%
%                  TOTIENT FUNCTION
%
\begin{ex}
\todo{Write this example on the Euler Totient function.}
\end{ex}


%
%                  EXAMPLE ON VECTOR SPACES
%
\begin{ex}
Up to this point
we have only seen valuations on distributive lattices.
We will now give an example
of a valuation on a non-distributive lattice.

Let $W$ be a vector space.
Let $L$ be the set of finite-dimensional linear subspaces
of~$W$ ordered by inclusion.
  Then~$L$ is a lattice, and 
for all~$A,B\in L$,
\begin{equation*}
A\wedge B \ =\ A\cap B,
\qquad A\vee B \ =\ \left<A\cup B\right>
\end{equation*}
where $\left< S \right>$ denotes the smallest
linear subspace containing $S$.
We have
\begin{equation*}
\dim (A\wedge B) \,+\, \dim(A\vee B)
\ =\ 
\dim A \,+\, \dim B 
\qquad\qquad(A,B\in L).
\end{equation*}
To see this,
apply the dimension theorem
to the map~$f\colon A\times B\ra A\vee B$ given by  $(a,b)\mapsto a+b$.
Hence the assignment $A \mapsto \dim A$
gives a valuation $\dim\colon L\ra \N$.

The lattice~$L$ might be distributive.
For instance, if~$W=\{ 0 \}$.
This occurs only seldomly:
if $W$ contains two linearly independent vectors,
then~$L$ is non-distributive.

Indeed,
let $v_1,v_2\in W$ be linearly independent vectors
and consider  $w\eqdf v_1 + v_2$.
One can verify that $v_i, w$ are linearly independent too.
So $\left< v_i \right> \cap \left< w \right> = \{0\}$.
Hence
\begin{equation*}
\left< w \right> \wedge(\left< v_1 \right>\vee\left< v_2 \right>)
\ =\ 
\left< v_1, v_2 \right>
\ \neq\ 
\{0\}
\ =\ 
(\left<w\right> \wedge \left<v_1\right>)
\,\vee\, (\left<w\right> \wedge \left<v_2\right>).
\end{equation*}
\end{ex}

We end this subsection
with some tame examples of valuations we 
need lateron.
\begin{ex}
\label{E:val-product}
Let $I=\{ 1,2\}$.
For each $i\in I$, 
let $L_i$ be a lattice,
$E_i$ an ordered Abelian group,
and $\varphi_i \colon L_i \ra E_i$
a valuation.
Then 
the map 
\begin{equation*}
\varphi_1 \times \varphi_2 \colon \,
L_1 \times L_2 \,\longrightarrow\, E_1 \times E_2
\end{equation*}
given by $(\varphi_1 \times\varphi_2)(a_1,a_2) = (\varphi_1(a),\varphi_2(b))$
for all~$a_i\in L_i$, is a valuation.

We call the valuation $\varphi_1 \times \varphi_2$
the \emph{product} of $\varphi_1$ and $\varphi_2$.
Of course,
one can similarly define a product of an $I$-indexed family
of valuations for any set~$I$.
\end{ex}

\begin{ex}
\label{E-val-opposite}
Let $L$ be a lattice.
If we reverse the order on~$L$,
i.e., consider the partial order on~$L$ 
given by $\smash{a \leq_{L^\mathrm{op}} b}
\iff a\geq_L b$,
then if a subset~$S\subseteq L$
has a supremum, $\bigvee S$,
then $\bigvee S$ is the 
\emph{infimum} of~$S$
with respect to~$\leq^\mathrm{op}$.
So we see that $\leq^\mathrm{op}$
gives us a lattice, $L^\mathrm{op}$.
(The \emph{opposite} lattice.)

Let~$E$ be an ordered Abelian group.
If we reverse the order on~$E$,
we obtain an ordered Abelian group~$E^\mathrm{op}$
with the same group structure,
but whose positive elements, $\smash{(E^\mathrm{op})^+}$,
are precisely the negative elements of~$E$.

Let~$\varphi\colon L\ra E$ be a modular map
(see Definition~\ref{D:val}).
Then one quickly sees that $\varphi$ is also modular considered as a map
$L^\mathrm{op} \ra E$.
However,
$\varphi\colon L^\mathrm{op}\ra E$
is a valuation (that is, also order preserving)
if and only if $\varphi\colon L\ra E$
is order \emph{reversing},
i.e., $a\leq b\implies \varphi(a)\geq \varphi(b)$ for
all $a,b\in L$.

Of course,
if $\varphi\colon L\ra E$ is a valuation,
then $\varphi$ is a valuation $L^\mathrm{op}\ra E^\mathrm{op}$.
\end{ex}

It is interesting to note that
there are some `connections'
between
modular maps (see Definition~\ref{D:val})
and \emph{modular lattices}.
Recall that a lattice~$L$ is modular if
\begin{equation*}
\ell \vee (a \wedge u) \ =\ (\ell \vee a) \wedge u
\end{equation*}
for all~$\ell,u,a \in L$ with $\ell \leq u$.
One such connection is given by the following lemma.
%
%                  LEMMA ON MODULARITY FOR MODULAR MAPS
%
\begin{lem}
\label{L:modular-map-modular}
For $\ell,u\in L$ with $\ell\leq u$, we have
\begin{equation}
\label{eq:modular-map}
\varphi(\,\ell \vee (a \wedge u)\,) 
\ =\ 
\varphi(\,(\ell\vee a)\wedge u\,)
\qquad (a\in L).
\end{equation}
\end{lem}
\begin{proof}
The trick is to consider the expression 
$\varphi(\ell) + \varphi(a) + \varphi(u)$.
On the one hand,
\begin{align*}
\varphi(\ell) + \varphi(a) + \varphi(u)
\ &=\ \varphi(\ell\wedge a) + \varphi(\ell \vee a) + \varphi(u) \\
\ &=\ \varphi(\ell \wedge a)
      + \varphi(\,(\ell\vee a)\wedge u\,)
      + \varphi(a\vee u),
\end{align*}
where we have used modularity twice.
On the other hand,
\begin{align*}
\varphi(\ell) + \varphi(a) + \varphi(u)
\ &=\ \varphi(\ell) + \varphi(a\wedge u) + \varphi(a \vee u) \\
\ &=\ \varphi(\ell \wedge a)
      + \varphi(\,\ell\vee (a\wedge u)\,)
      + \varphi(a\vee u).
\end{align*}
The difference,
$\varphi(\,(\ell\vee a)\wedge u\,)
- \varphi(\,\ell\vee (a\wedge u)\,)$,
must be zero.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%j%
%
%
%                  DISTANCE INDUCED BY 
%                     a valuation
%
%
\subsection{Distance induced by a valuation}
\label{SS:vals_d}
In this subsection,
we derive some facts 
concerning the following notion
of distance induced by a valuation.
\begin{dfn}
\label{D:d}
Let $E$ be an ordered Abelian group.
Let $L$ be a lattice.\\
Let $\varphi\colon L \ra E$ be a valuation.
Define $\ld\varphi\colon L\times L \ra E$ by
\begin{equation*}
\ld\varphi(a,b)\ =\  \varphi(a\vee b) - \varphi(a \wedge b)
\qquad\quad(a,b\in L).
\end{equation*}
\end{dfn}

To give the name ``distance'' for~$\ld\varphi$ some credibility,
we will prove that $\ld\varphi$ 
is a pseudometric (see Lemma~\ref{L:d-metric}).
After that,
we turn our attention to the following fact,
we will use often.
Given~$a\in L$, the map $x\mapsto a\wedge x$ is a \emph{contraction}, i.e.,
\begin{equation*}
\ld\varphi(\,a\wedge x,\, a\wedge y\,)
\ \ \leq\ \ \ld\varphi(x,y)
\qquad\quad(x,y\in L).
\end{equation*}
In fact,
we will prove the following, stronger, statement
(see Lemma~\ref{L:curry-wc-unif}).
\begin{equation*}
\ld\varphi(\,a\wedge x,\, a\wedge y\,) \ +\ 
\ld\varphi(\,a\vee x,\, a\vee y\,)
\ \ \leq\ \ \ld\varphi(x,y)
\qquad\quad(x,y\in L).
\end{equation*}

Before we do all this,
let us consider some examples.
%
%                  EXAMPLES OF DISTANCE
%
\begin{ex}
\label{E:d-riesz}
Let~$E$ be an ordered Abelian group.
Let $F$ be a Riesz space of functions,
and let $\varphi\colon F\ra E$ be
a positive and linear map
(see Example~\ref{E:val-riesz-space-of-functions}).

Let $f,g\in F$ be given.
The distance between $f$ and $g$ is the usual one,
\begin{equation*}
\ld\varphi(f,g) \,=\, \|f-g\|_1 \ \eqdf\ \varphi(\,|f-g|\,).
\end{equation*}
To see this,
note that since~$\varphi$ is linear,
we have 
\begin{equation*}
\ld\varphi(f,g) \,=\, 
\varphi(f\vee g) - \varphi(f\wedge g)
\,=\, \varphi(f\vee g - f\wedge g).
\end{equation*}
Further,
since we have the identity $\max\{x,y\} - \min\{x,y\} = |x-y|$
for reals $x,y$,
we have the identity $f\vee g - f\wedge g = |f-g|$
for functions.
\end{ex}
\begin{ex}
Let $E$ be an ordered Abelian group.
Let $\mathcal{A}$ be a ring of sets,
and let $\mu\colon \mathcal{A}\ra E$
be a positive additive map
(see Example~\ref{E:ring-val}).
Let $A,B\in\mathcal{A}$.
We have
\begin{equation*}
\ld\mu(A,B) \ =\ \mu(A \ominus B),
\end{equation*}
where $A\ominus B \eqdf A\backslash B \,\cup\, B\backslash A$
is the \emph{symmetric difference} of~$A$ and~$B$.
To see this, note that $A\cup B$ is the disjoint union
of $A\ominus B$ and $A\cap B$. So
since $\mu$ is additive, 
\begin{equation*}
\mu(A\cup B) \ =\ \mu(A\ominus B) + \mu(A\cap B).
\end{equation*}
\end{ex}
%
%                  LEMMA ON THE DISTANCE
%
\begin{lem}
\label{L:d-metric}
Let $E$ be an ordered Abelian group.\\
Let $L$ be a lattice,
and let $\varphi\colon L \ra E$ be a valuation.
We have:
\begin{enumerate}
\item \label{d-metric_pos}
$\ld\varphi(a,b)\geq 0$ for all~$a,b\in L$.
\item\label{d-metric_self} 
$\ld\varphi(a,a)=0$ for all $a\in L$.
\item\label{d-metric_sym}
$\ld\varphi(a,b)=\ld\varphi(b,a)$ for all $a,b\in L$.
\item\label{d-metric_triangle}
$\ld\varphi(a,b)\leq\ld\varphi(a,z)+\ld\varphi(z,b)$
for all $a,b,z\in L$.
\end{enumerate}
\end{lem}
\begin{proof}
Only point~\ref{d-metric_triangle} requires some work.
Let $a,b,z\in L$ be given.
We want to show that $\ld\varphi(a,b) \leq \ld\varphi(a,z)+\ld\varphi(z,b)$.
In other words:
\begin{equation}
\label{eq:d-metric}
\varphi(a\vee b) + \varphi(a\wedge z) + \varphi (z\wedge b)
\ \leq\ 
\varphi (a\vee z) + \varphi(z\vee b) + \varphi (a\wedge b)\text{.}
\end{equation}
By modularity,
the left-hand side equals
\begin{equation*}
\varphi(a\vee b) 
 + \varphi(\,(a\wedge z)\vee(b\wedge z)\,)
 + \varphi(a\wedge b\wedge z).
\end{equation*}
On the other hand,
using modularity
the right-hand side of Inequality~\eqref{eq:d-metric} becomes
\begin{equation*}
\varphi(a\vee b\vee z)
 + \varphi(\,(a\vee z)\wedge(b\vee z)\,)
 + \varphi(a\wedge b).
\end{equation*}
Note that $a\vee b \leq a\vee b\vee z$,
and $(a\wedge z)\vee (b\wedge z) \leq z \leq (a\vee z)\wedge (b\vee z)$,
and $a\wedge b\wedge z \leq a\wedge b$,
so that the monotonicity of~$\varphi$ yields Inequality~\eqref{eq:d-metric}.
\end{proof}
%
%                  DEFINITION HAUSDORFF
%
We might have $\ld\varphi(a,b) = 0$ while $a\neq b$
(see Example~\ref{E:eq-int}).
So in general, $\ld\varphi$ is not a metric
(but merely a  \emph{pseudo}metric).
Those $\varphi$ for which~$\ld\varphi$ is a metric
turn out to be useful. So let us give them a name.
\begin{dfn}
\label{D:val_Hausdorff}
Let $L$ be a lattice.
Let $E$ be an ordered Abelian group.\\
Let $\varphi\colon L \ra E$ be a valuation.
We say $\varphi$ is \keyword{Hausdorff}
provided that
\begin{equation*}
\ld\varphi(a,b) = 0 
\quad \implies \quad a=b \qquad\qquad(a,b\in L).
\end{equation*}
\end{dfn}
We return to Hausdorff valuations in Subsection~\ref{SS:vals_eq}.
%
%                  CURRY-WC-UNIF
%
\begin{lem}
\label{L:curry-wc-unif}
Let $E$ be an ordered Abelian group.\\
Let $L$ be a lattice,
and $\varphi\colon L \ra E$ a valuation.
We have
\begin{equation*}
\ld\varphi(a\wedge z,\,b\wedge z)\,+\,
 \ld\varphi(a\vee z,b\vee z)\ \leq\ \ld\varphi(a,b),
\end{equation*}
where $a,b,z\in L$.
\end{lem}
\begin{proof}
By expanding Definition~\ref{D:d},
we see that 
we need to prove that
\begin{equation}
\label{eq:curry-wc-unif}
\begin{split}
\varphi(a\vee b\vee z) \,+\,
\varphi(\,(a\wedge z)\vee(b\wedge z)\,) & \,+\,
\varphi(a\wedge b) \\
\ \leq\ 
\varphi(a \vee b)  \,+\, &
\varphi(\,(a\vee z)\wedge(b\vee z)\,) \,+\,
\varphi(a\wedge b \wedge z)
\end{split}
\end{equation}
By modularity,
the left-hand side equals
\begin{equation*}
\varphi(a\vee b \vee z) \,+\,
\varphi(\, (a\wedge z) \vee (b\wedge z) \vee (a\wedge b)\,) \,+\,
\varphi(\,a\wedge b\wedge ( (a \wedge z) \vee (b\wedge z) )\,).
\end{equation*}
To simplify the above expression,
we prove that
$a\wedge b\wedge ( (a \wedge z) \vee (b\wedge z)=a\wedge b\wedge z$.
To this end, note that
$a\wedge z \,\leq\, (a\wedge z)\vee (b\wedge z) \leq z$
so that 
\begin{equation*}
a\wedge b\wedge z
\, =\, a\wedge b \wedge (a\wedge z)
\,\leq\, a\wedge b \wedge ((a\wedge z)\vee (b\wedge z))
\,\leq\, a\wedge b \wedge z.
\end{equation*}
Hence the left-hand side of Inequality~\eqref{eq:curry-wc-unif}
equals
\begin{equation*}
\varphi(a\vee b \vee z) \,+\,
\varphi(\, (a\wedge z) \vee (b\wedge z) \vee (a\wedge b)\,) \,+\,
\varphi(\,a\wedge b\wedge z\,).
\end{equation*}
In a similar fashion,
one can show that the right-hand side of Inequality~\eqref{eq:curry-wc-unif}
equals
\begin{equation*}
\varphi(a\vee b \vee z) \,+\,
\varphi(\, (a\vee z) \wedge (b\vee z) \wedge (a\vee b)\,) \,+\,
\varphi(\,a\wedge b\wedge z\,).
\end{equation*}
So in order to prove Inequality~\eqref{eq:curry-wc-unif},
we must show that
\begin{equation*}
\varphi(\, (a\wedge z) \vee (b\wedge z) \vee (a\wedge b)\,) 
\ \leq\ 
\varphi(\, (a\vee z) \wedge (b\vee z) \wedge (a\vee b)\,).
\end{equation*}
Since $\varphi$ is order preserving,
it suffices to show that 
\begin{equation*}
 (a\wedge z) \vee (b\wedge z) \vee (a\wedge b)
\ \leq\ 
  (a\vee z) \wedge (b\vee z) \wedge (a\vee b).
\end{equation*}
Writing $c_1 = a$, $c_2 = b$, and $c_3 = z$,
we must prove that 
\begin{equation*}
\bv_{i\neq j}\, c_i \wedge c_j \ \leq\ \bw_{k\neq \ell}\, c_k \vee c_\ell.
\end{equation*}
That is,
we must show that 
$c_i \wedge c_j \leq c_k \vee c_\ell$
for given $i\neq j$ and $k\neq \ell$.
Now, note
\begin{equation*}
\# (\{ i,j \} \cap \{ k,\ell\}) + \#\{i,j,k,\ell\}
\ = \ \# \{i,j\} + \# \{k,\ell \}
\ = \ 4.
\end{equation*}
Since  $\# \{ i,j,k,\ell \}\leq 3$,
we see that $\# \{ i,j \} \cap \{ k,\ell\} \geq 1$.
So pick $m\in  \{ i,j \} \cap \{ k,\ell\}$.
Then $c_i \wedge c_j \leq c_m \leq c_k \vee c_\ell$.
\end{proof}
%
%                  WC-UNIF
%
\begin{lem}
\label{L:wv-unif}
Let $E$ be an ordered Abelian group.\\
Let $L$ be a lattice,
and $\varphi\colon L \ra E$ a valuation.
Then we have
\begin{equation*}
\ld\varphi(a\wedge w,b\wedge z) \,+\, \ld\varphi(a \vee w,b\vee z)
\ \leq\ 
\ld\varphi(a,b) \,+\, \ld\varphi(w,z),
\end{equation*}
where $a,b,w,z\in L$.
\end{lem}
\begin{proof}
By the triangle inequality (point~\ref{d-metric_triangle}
of Lemma~\ref{L:d-metric}),
we have
\begin{equation}
\label{eq:L:d-metric-1}
\begin{alignedat}{3}
\ld\varphi(a\wedge w,\,b\wedge z)\ &\leq\ 
\ld\varphi(a\wedge w,\,b\wedge w)\,+\,
\ld\varphi(b\wedge w,\,b\wedge z), \\
\ld\varphi(a\vee w,\,b\vee z)\ &\leq\ 
\ld\varphi(a\vee w,\,b\vee w)\,+\,
\ld\varphi(b\vee w,\,b\vee z).
\end{alignedat}
\end{equation}
On the other hand, Lemma~\ref{L:curry-wc-unif} gives us
\begin{equation}
\label{eq:L:d-metric-2}
\begin{alignedat}{3}
\ld\varphi(a\wedge w,b\wedge w) + \ld\varphi(a\vee w,b\vee w)
   \ &\leq\ \ld\varphi(a,b), \\
\ld\varphi(b\wedge w,b\wedge z) + \ld\varphi(b\vee w, b\vee z)
   \ &\leq\ \ld\varphi(w,z).
\end{alignedat}
\end{equation}
The sum of the right-hand sides of Equation~\eqref{eq:L:d-metric-1}
equals the sum of the left-hand sides of Equation~\eqref{eq:L:d-metric-2}.
Hence
$\ld\varphi(a\wedge w,b\wedge z) + \ld\varphi(a \vee w,b\vee z)
\leq 
\ld\varphi(a,b) + \ld\varphi(w,z)$.
\end{proof}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%                  EQUIVALENCE INDUCED BY A VALUATION
%
\subsection{Equivalence induced by a valuation}
\label{SS:vals_eq}
In measure theory
two functions are considered equivalent
if they are equal almost everywhere.

In this subsection, we extend this notion of equivalence 
to valuations. 
%
%                  DEFINITION OF EQUIVALENCE
%
\begin{dfn}
\label{D:eq}
Let $E$ be an ordered Abelian group.
Let $L$ be a lattice.\\
Let $\varphi\colon L\ra E$ be a valuation.
We define $\approx$ to be 
the binary relation on~$L$ given by
\begin{equation*}
a \approx b
\quad\iff\quad
\ld\varphi(a,b)=0\qquad\qquad(a,b\in L).
\end{equation*}
\end{dfn}
\begin{rem}
$\varphi$ is Hausdorff (see Definition~\ref{D:val_Hausdorff})
iff $a \approx b\iff a = b$.
\end{rem}
%
%                  EXAMPLES OF EQUIVALENCE
%
\begin{ex}
\label{E:eq-int}
We consider the Lebesgue integral $\Lphi\colon \LF\ra \R$
(see Example~\ref{E:int-val}).

Let $f,g\in \LF$ be given.
By Example~\ref{E:d-riesz} 
we know that
\begin{equation*}
f\approx g\quad\iff\quad \Lphi(\,|f-g|\,)=0.
\end{equation*}
For any $h\in \LF$
with $h\geq 0$,
we have that $\Lphi(h)=0$
iff $h(x)=0$ for almost all~$x$.
(That is, $\smash{ \Lmu(\{ x\in\R\colon h(x)\neq 0 \})} $
is zero.)
So we see that 
\begin{equation*}
f\approx g\quad\iff\quad f(x) = g(x)\quad\text{for almost all~$x$}.
\end{equation*}
\end{ex}
%
%                  PROP ON EQUIVALENCE
%
\begin{prop}
\label{P:eq}
Let $E$ be an ordered Abelian group.
Let $L$ be a lattice.\\
Let $\varphi\colon L\ra E$ be a valuation.
Let $\approx$ be as in Definition~\ref{D:eq}.
\begin{enumerate}
\item
\label{P:eq-1}
The relation $\approx$ is an equivalence.
\item 
\label{P:eq-2}
Let~$a_1,a_2\in L$ with 
with $a_1\approx a_2$ be given.
Then $\varphi(a_1)=\varphi(a_2)$.
\item
\label{P:eq-3}
Let $a_1,a_2 \in L$ with $a_1\approx a_2$,
and let $b_1,b_2 \in L$ with $b_1 \approx b_2$ be given.
Then
\begin{equation*}
a_1 \wedge b_1 \,\approx\, a_2 \wedge b_2
\qquad\text{and}\qquad 
a_1 \vee b_1 \,\approx\, a_2 \vee b_2.
\end{equation*}
\item
\label{P:eq-4}
Let $a_1,a_2 \in L$ with $a_1\approx a_2$,
and let $b_1,b_2 \in L$ with $b_1 \approx b_2$ be given.
Then
\begin{equation*}
\ld\varphi(a_1,b_1)\ =\ \ld\varphi(a_2,b_2).
\end{equation*}
\end{enumerate}
\end{prop}
\begin{proof}
\ref{P:eq-1}\ 
The relation~$\approx$ is clearly reflexive
and symmetric. So to prove~$\approx$ is an equivalence relation,
we will only show that~$\approx$ is transitive.
Let $a,b,c\in L$ with $a\approx b\approx c$ be given.
We must show that $a\approx c$.
Or in other words, $\ld\varphi(a,c) = 0$.

By Lemma~\ref{L:d-metric}, points~\ref{d-metric_pos}
and~\ref{d-metric_triangle},
we get
\begin{equation}
\label{eq:P:eq-1}
0 \ \leq\ \ld\varphi(a,c) \ \leq\ 
\ld\varphi(a,b) + \ld\varphi(b,c).
\end{equation}
But 
$\ld\varphi(a,b)=0$ and $\ld\varphi(b,c)=0$,
since $a\approx b$ and $b\approx c$, respectively.

So we see that Statement~\eqref{eq:P:eq-1} implies $\ld\varphi(a,c)= 0$.
Hence $a\approx c$.

\vspace{.3em}
\ref{P:eq-2}\ 
Let $a_1,a_2\in L$ with $a_1\approx a_2$ be given.
We must prove $\varphi(a_1) = \varphi(a_2)$.

Let $i\in\{1,2\}$ be given. 
Note that $a_1 \wedge a_2 \leq a_i \leq a_1 \vee a_2$.
So we have 
\begin{equation}
\label{eq:P:eq-2}
\varphi(a_1 \wedge a_2) \ \leq\ \varphi(a_i) \ \leq\  \varphi(a_1 \vee a_2).
\end{equation}
Since $\ld\varphi(a_1,a_2)=0$,
we know that $\varphi(a_1 \vee a_2) = \varphi(a_1 \wedge a_2)$.
So Statement~\eqref{eq:P:eq-2}
implies that
$\varphi(a_1 \vee a_2) = \varphi(a_i) = \varphi(a_1 \wedge a_2)$.
Hence $\varphi(a_1) = \varphi(a_2)$.

\vspace{.3em}
\ref{P:eq-3}
Let $a_1,a_2 \in L$ with $a_1 \approx a_2$ be given.
Let $b_1,b_2 \in L$ with $b_1 \approx b_2$ be given.
We will only show that $a_1 \wedge b_1 \approx a_2 \wedge b_2$;
the proof of $a_1 \vee b_1 \approx a_2 \vee b_2$ is similar.

Note that we have the following inequalities by 
Lemma~\ref{L:d-metric}\ref{d-metric_pos}
and Lemma~\ref{L:wv-unif}.
\begin{equation}
\label{eq:P:eq-3}
0\ \leq\ 
\ld\varphi(a_1 \wedge b_1, a_2\wedge b_2)
\ \leq\ 
\ld\varphi(a_1,b_1) + \ld\varphi(a_2,b_2)
\end{equation}
Since $a_1 \approx a_2$ and $b_1 \approx b_2$,
we have $\ld\varphi(a_1,b_1)=0$
and $\ld\varphi(a_2,b_2)=0$,
respectively.
Hence Statement~\eqref{eq:P:eq-3} implies 
$\ld\varphi(a_1\wedge b_1, a_2 \wedge b_2)=0$.
So $a_1 \wedge b_1 \approx a_2 \wedge b_2$.

\vspace{.3em}
\ref{P:eq-4}
Let $a_1,a_2 \in L$ with $a_1 \approx a_2$ be given.
Let $b_1,b_2 \in L$ with $b_1 \approx b_2$ be given.
We must prove that $\ld\varphi(a_1,b_1)=\ld\varphi(a_2,b_2)$.
Note that
by point~\ref{P:eq-3}
we have
\begin{alignat*}{5}
a_1 \wedge b_1 &\approx a_2 \wedge b_2&
\qquad&\text{and}\qquad&
a_1 \vee b_1 &\approx a_2 \vee b_2.&
\shortintertext{%
So by point~\ref{P:eq-2} of this lemma, we get}
\varphi(a_1 \wedge b_1) &= \varphi( a_2 \wedge b_2)&
\qquad&\text{and}\qquad&
\varphi(a_1 \vee b_1)&=\varphi(a_2 \vee b_2).&
\end{alignat*}
So if we unfold Definition~\ref{D:d},
we see that
\begin{alignat*}{5}
\ld\varphi(a_1,b_1)
\ &=\ 
\varphi(a_1 \vee b_1) - \varphi(a_1 \vee b_1) \\
\ &=\ 
\varphi(a_2 \vee b_2) - \varphi(a_2 \vee b_2)
\ =\ 
\ld\varphi(a_2,b_2).\qedhere
\end{alignat*}
\end{proof}
%
%                  THE QUOTIENT LATTICE
%
When studying the Lebesgue integrable functions, $\LF$,
it is sometimes convenient
to consider the space~$\mathrm{L}^1 = \qvL{\LF}$
of integrable functions modulo equal-almost-everywhere
(see Example~\ref{E:eq-int}).
Of course,
one can consider the space $\qvL{L}$ 
for any valuation~$\varphi\colon L\ra E$.
We list some of the properties of~$\qvL{L}$
in Proposition~\ref{P:quotient-lattice}.
\begin{prop}
\label{P:quotient-lattice}
Let $E$ be an ordered Abelian group.\\
Let $L$ be a lattice.
Let $\varphi\colon L \ra E$ be a valuation.
Let $\approx$ be as in Definition~\ref{D:eq}.\\
Let $\qvL{L}$ denote the quotient set,
and let $q\colon L\ra \qvL{L}$ be the
quotient map. Then:
\begin{enumerate}
\item
The set
$\qvL{L}$ is lattice if
the operations are given by
\begin{equation*}
qa \wedge qb \,=\, q(a\wedge b),
\qquad
 qa \vee qb \,=\, q(a\vee b)\qquad\quad(a,b\in L).
\end{equation*} 
Then, in particular, $q\colon L\ra \qvL{L}$ is a lattice homomorphism.

\item
There is a unique map $\psi\colon\qvL{L}\ra E$
such that
\begin{equation*}
\psi(q(a)) \ =\ \varphi(a)\qquad\quad(a\in L).
\end{equation*}
This map ---
let us denote it by
 $\qvphi\varphi\colon \qvL{L}\ra E$ --- 
is a valuation.

\item
We have
\ $\ld{\qvphi{\varphi}}(\,qa,\,qb\,) \,=\,\ld\varphi(a,b)$\ \ 
for all \ $a,b\in L$.

\item
We have
\ $\ld{\qvphi{\varphi}}(\mathfrak{a},\mathfrak{b})=0
\ \implies \ 
\mathfrak{a} = \mathfrak{b}$\ \ 
for all \ $\mathfrak{a},\mathfrak{b}\in \qvL{L}$.
\end{enumerate}
\end{prop}
\begin{proof}
Follows from Proposition~\ref{P:eq}.
We leave the verification to the reader.
\end{proof}
\begin{rem}
Note that $\qvphi\varphi$ is Hausdorff 
(see Definition~\ref{D:val_Hausdorff}).
\end{rem}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%                  C O M P L E T E      V A L U A T I O N S 
%
%
%
\section{Complete Valuations}
\label{S:complete-val}
We now turn to the study of complete valuations
(see Definition~\ref{D:complete-val}).
Among all valuations
the complete valuations resemble
the Lebesgue measure
and the Lebesgue integral most closely.
To support this claim,
we will
prove generalisations of
some of the classical convergence theorems of integration
in Subsection~\ref{SS:complete-val_convergence}.

The notion of complete valuation is not at the end of the 
road.  We will study the
slightly more sophisticated 
\emph{valuation systems}
(see Definition~\ref{D:system})
and \emph{complete valuation systems}
(see Definition~\ref{D:system-complete})
in Section~\ref{S:valuation-systems}.

%
%                  PHI-CONVERGENCE
%
\subsection{Introduction}
\begin{dfn}
\label{D:phi-conv}
Let $E$ be an ordered Abelian group.\\
Let $L$ be a lattice, and let $\varphi\colon L \ra E$ be a valuation.

Consider a sequence
$a_1 \geq a_2 \geq \dotsb$ from~$L$.
We say
\begin{equation*}
a_1 \geq a_2 \geq \dotsb \text{ is \keyword{$\varphi$-convergent}}
\qquad\text{if}\qquad \bw_n \varphi(a_n)\text{ exists.}
\end{equation*}

Similarly,
if
$b_1 \leq b_2 \leq \dotsb$ is
a sequence in~$L$, then 
\begin{equation*}
b_1 \leq b_2 \leq \dotsb \text{ is \keyword{$\varphi$-convergent}}
\qquad\text{if}\qquad \bv_n \varphi(b_n)\text{ exists.}
\end{equation*}
\end{dfn}
%
%                  COMPLETE VALUATION
%
\begin{dfn}
\label{D:complete-val}
Let $E$ be an ordered Abelian group. Let $L$ be a lattice.\\
Let $\varphi\colon L \ra E$ be a valuation.
We say $\varphi$ is \keyword{$\Pi$-complete} if
\begin{alignat*}{5}
a_1 \geq a_2 \geq \dotsb \text{$\varphi$-convergent }
  \quad &\implies \quad 
  & \bw_n a_n &\text{ exists}\quad 
  &&\text{and}\quad
  &\varphi(\,\bw_n a_n\,) &= \bw_n \varphi(a_n);
\shortintertext{%
We say $\varphi$ is \keyword{$\Sigma$-complete} if
}
b_1 \leq b_2 \leq \dotsb \text{$\varphi$-convergent }
  \quad &\implies \quad 
  & \bv_n b_n &\text{ exists}\quad 
  &&\text{and}\quad
  &\varphi(\,\bv_n b_n\,) &= \bv_n \varphi(b_n).
\end{alignat*}
We say $\varphi$ is \keyword{complete}
if $\varphi$ is both $\Pi$-complete and $\Sigma$-complete.
\end{dfn}
%
%                  THE LEBESGUE MEASURE IS A COMPLETE VALUATION
%
\begin{ex}
\label{E:lmeas-complete-val}
The Lebesgue measure~$\Lmu$
(see Example~\ref{E:lmeas-val})
is a complete valuation.
We must show that~$\Lmu$ is
both $\Pi$-complete and $\Sigma$-complete
(see Definition~\ref{D:complete-val}).

Let us prove~$\Lmu$ is $\Sigma$-complete.
Let $B_1 \subseteq B_2 \subseteq \dotsb$ in~$\LA$
be $\Lmu$-convergent.
We must prove that $\bv_n B_n$ exists in~$\LA$
and that 
$\Lmu(\bv_n B_n) = \bv_n \Lmu(B_n)$.
Note that $\bigcup_n B_n$
is Lebesgue measurable and that
\begin{equation}
\label{eq:E:lmeas-complete-val-1}
\textstyle
\Lmu(\,\bigcup_n B_n\,) \ =\  \bv_n \Lmu(B_n).
\end{equation}
So we are done if we prove that $\bigcup_n B_n = \bv_n B_n$.
We know that $\bigcup_n B_n$ is of
finite Lebesgue measure
by Equation~\eqref{eq:E:lmeas-complete-val-1}.
Hence $\bigcup_n B_n \in \LA$.
Since $\bigcup_n B_n$ is the smallest subset of~$\R$
containing all~$B_n$
(i.e. $\bigcup_n B_n$ is the supremum of the~$B_n$ in~$\wp \R$),
$\bigcup_n B_n$ is also the smallest
subset \emph{of finite Lebesgue measure} containing all~$B_n$.
So
 $\bigcup_n B_n = \bv_n B_n$.
Hence $\Lmu$ is $\Sigma$-complete.

Using a similar reasoning one can prove that $\Lmu$
is $\Pi$-complete.
\end{ex}
%
%                  THE LEBESGUE INTEGRAL IS A COMPLETE VALUATION
%
\begin{ex}
\label{E:int-complete-val}
The Lebesgue integral~$\Lphi$
(see Example~\ref{E:int-val})
is a complete valuation.
We must show that~$\Lmu$ is
both $\Pi$-complete and $\Sigma$-complete
(see Definition~\ref{D:complete-val}).

Let us prove~$\Lmu$ is $\Sigma$-complete.
Let $f_1\leq f_2 \leq\dotsb$ in~$\LF$
be $\Lphi$-convergent.
We must prove that $\bv_n f_n$ exists in~$\LF$
and that 
\begin{equation*}
\Lphi(\bv_n f_n) \ =\  \bv_n \Lphi(f_n).
\end{equation*}
Of course,
this follows immediately from
Levis Monotone Convergence Theorem;
the supremum $\bv_n f_n$ in~$\LF$ is simply
the pointwise supremum
(which is the supremum of~$f_1\leq f_2 \leq\dotsb$ in $\E^\R$).
So we see $\Lphi$ is $\Sigma$-complete.

With a similar argument one can see that $\Lphi$ is $\Pi$-complete.
\end{ex}
%
%                  THE SIMPLE MEASURABLE SETS ARE NOT COMPLETE
%
\begin{ex}
The valuation $\Smu$ 
(see Example~\ref{E:smeas-val}) is \emph{not} complete.

It takes some effort to prove this.
Let $A_1,A_2,\dotsc$ be a
disjoint sequence of distinct open non-empty intervals in $[0,1]$,
and let $A\eqdf \bigcup_{n} A_n$.
For instance, if one chooses
\begin{equation*}
A_1 = (\nicefrac{1}{3},\nicefrac{2}{3}),\qquad
A_2 = (\nicefrac{1}{9},\nicefrac{2}{9}),\qquad
A_3 = (\nicefrac{7}{9},\nicefrac{8}{9}),\qquad\dotsb
\end{equation*}
as in the definition of the Cantor set,
one sees that~$A$ will be
the  complement of the Cantor set in~$[0,1]$,
which is called  the \emph{Cantor Discontinuoum}.

Now, define $B_N \eqdf A_1 \cup \dotsb \cup A_N$
for each~$N\in \N$.
We will prove that
\begin{equation*}
B_1 \ \subseteq \ B_2 \ \subseteq \ B_3 \ \subseteq \ \dotsb
\end{equation*}
is $\Smu$-convergent.
That is,
we must show that $\bv_N \Smu (B_N)$ exists.
Note that we have
$B_N \subseteq [0,1]$ for all~$N\in\N$.
So the sequence $\Smu(B_1) \leq \Smu(B_2) \leq \dotsb$ in~$\R$
is bounded from above by~$\Smu([0,1])$.
Hence $\bv_N \Smu(B_N)$ exists.

Note that if~$\Smu$ is complete,
then $B_1 \subseteq B_2 \subseteq \dotsb$
has a supremum in~$\SA$
as $B_1 \subseteq B_2 \subseteq \dotsb$
is $\Smu$-convergent.
So to prove that $\Smu$
is not complete,
we will show that 
$B_1 \subseteq B_2 \subseteq \dotsb$
has no supremum in~$\SA$.

Let~$U\in\SA$ be an upper bound of~$B_1\subseteq B_2 \subseteq \dotsb$
in~$\SA$. We will find an upper bound~$U'\in\SA$
of $B_1\subseteq B_2 \subseteq\dotsb$
such that $U'\subset U$.
Then~$U$ is not the smallest upper bound of
$B_1 \subseteq B_2 \subseteq\dotsb$.
Hence there is no supremum of~$B_1 \subseteq B_2 \subseteq\dotsb$.

Since $U\in\SA$,
we can write $U\equiv I_1 \cup \dotsb \cup I_N$
where $I_1,\dotsc,I_N$ are non-empty disjoint intervals.
We can choose $I_1,\dotsc,I_N$
in such a way that $I_1,\dotsc,I_N$ are precisely
the connected components of~$A$.
Let $k\in\N$ be given.
Note that 
\begin{equation*}
A_k\ \subseteq \ B_k \ \subseteq \ U \ =\ 
I_1 \,\cup\,\dotsb\,\cup\,I_N.
\end{equation*}
So since~$A_k$ is connected and non-empty,
there is a unique~$n\in \N$ such that $A_k \subseteq I_n$.

Now, let $f\colon \N \ra \{ 1, \dotsc, N \}$
be given by $A_k \subseteq I_{f(k)}$ for all~$k\in \N$.
Then by the pidgeon hole principle
there is an~$n\in\N$ such that $\#\{k\in\N\colon f(k)=n\} \geq 2$.

Let $k_1,\dotsc,k_M$
be an enumeration of all~$k\in\N$ with $A_k\subseteq I_n$.
Then 
\begin{equation*}
A_{k_1} \,\cup\, \dotsb \,\cup\, A_{k_M} \ \subseteq \ I_n.
\end{equation*}
Since~$I_n$ is connected, and  $A_{k_1},\dotsc,A_{k_M}$
is disjoint,
and each $A_{k_m}$ is open and non-empty,
there is an~$x\in I_n \,\backslash\  A_{k_1}\cup\dotsb\cup A_{k_M}$.
Now, define
\begin{equation*}
I^a_n \ \eqdf\  I_n \,\cap\, [0,x)
\qquad\text{and}\qquad
I^b_n \ \eqdf\  I_n \,\cap\, (x,1].
\end{equation*}
Then $I^a_n$ and $I^b_n$ are intervals
and we have
\begin{equation*}
A_{k_1} \,\cup\,\dotsb\,\cup\, A_{k_M}
\ \subseteq\ 
I^a_n \,\cup\, I^b_n
\ \subset \ I_n
\end{equation*}
So let us define~$U'\in\SA$ by 
\begin{equation*}
U'\ \eqdf\ 
I_1 \,\cup\,\dotsb \,\cup\, I_{n-1} \,\cup\,
I_{n+1} \,\cup\, \dotsb \,\cup\, I_N\,\cup\,I^a_n\,\cup\,I^b_n.
\end{equation*}
Then we have $A_k \subseteq U'$ for all~$k$,
and $U'\subset U$.
So we see that $U'$ is an upper bound for $B_1\subseteq B_2 \subseteq\dotsb$
that is strictly smaller than~$U$.
Hence $\Smu$ is not complete.
\end{ex}
%
%                  THE STEP FUNCTIONS ARE NOT COMPLETE
%
\begin{ex}
The valuation $\Sphi$ 
(see Example~\ref{E:sint-val}) is also \emph{not} complete.\\
We leave it to the reader to prove this fact.
\end{ex}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\subsection{$R$-completeness}
\label{SS:complete-val_R-completeness}
Let $\varphi\colon L \ra E$ be a valuation.
Let $a_1 \leq a_2 \leq \dotsb$
and $b_1 \leq b_2 \leq \dotsb$ be $\varphi$-convergent sequences in~$L$.
(see Definition~\ref{D:phi-conv}).

For the development of the theory,
it would be convenient if also 
\begin{equation}
\label{eq:SS-complete-val_R-completeness-1}
a_1 \vee b_1 \ \leq\  a_2 \vee b_2 \ \leq\  \dotsb
\qquad\text{is $\varphi$-convergent.}
\end{equation}
Unfortunately, 
this is not always the case (see Example~\ref{E:P:R-main}).
However,
if~$E$ is Dedekind complete 
(such as~$\R$),
then one can prove that Statement~\eqref{eq:SS-complete-val_R-completeness-1}
holds.

In fact,
if we assume that
$E$ is only \emph{$R$-complete} (see Definition~\ref{D:R-complete}) ---
which is a weaker assumption than that~$E$ is Dedekind-complete ---
then we can still prove that 
that Statement~\eqref{eq:SS-complete-val_R-completeness-1}
holds (see Proposition~\ref{P:R-main}).

We study the notion of $R$-completeness in this subsection.
%
%                  R-completeness
%
\begin{dfn}
\label{D:R-complete}
Let $E$ be an ordered Abelian group.
Consider the following.
\begin{equation*}
\left[\quad 
\begin{minipage}{.7\columnwidth}
Let $x_1 \leq x_2 \leq \dotsb$
and $y_1 \leq y_2 \leq \dotsb$ be from~$E$
such that
\begin{equation*}
x_{n+1} - x_n \ \leq\ y_{n+1} - y_n\qquad \text{for all }n.
\end{equation*}
Then $\bv x_n $ exists whenever $\bv y_n$ exists.
\end{minipage}
\right.
\end{equation*}
If the above statement holds,
we say~$E$ is \keyword{$R$-complete}.
\end{dfn}

\begin{exs}
\begin{enumerate}
\item
Any $\sigma$-Dedekind complete Riesz space is $R$-complete.

\item
In particular, $\R$ is $R$-complete.

\item
The lexicographic plane~$\Lex$ (see Example~\ref{E:oag_lex})
is $R$-complete,
but~$\Lex$ is not $\sigma$-Dedekind complete.

\item
The ordered Abelian group~$\Q$ is \emph{not} $R$-complete.\\
To see this,
and pick $q_1 \leq q_2\leq \dotsb$ in~$\Q$
with
\begin{equation*}
q_{n+1} - q_{n} \ \leq\ 2^{-(n+1)}
\qquad\quad\text{and}\quad\qquad  \text{$\bv_n q_n = \sqrt{2}$ in~$\R$.}
\end{equation*}
Note that $q_1\leq q_2 \leq \dotsb$
has no supremum in~$\Q$.

Now, let $y_n \eqdf 1-2^{-n}$ for all~$n\in\N$.
Then $y_1 \leq y_2 \leq \dotsb$ has an supremum, namely~$1$,
and we have $y_{n+1} - y_{n} = 2^{-(n+1)}$. So we see that
\begin{equation*}
q_{n+1} - q_{n} \ \leq\ y_{n+1} - y_n
\qquad\quad(n\in\N).
\end{equation*}
If~$\Q$ would be~$\R$-complete,
then the above implies 
 $q_1 \leq q_2 \leq \dotsb$
has a supremum in~$\Q$,
which it does not.
Hence~$\Q$ is not $R$-complete.
\end{enumerate}
\end{exs}

\begin{rem}
Let~$E$ be an ordered Abelian group.
Using the map $x\mapsto -x$,
one can easily verify
that $E$ is $R$-complete
if and only if the following statement holds.
\begin{equation*}
\left[\quad 
\begin{minipage}{.7\columnwidth}
Let $x_1 \geq x_2 \geq \dotsb$
and $y_1 \geq y_2 \geq \dotsb$ be from~$E$
such that
\begin{equation*}
x_{n} - x_{n+1}\ \leq\ y_{n} - y_{n+1}\qquad \text{for all }n.
\end{equation*}
Then $\bw y_n $ exists implies that $\bw x_n$ exists.
\end{minipage}
\right.
\end{equation*}
\end{rem}

\begin{prop}
\label{P:R-main}
Let  $E$ be an ordered Abelian group which is $R$-complete.
Let~$L$ be a lattice and $\varphi\colon L \ra E$ a valuation.
\begin{enumerate}
\item
\label{P:R-main-descending}
If  $a_1 \geq a_2 \geq \dotsb$,
$b_1 \geq b_2 \geq \dotsb$
are  $\varphi$-convergent
sequences from~$L$,
then
\begin{equation*}
a_1 \wedge b_1 \,\geq\, a_2 \wedge b_2 \,\geq\, \dotsb
\qquad\text{and}\qquad
a_1 \vee b_1 \,\geq\, a_2 \vee b_2 \,\geq\, \dotsb
\end{equation*}
are $\varphi$-convergent.

\item
\label{P:R-main-ascending}
If  $a_1 \leq a_2 \leq \dotsb$,
$b_1 \leq b_2 \leq \dotsb$
are  $\varphi$-convergent
sequences from~$L$,
then
\begin{equation*}
a_1 \wedge b_1 \,\leq\, a_2 \wedge b_2 \,\leq\, \dotsb
\qquad\text{and}\qquad
a_1 \vee b_1 \,\leq\, a_2 \vee b_2 \,\leq\, \dotsb
\end{equation*}
are $\varphi$-convergent.
\end{enumerate}
\end{prop}
\begin{proof}
\ref{P:R-main-descending}\ 
We prove that $a_1 \wedge b_1 \geq a_2 \wedge b_2 \geq\dotsb$
is $\varphi$-convergent.
For this we need to show that $\bw_n \varphi(a_n\wedge b_n)$ exists.
Note that since $\bw_n \varphi (a_n)$
and $\bw_n \varphi(b_n)$ exist,
we know that $\bw_n \ (\varphi(a_n) + \varphi(b_n))$
exists (see Lemma~??).
So by $R$-completeness,
in order to show $\bw_n\varphi(a_n \wedge b_n)$ exists,
it suffices to prove that 
\begin{equation*}
\varphi(a_{n}\wedge b_{n}) \,-\, \varphi(a_{n+1} \wedge b_{n+1}) 
\ \leq\ 
(\,\varphi(a_{n}) + \varphi(b_{n})\,) 
\,-\, (\,\varphi(a_{n+1}) + \varphi(b_{n+1})\,).
\end{equation*}
Phrased differently
using the distance ``$d_\varphi$'',
we need to prove that
\begin{equation*}
d_\varphi(a_{n}\wedge b_{n},\, a_{n+1} \wedge b_{n+1}) 
\ \leq\ 
d_\varphi(a_{n},a_{n+1}) + d_\varphi(b_{n},b_{n+1}).
\end{equation*}
This follows immediately from Lemma~\ref{L:wv-unif}.

The 
proof that~$a_1 \vee b_1 \geq a_2 \vee b_2 \geq \dotsb$
is $\varphi$-convergent is similar.

\ref{P:R-main-ascending}.  Again, similar.
\end{proof}
%
%                  COUNTEREXAMPLE ON PROP R-MAIn
%
\begin{ex}
\label{E:P:R-main}
We will prove that Proposition~\ref{P:R-main}
does not hold if we drop the assumption that~$E$ is $R$-complete.

Let $\mathcal{A}$ be the ring of subsets (see Example~\ref{E:ring-val})
of~$\R$ generated by the non-empty closed intervals with rational endpoints,
i.e., subsets of the form $[q,r]$ where $q,r\in \Q$ and $q<r$.
Then there is a unique positive and additive map $\mu\colon\mathcal{A} \ra\Q$
such that
\begin{equation*}
\mu(\,[q,r]\,) \ =\ r-q\qquad\quad\text{for all $q<r$ from $\Q$}.
\end{equation*}

Recall that~$\Q$ is not $R$-complete.
To prove that the conclusion of Proposition~\ref{P:R-main}
does not hold for~$E=\Q$, we will find $\mu$-convergent sequences $A_1 \subseteq A_2 \subseteq$
and $B_1 \subseteq B_2 \subseteq\dotsb$ such that
$A_1 \cup B_1 \ \subseteq\ A_2 \cup B_2 \ \subseteq\ \dotsb$
is not $\mu$-convergent.

Find rational numbers $ \dotsb \leq r_2 \leq r_1 < q_1 \leq q_2 \leq \dotsb$
such that, in~$\R$,
\begin{equation*}
\bv_n q_n \ = \ \sqrt2\qquad\text{and}\qquad
\bw_n r_n \ = \ \sqrt2-1.
\end{equation*}
Now, let us define $A_1 \subseteq A_2 \subseteq \dotsb$
and $B_1 \subseteq B_2 \subseteq \dotsb$ in~$\mathcal{A}$
by, for~$n\in\N$,
\begin{equation*}
A_n  \ = \ [0, r_1]
\qquad\text{and}\qquad
B_n \ = \ [r_n,q_n].
\end{equation*}
Then clearly $A_1 \subseteq A_2 \subseteq \dotsb$
is $\varphi$-convergent.
Note that $\mu(B_n) = q_n - r_n$. So, in~$\R$,
\begin{equation*}
\bv_n \mu(B_n) \ =\ \bv_n q_n - \bw_n r_n \ =\ 1.
\end{equation*}
Hence $\bv_n \mu(B_n)=1$ in~$\Q$.
So  $B_1 \subseteq B_2 \subseteq \dotsb$
is a $\mu$-convergent sequence.

However $A_n \cup B_n = [0,q_n]$,
and thus $\mu(A_n \cup B_n) = q_n$.
So we see that, in~$\R$,
\begin{equation*}
\bv_n \mu(\,A_n \cup B_n \,) \ =\ \bv_n q_n \ =\ \sqrt{2}.
\end{equation*}
So $\mu(A_1 \cup B_1) \ \leq\  \mu(A_2 \cup B_2) \ \leq\ \dotsb$
has no supremum in~$\Q$.

Hence $A_1 \cup B_2 \ \subseteq\ A_2 \cup B_2 \ \subseteq\ \dotsb$
is not $\mu$-convergent.




\end{ex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Convergence Theorems}
\label{SS:complete-val_convergence}
%
%                  CONVERGENCE OF SEQUENCES
%
The notion of a complete valuation
has been based on Levi's Monotone Convergence Theorem
(see Example~\ref{E:complete-lint}).
In this subsection,
we prove variants of some of the other classical convergence theorems
of integration theory.
For example, 
Lebesgue's Dominated Convergence Theorem.
It states:
\begin{equation}
\label{eq:Lebesgue}
\left[\quad
\begin{minipage}{.7\columnwidth}
Let $f_1,\,f_2,\,\dotsc$ be a sequence in~$\LF$.

Assume $f_1(x),\,f_2(x),\,\dotsc$
converges for almost all~$x\in\R$.

Assume that $f_1,\,f_2,\,\dotsc$
 is dominated in the sense that
 $|f_n|\leq D$ for all~$n$
for some $D\in\LF$.

Then there is an $f\in\LF$
with $f_1(x),\,f_2(x),\,\dotsc$ converges to $f(x)$ for 
for almost all~$x\in\R$,
and $\Lphi(f) = \lim_n\Lphi(f_n)$.
\end{minipage}
\right.
\end{equation} 
The difficulty 
in the setting of  valuation systems
is not the proof of the theorem,
but its formulation.
For instance, it 
not clear how we should interpret 
\begin{equation*}
\text{``$f_1(x),\,f_2(x),\,\dotsb$
converges for almost all~$x$'',}
\end{equation*}
when the objects $f_n$ are not necessarily functions,
but elements of a lattice~$V$.

%
%                  CONVERGENCE IN A LATTICE
%
Let us begin by generalising the notion of convergence in~$\R$
to any lattice~$L$.
Recall that a sequence $a_1,a_2,\dotsc$ in~$\R$
is convergent (in the usual sense) if and only if 
the \emph{limit inferior}, $\lim_{N} \inf_{n\geq N}\,a_n$,
and the \emph{limit superior}, $\lim_{N} \sup_{n\geq N}\,a_n$,
exist and are equal.
This leads us to the following definitions.
\begin{dfn}
\label{D:conv}
Let~$L$ be a lattice.
Let $a_1,\,a_2,\,\dotsc$ be a sequence in~$L$.
\begin{enumerate}
\item
We say $a_1,\,a_2,\,\dotsc$ 
is \keyword{upper convergent}
if the following exists.
\begin{equation*}
\ulim{n} a_n \ \eqdf\ \bw_N\bv_{n\geq N}\ a_N \vee \dotsb \vee a_n.
\end{equation*}
Similarly,
we say $a_1,\,a_2,\,\dotsc$ is \keyword{lower convergent}
if the following exists.
\begin{equation*}
\llim{n} a_n \ \eqdf\ \bv_N\bw_{n\geq N}\ a_N \wedge \dotsb \wedge a_n.
\end{equation*}

\item
We say $a_1,\,a_2,\,\dotsc$ is \keyword{convergent}
if it is both upper and lower convergent,
and in addition $\ulim{n}a_n = \llim{n}a_n$.
In that case,
we write $\lim_n a_n\eqdf\ulim{n} a_n$.

\item
Let $a\in L$ be given.
We say $a_1,\,a_2,\,\dotsc$ 
\keyword{converges to}~$a$
if  $a=\lim_n a_n$.
\end{enumerate}
\end{dfn}
%
%                  REMARK ON THE INEQUALITY OF LIMSUP AND LIMINF
%
\begin{rem}
\label{R:conv}
Let~$L$ be a lattice.
Let $a_1,\,a_2,\,\dotsc$ be a sequence in~$L$,
which is  upper convergent and lower convergent.
Then we have the following inequality.
\begin{equation*}
\llim{n} a_n \ \leq\ \ulim{n} a_n
\end{equation*}
Indeed,
this follows immediately from the observation
that, for every~$N\in\N$,
\begin{equation*}
\bv_{n\geq N}\ a_N \wedge \dotsb \wedge a_n
\ \leq\ 
\bw_{n\geq N}\ a_N \vee \dotsb \vee a_n.
\end{equation*}
\end{rem}
%
%                  EXAMPLES OF CONVERGENCE IN A LATTICE
%
\begin{exs}
\label{E:conv}
\begin{enumerate}
\item 
In~$\R$ we have:
A sequence $a_1,\,a_2,\,\dotsc$
is convergent in the sense of Definition~\ref{D:conv}
if and only if $a_1,\,a_2,\,\dotsc$
is convergent as usual.\\
Moreover, if $a_1,\,a_2.\,\dotsc$
is convergent, then $\lim_n a_n$ from Definition~\ref{D:conv}
is also the limit of $a_1,\,a_2,\,\dotsc$
in the usual sense.

\item
Similarly,
in~$\R^X$, where~$X$ is any set,
``convergent'' from Definition~\ref{D:conv}
coincides with the usual ``pointwise convergent''.

\item
It is not hard to see
(cf. Example~\ref{E:int-complete-val}),
that in the space~$\LF$ of Lebesgue integrable functions,
``convergent'' from Definition~\ref{D:conv}
coincides with the usual ``pointwise convergent''.
\end{enumerate}
\end{exs}

%
%                  PHI-CONVERGENCE
%
Let us now think about ``almost everywhere convergent''.
Let $\varphi\colon L \ra E$ be a valuation.
We would like to define a notion of convergence
for sequences in~$L$ 
which coincides with ``convergence
almost everywhere'' in the case of the  Lebesgue integral,
that is, for $\varphi = \Lphi$
(see Example~\ref{E:int-val}).

Unfortunately,
we have not found such a notion of convergence.
Instead,
we study a specific notion of convergence,
which we call \emph{$\varphi$-convergence}
(see Definition~\ref{D:seq-phi-conv})
that \emph{almost} coincides with ``convergence almost everywhere''.

More precisely,
given Lebesgue integrable functions $f_1,\,f_2,\,\dotsc$
and $f$, we have that the
$f_1,\,f_2,\,\dotsc$ $\Lphi$-converges to~$f$
if and only if 
\begin{equation*}
\left[\quad
\begin{minipage}{.7\columnwidth}
 $f_1(x),\,f_2(x),\,\dotsc$ 
converges to~$f(x)$ for allmost all~$x\in \R$,
\emph{and}
there is a Lebesgue integrable function~$D$
such that 
\begin{equation*}
|f_n|\,\leq\, D\qquad\quad(n\in\N).
\end{equation*}
\end{minipage}
\right.
\end{equation*}
We postpone the proof of this 
statement to Example~\ref{E:lebesgue}
as it requires some theory.
\begin{dfn}
\label{D:seq-phi-conv}
Let $E$ be an ordered Abelian group.
Let~$L$ be a lattice.\\
Let $\varphi\colon L \ra E$ be a valuation.
Let $a_1,\, a_2\, \dotsc\in L$ be given.
\begin{enumerate}
\item
We say $a_1,\,a_2,\, \dotsc$
is \keyword{upper $\varphi$-convergent}
if the following exists.
\begin{equation*}
\pulim\varphi{n}a_n \ \eqdf\ 
\bw_N \bv_{n\geq N}\ \varphi(a_N\vee\dotsb\vee a_n)
\end{equation*}
Similarly,
we say $a_1,\,a_2,\,\dotsc$ is \keyword{lower $\varphi$-convergent} if
the following exists.
\begin{equation*}
\pllim\varphi{n}a_n \ \eqdf\ 
\bv_N \bw_{n\geq N}\ \varphi(a_N\wedge\dotsb\wedge a_n)
\end{equation*}

\item
We say $a_1,\,a_2,\,\dotsc$
is \keyword{$\varphi$-convergent}
if it is lower and upper $\varphi$-convergent,
and in addition $\pulim\varphi{n}a_n = \pllim\varphi{n}a_n$.
\end{enumerate}
\end{dfn}
%
%                  REMARK ON INEQUALITY BETWEEN UPPER AND LOWER PHI-LIM
%
\begin{rem}
\label{R:seq-phi-conv}
Let $E$ be an ordered Abelian group.
Let~$L$ be a lattice.\\
Let $\varphi\colon L \ra E$ be a valuation.
Let $a_1,\,a_2,\,\dotsb$ be a sequence in~$L$,
which is upper and lower $\varphi$-convergent.
We have the following inequality
(cf. Remark~\ref{R:conv}).
\begin{equation*}
\pllim\varphi{n}a_n \ \leq\ \pulim\varphi{n}a_n.
\end{equation*}
\end{rem}
%
%                  EXAMPLE OF PHI-CONV SEQUENCE
%
\begin{ex}
\label{E:seq-phi-conv}
Let $f_1,f_2,\dotsc$ be  Lebesgue integrable functions
(see Example~\ref{E:int-val}). \\
We will prove the following.
\begin{equation}
\label{eq:E:seq-phi-conv}
\left[\quad
\begin{minipage}{.7\columnwidth}
 The sequence $f_1,f_2,\dotsc$
is  upper and lower $\varphi$-convergent.
\begin{center}
$\Longleftrightarrow$
\end{center}
There is a Lebesgue integrable~$D$ with $|f_n| \leq D$ for all~$n$.
\end{minipage}
\right.
\end{equation}

\vspace{.3em}
\emph{($\Longrightarrow$)}
Assume that  $f_1,\,f_2,\,\dotsc$ is upper and lower $\Lphi$-convergent.
We prove there is a Lebesgue integrable function~$D$ such that
$|f_n|\leq D$ for all~$n\in \N$.

Since $f_1,\,f_2,\,\dotsc$ is  upper $\Lphi$-convergent
(see Definition~\ref{D:seq-phi-conv}), 
we know that 
\begin{equation*}
\bv_n \ \Lphi(f_1 \vee \dotsb \vee f_n)\qquad\text{exists.}
\end{equation*}
In other words, we know that
\begin{equation*}
f_1 \ \leq\ f_1 \vee f_2 \ \leq\ \dotsb \qquad\text{is $\Lphi$-convergent.}
\end{equation*}
Since $\Lphi$ is complete
(see Example~\ref{E:int-complete-val}),
this implies that $\ol{D}\eqdf \bv_n f_n$ exists in~$\LF$.
Note that $f_n \leq \ol{D}$ for all~$n\in \N$.

By a similar reasoning, but using the fact that
$f_1,\,f_2,\,\dotsc$ is lower $\varphi$-convergent,
we can find a Lebesgue integrable function~$\underline{D}$
such that $\ul{D}\leq f_n$ for all~$n\in \N$.

Now define $D\eqdf  \ol{D} \vee (-\ul{D})$. We see that $|f_n|\leq D$
for all~$n\in \N$.

\vspace{.3em}

\emph{($\Longleftarrow$)}
Assume
 that $|f_n|\leq D$
for some Lebesgue integrable~$D$.

We prove that $f_1,\,f_2,\,\dotsc$
is upper $\Lphi$-convergent.
For this,
we must show that the following exists
(see Definition~\ref{D:seq-phi-conv}).
\begin{equation}
\label{eq:E:seq-phi-conv-1}
\bw_N \bv_{n\geq N}\ \Lphi(f_N\vee\dotsb\vee f_n)
\end{equation}
Let $N\in\N$ and $n\geq N$ be given.
Note that we have 
\begin{equation*}
f_N \vee \dotsb\vee f_n \ \leq\ D.
\end{equation*}
Since $\Lphi$ is order preserving, this gives us
\begin{equation*}
\Lphi(f_N \vee \dotsb \vee f_n)
\ \leq\  \Lphi(D).
\end{equation*}
So we see that 
$\Lphi(f_N) \ \leq\ 
\Lphi(f_N \vee f_{N+1}) \ \leq\ \dotsb$
is an ascending sequence in~$\R$
bounded from above by~$\Lphi(D)$.
Hence $\bv_{n\geq N} \, \Lphi(f_N\vee \dotsb \vee f_n)$ exists.
I.e.,
\begin{equation*}
f_N \ \leq \ f_N \vee f_{N+1} \ \leq\ \dotsb
\qquad\text{is $\Lphi$-convergent.}
\end{equation*}
Since $\Lphi$ is complete
(see Example~\ref{E:int-complete-val}),
this implies $\ol{f}_N \eqdf \bv_{n\geq N}\,f_n$ exists and
\begin{equation*}
\Lphi(\ol{f}_N)\ 
=\ \bv_{n\geq N}\ \Lphi(f_N \vee \dotsb \vee f_n).
\end{equation*}
Recall that we must prove that Expression~\eqref{eq:E:seq-phi-conv-1}
exists.
In other words,
we must show that $\bw_N \Lphi(\ol{f}_N)$ exists.
Note that $-D \leq  f_n$ for all~$n\in\N$.
Hence 
\begin{equation*}
-D \ \leq \   \ol{f}_N  = \bv_{n \geq N}\, f_n .
\end{equation*}
We now have a descending sequence
$\Lphi(\ol{f}_1) \,\geq\, \Lphi(\ol{f}_2) \,\geq\,\dotsb$
in~$\R$ that is bounded from below by $\Lphi(-D)$.
Hence $\bw_N \Lphi(\ol{f}_N)$ exists.

We have proven that $f_1,\,f_2,\,\dotsc$ 
is upper $\varphi$-convergent.
With a similar reasoning one can prove that
$f_1,\,f_2,\,\dotsc$
is lower $\varphi$-convergent.

This completes the proof of Statement~\eqref{eq:E:seq-phi-conv}.

\vspace{.3em}

Again, let $f_1,\,f_2,\,\dotsc$ be a sequence of
Lebesgue integrable functions.
Assume that $f_1,\,f_2,\,\dotsc$
is upper and lower $\Lphi$-convergent.
Note that the sequence $f_1,\,f_2,\,\dotsc$
need not be $\Lphi$-convergent.
That is,
we do not always have
 $\pulim\Lphi{n} f_n = \pllim\Lphi{n} f_n$.
Indeed, consider $f_1,\,f_2,\,\dotsc$
given by
\begin{equation*}
f_n \ =\  (-1)^{n}\cdot \mathbf{1}_{[0,1]}.
\end{equation*}
Then we have the following inequality
\begin{equation*}
\pllim\Lphi{n} f_n \ = \ -1 \ <\  +1 \ = \ 
\pulim\Lphi{n} f_n.
\end{equation*}

Indeed,
as has been remarked before,
we will lateron prove 
(see Example~\ref{E:lebesgue}) that such a sequence $f_1,\,f_2,\,\dotsc$
is $\varphi$-convergent if and only if 
\begin{equation*}
\text{ $f_n(x)$ converges for almost all~$x\in\R$.}
\end{equation*}
\end{ex}

%
%                  LEMMA OF FATOU
%
\noindent The following lemma
is a generalisation of the Lemma of Fatou.
\begin{lem}
\label{L:fatou}
Let $E$ be an ordered Abelian group.\\
Let $L$ be a lattice,
and let $\varphi\colon L \ra E$ be a 
complete valuation.\\
Let $a_1,a_2,\dotsc$ be an upper $\varphi$-convergent
sequence in~$L$
(see Definition~\ref{D:seq-phi-conv}) \\
Then $a_1,a_2,\dotsc$ is upper convergent
(see Definition~\ref{D:conv}),
and we have
\begin{equation*}
\varphi(\ulim{n}a_n) \ =\ 
\pulim\varphi{n} a_n.
\end{equation*}
Moreover,
if $E$ is a lattice, and
if $\ulim{n}\varphi(a_n)$ exists
(see Definition~\ref{D:conv}), then
\begin{equation*}
\pulim\varphi{n}a_n
\ \geq\  
\ulim{n}\varphi(a_n) .
\end{equation*}
\end{lem}
\begin{proof}
Let $a_1,\,a_2,\,\dotsc$ be a upper $\varphi$-convergent sequence.
We prove that $a_1,a_2,\dotsc$ is upper convergent
(see Definition~\ref{D:conv}),
and that $\varphi(\ulim{n}a_n) = \pulim\varphi{n}a_n$.

Let $N\in\N$ be given.
Note that $\bv_{N\geq n} \,\varphi(a_N\vee\dotsb\vee a_n)$
exists because the sequence~$a_1,\,a_2,\,\dotsc$ is upper $\varphi$-convergent
(see Definition~\ref{D:seq-phi-conv}).
So the sequence
\begin{equation*}
a_N \,\leq\ a_N \vee a_{N+1} 
    \ \leq\quad a_N\vee a_{N+1} \vee a_{N+2} 
    \quad \leq\qquad \dotsb
\end{equation*}
is $\varphi$-convergent (in the sense of
Definition~\ref{D:phi-conv}).
For brevity,
let us write
\begin{equation*}
\overline{a}_{N}^n \ \eqdf \ a_N\vee\dotsb\vee a_{N+n},
\end{equation*}
Since~$\varphi$ is complete,
and $\overline{a}_N^0 \leq \overline{a}_N^1 \leq\dotsb$
is $\varphi$-convergent,
we get $\bv_n\,\overline{a}_N^n$ exists,
and 
\begin{equation*}
\varphi(\overline{a}_N) 
\ =\ 
 \bv_{n}\ \varphi( \overline{a}_N^n).
\end{equation*}
For brevity, let us write
\begin{equation*}
\overline{a}_N \ \eqdf \ \bv_n \,\overline{a}_N^n.
\end{equation*}

Note that $\overline{a}_1 \geq \overline{a}_2 \geq\dotsb$
is $\varphi$-convergent,
because
\begin{equation*}
\pulim\varphi{n} a_n
\ \eqdf\ 
\bw_N\bv_n\, \varphi(\overline{a}^n_N)
\end{equation*}
exists as $a_1,\,a_2,\,\dotsc$ is upper $\varphi$-convergent.
Since $\varphi$ is complete,
this implies that
\begin{equation*}
\bw_n \overline{a}_N\text{ exists}
\qquad\text{and}\qquad \varphi(\bw_n \overline{a}_N)
\,=\,
\bw_n\varphi(\overline{a}_N).
\end{equation*}
Now,
note that  we have the following equality.
\begin{equation*}
\bw_N \overline{a}_N \ =\ 
\bw_N \bv_{n\geq N} \, a_n
\end{equation*}
So we see that $a_1,a_2,\dotsc$
is $\varphi$-convergent and that
\begin{equation*}
\varphi(\ulim{n}a_n)
\ =\ 
\bw_N\varphi(\overline{a}_N)
\ =\ 
\bw_N \bv_n \varphi(\overline{a}^n_N)
\ =\ 
\pulim\varphi{n} a_n.
\end{equation*}
We have proven the first part of the lemma.

Assume $E$ is a lattice and $\ulim{n}\varphi(a_n)$ exists
(see Definition~\ref{D:conv}).
To prove the remainder of the theorem,
we need to show that 
$\pulim\varphi{n}a_n \geq \ulim{n}\varphi(a_n)$.
That is,
\begin{equation*}
\bw_N \bv_{n\geq N} \ \varphi(a_N \vee \dotsb \vee a_n)
\ \geq \ 
\bw_N \bv_{n\geq N} \ \varphi(a_N) \vee \dotsb \vee \varphi(a_n).
\end{equation*}
This is easy.  It follows immediately
from the fact that
\begin{equation*}
\varphi(a_N\vee \dotsb\vee a_n)
\ \geq\ \varphi(a_N)\vee \dotsb \vee \varphi(a_n)
\end{equation*}
for all~$N\in\N$ and $n\geq N$.
\end{proof}

%
%                  LEBESGUE'S DOMINATED CONVERGENCE THEOREM
%
\noindent
The following theorem contains the gist
of Lebesgue's Dominated Convergence Theorem,
although there is no word of domination.
\todo{Add remark on
$E$ which are conditionally complete.}
\begin{thm}
\label{T:lebesgue}
Let $E$ be an ordered Abelian group.\\
Let $L$ be a lattice,
and let $\varphi\colon L \ra E$ be a 
complete valuation.\\
Let $a_1,a_2,\dotsc$ be a $\varphi$-convergent
sequence in~$L$
(see Definition~\ref{D:seq-phi-conv}). \\
Then $a_1,a_2,\dotsc$ is convergent
(see Definition~\ref{D:conv}),
and we have
\begin{equation*}
\varphi(\llim{n}a_n) \,=\, \varphi(\ulim{n} a_n) 
\ =\ 
\plim\varphi{n} a_n.
\end{equation*}
Moreover,
if $E$ is a lattice,
and if $\llim{n}\varphi(a_n)$ and $\ulim{n}\varphi(a_n)$ exists, 
then
\begin{equation*}
\plim\varphi{n}a_n \ =\ {\lim}_n \varphi(a_n).
\end{equation*}
\end{thm}
\begin{proof}
Let $a_1,a_2,\dotsc$ be a $\varphi$-convergent sequence.
The first part of the theorem follows immediately from Lemma~\ref{L:fatou}
and its dual.

Assume that $\llim{n}\varphi(a_n)$ and $\ulim{n}\varphi(a_n)$ exist.
By Lemma~\ref{L:fatou}, we see that
\begin{equation*}
\pllim\varphi{n}a_n \,\leq\,
\llim{n}\varphi(a_n) \,\leq\,
\ulim{n}\varphi(a_n) \,\leq\,
\pulim\varphi{n}a_n.
\end{equation*}
But $\pllim\varphi{n}a_n = \pulim\varphi{n}a_n$,
since $a_1,a_2,\dotsc$ is $\varphi$-convergent.
So we get 
\begin{equation*}
\pllim\varphi{n}a_n \,=\,
\llim{n}\varphi(a_n) \,=\,
\ulim{n}\varphi(a_n) \,=\,
\pulim\varphi{n}a_n.
\end{equation*}
In particular,
$\varphi(a_1),\,\varphi(a_2),\,\dotsc$
is convergent and ${\lim}_n\varphi(a_n) = \plim\varphi{n}a_n$.
\end{proof}
%
%                  EXAMPLE ON LEBESGUE'S THEOREM
%
\begin{ex}
\label{E:lebesgue}
Let us now prove Lebesgue's Dominated Convergence Theorem,
 see Statement~\eqref{eq:Lebesgue}, 
using Theorem~\ref{T:lebesgue}.
To do this, we first establish the relationship between
``$\Lphi$-convergent'' and ``convergent almost everywhere''.

Let $f_1,\,f_2,\,\dotsc$ be a sequence of Lebesgue integrable functions.
We will prove:
\begin{equation}
\label{eq:E:lebesgue-0}
\left[\quad
\begin{minipage}{.7\columnwidth}
The sequence $f_1,\,f_2,\,\dotsc$
is $\Lphi$-convergent
\begin{center}
if and only if
\end{center}
 $f_1(x),\,f_2(x),\,\dotsc$ converges
for allmost all~$x\in \R$,
and
there is a Lebesgue integrable function~$D$
such that 
\begin{equation*}
|f_n|\,\leq\, D\qquad\quad(n\in\N).
\end{equation*}
\end{minipage}
\right.
\end{equation}

\vspace{.3em}
\emph{($\Longrightarrow$)}
Suppose that $f_1,\,f_2,\,\dotsc$
is $\Lphi$-convergent.
By Example~\ref{E:seq-phi-conv}
we know that there is a Lebesgue integrable function~$D$
such that $|f_n|\leq D$ for all~$n\in\N$.
So we only need to show that $f_n(x)$ 
converges for almost all~$x\in\R$.

By Theorem~\ref{T:lebesgue},
we know that $f_1,\,f_2,\,\dotsc$
is upper convergent and lower convergent
(see Definition~\ref{D:conv}),
and that 
\begin{equation}
\label{eq:E:lebesgue}
\Lphi(\llim{n} f_n) \ =\  \Lphi(\ulim{n} f_n).
\end{equation}
Since $\llim{n} f_n \leq \ulim{n} f_n$
(see Remark~\ref{R:conv})
one can easily see that Statement~\eqref{eq:E:lebesgue}
implies that 
 $\llim{n} f_n \approx \ulim{n} f_n$
(see Definition~\ref{D:eq}).
Hence (see Example~\ref{E:eq-int})
\begin{equation*}
 (\,\llim{n} f_n\,)(x) \ =\  (\,\ulim{n} f_n\,)(x)
\qquad\text{for almost all~$x\in \R$.}
\end{equation*}
Or in other words
(cf.~Example~\ref{E:conv}) we have
\begin{equation*}
 \llim{n} (\,f_n(x)\,) \ =\  \ulim{n} (\,f_n(x)\,)
\qquad\text{for almost all~$x\in \R$.}
\end{equation*}
So we see that $f_1(x),\,f_2(x),\,\dotsc$
is convergent
for almost all~$x\in \R$.

\vspace{.3em}
\emph{($\Longleftarrow$)}
Suppose that there is a Lebesgue integrable function~$D$
such that $|f_n|\leq D$ for all~$n\in\N$,
and that  $f_1(x),\,f_2(x),\,\dotsc$
is convergent for almost all~$x\in\R$.

We must prove that $f_1,\,f_2,\,\dotsc$
is $\Lphi$-convergent (see Definition~\ref{D:seq-phi-conv}).

By Example~\ref{E:seq-phi-conv}
we know that $f_1,\,f_2,\,\dotsc$ is upper and lower
$\Lphi$-convergent.
So the only thing that we must still show is
that 
\begin{equation}
\label{eq:E:lebesgue-2}
\pulim\Lphi{n} f_n \ =\  \pllim\Lphi{n} f_n.
\end{equation}

By Lemma~\ref{L:fatou}, our generalisation of the Lemma of Fatou,
we know that $f_1,\,f_2,\,\dotsc$
is upper and lower convergent in~$\LF$,
and that 
\begin{equation*}
\Lphi(\,\llim{n} f_n\,) \ =\ 
\pllim\Lphi{n} f_n
\qquad\text{and}\qquad
\Lphi(\,\ulim{n} f_n\,) \ =\ 
\pulim\Lphi{n} f_n.
\end{equation*}
So to prove Equality~\eqref{eq:E:lebesgue-2},
it suffices to show that 
\begin{equation}
\label{eq:E:lebesgue-3}
\Lphi(\,\llim{n} f_n\,) \ =\  \Lphi(\,\ulim{n} f_n\,).
\end{equation}
Recall that $\llim{n}f_n \leq \ulim{n}f_n$
(see Remark~\ref{R:conv}).
So Statement~\ref{eq:E:lebesgue-3}
is equivalent to 
\begin{equation*}
\llim{n} f_n \ \approx\  \ulim{n} f_n,
\end{equation*}
where $\approx$ is from Definition~\ref{D:eq}.
So we must prove that (see Example~\ref{E:eq-int})
\begin{equation*}
(\,\llim{n} f_n\,) (x)\ \approx\  (\,\ulim{n} f_n\,)(x)
\qquad\text{for almost all $x\in\R$}.
\end{equation*}
The reader can verify that this is equivalent to
\begin{equation*}
\llim{n} (\, f_n (x)\,)\ \approx\  \ulim{n} (\,f_n(x)\,)
\qquad\text{for almost all $x\in\R$}.
\end{equation*}
That is (see Example~\ref{E:conv}),
we must prove that $f_1(x),\,f_2(x),\,\dotsc$
converges for almost all~$x\in\R$.
But this is exactly what we have assumed.

This concludes the proof of Statement~\eqref{eq:E:lebesgue-0}.

\vspace{.3em}
\noindent
We now prove Lebesgue's Dominated Convergence Theorem
(see Stat.~\eqref{eq:Lebesgue}).

Let $f_1,\,f_2,\,\dotsc$ be a sequence of Lebesgue integrable functions.
Assume there is an Lebesgue integrable function~$D$ such that
$|f_n|\leq D$ for all~$n\in\N$,
and assume that $f_1(x),\,f_2(x),\,\dotsc$ converges
for almost all~$x\in\R$.

We must prove that there is a Lebesgue integrable~$f$
such that $f_1(x),\,f_2(x),\,\dotsc$
converges to~$f(x)$ for almost all~$x\in\R$,
and $\Lphi(f) = \lim_n \Lphi(f_n)$.

By Statement~\eqref{eq:E:lebesgue-0},
we know that $f_1,\,f_2,\,\dotsc$
$\Lphi$-converges.
So by Theorem~\ref{T:lebesgue} we get that
$f_1,\,f_2,\,\dotsc$
is upper convergent
and that $\Lphi(\,\ulim{n}f_n\,) = \lim_n \Lphi(f_n)$.

So let us define $f\eqdf \ulim{n}f_n$.
To complete the proof,
we must show that $f_1(x),\,f_2(x),\,\dotsc$
converges to~$f(x)=\ulim{n}f_n(x)$
for almost all~$x\in \R$.
This is indeed the case since~$f_1(x),\,f_2(x),\,\dotsc$
converges for almost all~$x\in\R$.
\end{ex}

There is a small gap that needs to filled before
we continue with another topic.
Let $\varphi\colon L \ra E$ be a valuation.
We have defined what it means
for a sequence $a_1,\,a_2,\,\dotsc$ in~$L$ to be
$\varphi$-convergent
(see Definition~\ref{D:seq-phi-conv}),
but we have not yet given the meaning of
``$a_1,\,a_2,\,\dotsc$ converges \emph{to} $a$''.
We will do this in Definition~\ref{D:seq-phi-conv-to}.
%
%                  DEFINITION OF PHI-CONV TO AN ELEMENT
%
\begin{dfn}
\label{D:seq-phi-conv-to}
Let $L$ be a lattice.
Let $E$ be an ordered Abelian group.\\
Let $\varphi\colon L\ra E$ be a valuation.
Let $a_1,\,a_2,\,\dotsc$ be a sequence in~$L$.\\
Let $a\in L$ be given.
We say $a_1,\,a_2,\,\dotsc$
\keyword{$\varphi$-converges} to~$a$
provided that 
\begin{equation*}
a_1,\,a,\,a_2,\,a,\,\dotsc\qquad\text{is $\varphi$-convergent.}
\end{equation*}
\end{dfn}

Let $\varphi\colon L \ra E$ be a valuation.
While Definition~\ref{D:seq-phi-conv-to} is certainly reasonable,
it is also quite silly, 
and so one wonders if there is a more direct description
of when a sequence $\varphi$-converges to an element~$a\in L$.
If we assume~$\varphi$ is complete,
then there is a slightly 
better description (see Proposition~\ref{P:seq-phi-conv-to}).

In Section~\ref{S:fub} we study a notion of convergence
(see Definition~\ref{D:weak-phi-conv})
which was intended to be a more aesthetically pleasing
definition of $\varphi$-convergence,
but which turns out to be strictly weaker 
than $\varphi$-convergence
(see Example~\ref{E:weak-phi-conv}).
%
%                  EXAMPLE PHI-CONVERGENCE
%
\begin{prop}
\label{P:seq-phi-conv-to}
Let~$E$ be an ordered Abelian group.\\
Let~$L$ be a lattice, and $\varphi\colon L \ra E$ be a complete valuation.\\
Let $a_1,\,a_2,\,\dotsc$ be a $\varphi$-convergent sequence in~$L$,
and let $a\in L$.\\
Then $a_1,\,a_2,\,\dotsc$
$\varphi$-converges to~$a$
if and only if (see Definition~\ref{D:eq})
\begin{equation*}
a\ \approx\ \ulim{n}a_n.
\end{equation*}
(Recall that
$a_1,\,a_2,\,\dotsc$
is upper convergent
(see Definition~\ref{D:conv})
by Lemma~\ref{L:fatou}.)
\end{prop}
\begin{proof}
We leave this to the reader.
\end{proof}


\section{Valuation Systems}
\label{S:valuation-systems}
\subsection{Simple valuation systems}
%
%                  SYSTEMS
%
\begin{dfn}
\label{D:simple-system}
We say $\vs{V}{L}{\varphi}{E}$
 is a \keyword{simple valuation system}
provided that
\begin{enumerate}
\item \label{D:simple-system-1}
$V$ is a lattice
such that $\bw_n c_n$ and $\bv_n c_n$ exists for all~$c_1,c_2,\dotsc\in V$;
\item \label{D:simple-system-2}
$L$ is a sublattice of~$V$;
\item \label{D:simple-system-3}
$E$ is an ordered Abelian group;
\item \label{D:simple-system-4}
$\varphi\colon L\ra E$ is a valuation.
\end{enumerate}
\end{dfn}

%
%                  RING AS SIMPLE VALUATION SYSTEM
%
\begin{ex}
\label{E:ring-system}
Let $E$ be an ordered Abelian group.
Let~$X$ be a set, 
$\mathcal{A}$ a ring of sets,
and $\mu\colon \mathcal{A}\ra E$
a positive and additive map.
(See Example~\ref{E:ring-val}.)

Then we have the following simple valuation system.
\begin{equation*}
\vs{\wp X}{\mathcal{A}}\mu{E}
\end{equation*}
Indeed, $\wp X$ is lattice with 
$\bw_n A_n = \bigcap_n A_n$
and $\bv_n A_n = \bigcup_n A_n$ for all $A_i \in \wp X$,
$\mathcal{A}$ is a sublattice of~$\wp X$
by definition
and we have already seen 
that $\mu\colon \mathcal{A}\ra E$ is a valuation
(in Example~\ref{E:ring-val}).

In particular,
we have the following valuation systems.
\begin{equation*}
\vsLA \qquad\qquad \vsSA.
\end{equation*}
See Example~\ref{E:lmeas-val}) and  Example~\ref{E:smeas-val}.
\end{ex}

%
%                  RIESZ SPACE OF FUNCTIONS AS SIMPLE VALUATION SYSTEM
%
\begin{ex}
\label{E:riesz-function-space-simple-system}
Let $E$ be an ordered Abelian group.
Let $F$ be a Riesz space of functions on a set~$X$,
and $\varphi\colon F\ra E$ be a positive and linear map
(see Example~\ref{E:val-riesz-space-of-functions}).

Then we have the following simple valuation system.
\begin{equation*}
\vs{[-\infty,\infty]^X}{F}\varphi{E}
\end{equation*}
Indeed, $[-\infty,\infty]$ is a complete lattice
and hence so is $[-\infty,\infty]^X$.
Further, $F$ is a sublattice of~$\R^X$
which is in turn a sublattice of $[-\infty,\infty]^X$,
and we already know that
$\varphi$ is a valuation (see Example~\ref{E:val-riesz-space-of-functions}).
\end{ex}

\begin{ex}
Let $I=\{1,2\}$.
For each~$i\in I$,
let $\vs{V_i}{L_i}{\varphi_i}{E_i}$
be a simple valuation system.
Then we have the following simple valuation system
(see Example~\ref{E:val-product}).
\begin{equation*}
\vs{V_1\times V_2}{L_1 \times L_2}{\varphi_1 \times \varphi_2}{E_1 \times E_2}.
\end{equation*}
We call this system
the \emph{product} of $\vs{V_1}{L_1}{\varphi_1}{E_1}$
and $\vs{V_2}{L_2}{\varphi_2}{E_2}$.

Of course,
one can define a product of simple valuation systems
for any set~$I$.
\end{ex}

%
%                  NOTATION CONCERNING SUPREMA AND INFIMA IN L
%
\begin{nt}
Let $\vs{V}{L}\varphi{E}$ be a simple valuation system.
Let $a_1, a_2, \dotsc$ be from~$L$.
Then $a_1, a_2,\dotsc$ has a supremum
in~$V$ and might have a supremum in~$L$.
We ignore the latter:
\emph{With $\bv_n a_n$
we always mean the supremum of~$a_1, a_2,\dotsc $ in~$V$}.
\end{nt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%                  COMPLETENESS
%
%
\subsection{Completeness}
%
%                  COMPLETE SYSTEMS
%
\begin{dfn}
Let $\vs{V}{L}\varphi{E}$ be a simple valuation system.

Consider a sequence
$a_1 \geq a_2 \geq \dotsb$ from~$L$.
We say
\begin{equation*}
a_1 \geq a_2 \geq \dotsb \text{ is \keyword{$\varphi$-convergent}}
\qquad\text{if}\qquad \bw_n \varphi(a_n)\text{ exists.}
\end{equation*}

Similarly,
if
$b_1 \leq b_2 \leq \dotsb$ is
a sequence in~$L$, then 
\begin{equation*}
b_1 \leq b_2 \leq \dotsb \text{ is \keyword{$\varphi$-convergent}}
\qquad\text{if}\qquad \bv_n \varphi(b_n)\text{ exists.}
\end{equation*}
\end{dfn}
%
%                  COMPLETE VALUATION
%
\begin{dfn}
\label{D:system-complete}
Let $\vs{V}{L}\varphi{E}$ be a simple valuation system.\\
We say $\vs{V}{L}\varphi{E}$
is \keyword{complete}
(or simply, $\varphi$ is \keyword{complete})
 provided that
\begin{alignat*}{5}
a_1 \geq a_2 \geq \dotsb \text{$\varphi$-convergent }
  \quad &\implies \quad 
  & \bw_n a_n &\in L\quad 
  &&\text{and}\quad
  &\varphi(\,\bw_n a_n\,) &= \bw_n \varphi(a_n); \\
b_1 \leq b_2 \leq \dotsb \text{$\varphi$-convergent }
  \quad &\implies \quad 
  & \bv_n b_n &\in L\quad 
  &&\text{and}\quad
  &\varphi(\,\bv_n b_n\,) &= \bv_n \varphi(b_n).
\end{alignat*}
\end{dfn}
%
%                  THE LEBESGUE INTEGRAL IS COMPLETE
%
\begin{ex}
\label{E:complete-lint}
The Lebesgue integral
given us the  simple valuation system
\begin{equation*}
\vsLF
\end{equation*}
(see Example~\ref{E:int-val} and
Example~\ref{E:riesz-function-space-simple-system});
we will prove that this system is complete.

Let $f_1 \leq f_2 \leq \dotsb$
be a $\Lphi$-convergent sequence (in $\LF$).
We must prove that 
\begin{equation*}
\bv_n f_n \in\LF
\qquad\text{and}\qquad\Lphi(\bv_n f_n) = \bv_n \Lphi (f_n).
\end{equation*}
This follows immediately from Levi's Monotone Convergence Theorem.

Similarly,
if $g_1 \geq g_2 \geq \dotsb$
is a $\Lphi$-convergent sequence,
then
\begin{equation*}
\bw_n g_n \in\LF
\qquad\text{and}\qquad\Lphi(\bw_n g_n) = \bw_n \Lphi (g_n).
\end{equation*}
So the simple valuation system $\vsLF$ is complete.
\end{ex}
\todo{Add examples: the Lebesgue measure.
Maybe: add a description
of complete systems based on a ring of sets
and of complete systems based on a riesz space of functions.}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%                  SIGMA DISTRIBUTIVITY AND R COMPLETENESS
%
\subsection{$\sigma$-distributivity and $R$-completeness}
We would like to
extend a given simple valuation system to a complete one.
To make the theory work,
we concern ourselves only with certain
simple valuation systems,
which we just call ``valuation systems'' (see Definition~\ref{D:system}).
%
%                  ADDITIONAL RESTRICIONS ON THE SYSTE
% 
\begin{dfn}
\label{D:sigma-distributive}
Let~$V$ be a lattice.
We say~$V$ is
\keyword{$\sigma$-distributive}
provided that
\begin{alignat*}{3}
\bw_n c_n\text{ exists} \quad&\implies\quad
&a \vee \bw_n c_n \,&=\, \bw_n\  a\vee c_n, \\
\bv_n c_n\text{ exists} \quad&\implies\quad
&a \wedge \bv_n c_n \,&=\, \bv_n\  a\wedge c_n,
\end{alignat*}
where $a,\,c_1,c_2,\dotsc\in V$.
\end{dfn}
\begin{exs}
\begin{enumerate}
\item
Let $X$ be a set. Then $\wp(X)$ is $\sigma$-distributive.
Indeed,  
\begin{equation*}
\textstyle{
A \cup \bigcap_n C_n \,=\, \bigcap_n A \cup C_n
\qquad
A\cap \bigcup_n C_n \,=\, \bigcup_n A \cap C_n}
\end{equation*}
for all $A,\, C_1,C_2,\dotsc \subseteq X$.
Moreover,
we have already seen
that~$\wp(X)$ satisfies condition~\ref{D:simple-system-1}
of Definition~\ref{D:simple-system}
(see Example~\ref{E:ring-system}).
\item
Let $C$ be totally ordered. Then $C$ is $\sigma$-distributive.

Indeed,
let $a,\,c_1,c_2,\dotsc \in C$ be such that $\bw_n c_n$
exists. We need to prove that $a \vee \bw_n c_n$
is the supremum of~$a\vee c_1,\,a\vee c_2,\,\dotsc$.
To this end note that 
\begin{equation*}
b \leq d_1 \vee d_2 \quad\iff\quad 
b\leq d_1\quad\text{or}\quad b\leq d_2
\qquad\quad(b,d_i\in C).
\end{equation*}
(To see this, recall that $d_1 \vee d_2 = \max\{d_1,d_2\}$.)
Now, for $\ell \in C$, we have
\begin{alignat*}{3}
\forall n [\ \ell \leq a \vee c_n \ ]
\quad&\iff\quad
\ell \leq a
    \quad\text{or}\quad
    \forall n[\ \ell \leq c_n\ ] \\
\quad&\iff\quad
\ell \leq a
    \quad\text{or}\quad
    \ell \leq \bw_n c_n \\
\quad&\iff\quad
\ell \leq a\vee \bw_n c_n.
\end{alignat*}
So we see 
that $a\vee\bw_n c_n$ is the greatest 
lower bound of~$a\vee c_1,\,a\vee c_2\,\dotsc$.

With the same argument,
one can prove that $a \wedge \bv_n c_n = \bv a \wedge c_n$
for all $a,\,c_1,c_2,\dotsc \in C$
such that $\bv_n c_n$ exists.
Hence $C$ is $\sigma$-distributive.

\item
The lattice of the real numbers~$\R$ is a chain
and hence $\sigma$-distributive.

However,
$\R$ does not satisfy condition~\ref{D:simple-system-1}
of Definition~\ref{D:simple-system}.
Indeed,
a sequence $c_1,c_2,\dotsc$ in~$\R$ has a supremum
if and only if it is bounded from above,
i.e. there is an~$a\in \R$ such that $c_n \leq a$
for all~$n$.
Similarly,
a sequence
$c_1,c_2,\dotsc\in \R$ has an infimum
if and only if it is bounded from below.

\item
Let $\E$ be the lattice of the extended real numbers.
Then $\E$ is a chain and hence $\sigma$-distributive.
Moreover,
$\E$ also satisfies condition~\ref{D:simple-system-1}
of Definition~\ref{D:simple-system}.

\item
Let $I$ be a set,
and for each~$i\in I$,
let $L_i$ be a $\sigma$-distributive lattice.
Then the product $L\eqdf \prod_{i\in I} L_i$
is $\sigma$-distributive.
Moreover,
if $L_i$ satisfies condition~\ref{D:simple-system-1}
of Definition~\ref{D:simple-system}
for every~$i\in I$,
then $L$ satisfies condition~\ref{D:simple-system-1}
of Definition~\ref{D:simple-system}.

\item
Let $X$ be a set.
Then lattice $\EX$ of functions from~$X$ to~$\E$
is $\sigma$-distributive and satisfies
condition~\ref{D:simple-system-1}
of Definition~\ref{D:simple-system}.
\end{enumerate}
\end{exs}

\subsection{Valuation systems}
\begin{dfn}
\label{D:system}
A \keyword{valuation system} $\vs{V}{L}\varphi{E}$
is a simple valuation system
for which~$V$ is $\sigma$-distributive
and $E$ is $R$-complete.
\end{dfn}
\todo{Add examples of valuation systems.}

For the following it is convenient to `split' the notion
of complete valuation.
\begin{dfn}
\label{D:system-complete2}
Let $\vs{V}{L}\varphi{E}$ be
a valuation system.
\begin{enumerate}
\item
\label{D:system-complete2-pi}
We say $\vs{V}{L}\varphi{E}$ is \keyword{$\Pi$-complete}
provided that
\begin{alignat*}{5}
a_1 \geq a_2 \geq \dotsb \text{$\varphi$-convergent }
  \quad &\implies \quad 
  & \bw_n a_n &\in L\quad 
  &&\text{and}\quad
  &\varphi(\,\bw_n a_n\,) &= \bw_n \varphi(a_n). 
\end{alignat*}
\item
We say $\vs{V}{L}\varphi{E}$ is \keyword{$\Sigma$-complete}
provided that
\begin{alignat*}{5}
b_1 \leq b_2 \leq \dotsb \text{$\varphi$-convergent }
  \quad &\implies \quad 
  & \bv_n b_n &\in L\quad 
  &&\text{and}\quad
  &\varphi(\,\bv_n b_n\,) &= \bv_n \varphi(b_n). 
\end{alignat*}

\end{enumerate}
\end{dfn}

\end{document}
